### 2022-10-18 src/Draft_Docs.Rmd
trans MPI + clean MPI, edu + preliminary corr

---
title: "Draft"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-10-18 src/clean_edu.R

@@ -1,0 +1,39 @@
+library(dplyr)
+library(tidy)
+library(tibble)
+
+#custom plus operator to ignore NAs
+`%+%` <- function(x, y) {
+  mapply(sum, x, y, MoreArgs = list(na.rm = TRUE))
+}
+
+non.NA <- function(...)
+
+#import + clean data
+edu.expense <- read.csv('data/edu.percent.gdp.csv')
+edu.expense <- as_tibble(edu.expense) %>% 
+  dplyr::select(economy, country.code, edu.2016 = X2012, edu.2017 = X2013, edu.2018 = X2014)
+
+edu.expense <- edu.expense %>% mutate(no.avail = 3 - rowSums(is.na(.)), mean.edu = (edu.2016 %+% edu.2017 %+% edu.2018) / no.avail) %>% 
+  filter(!is.nan(mean.edu)) %>% 
+  dplyr::select(economy, country.code, mean.edu)
+
+# MPI with edu col
+dat <- fin.df %>%
+  dplyr::select(country.code, dep.mpi.diff, dep.mpi.diff.trans) %>% 
+  inner_join(edu.expense %>% 
+               dplyr::select(country.code, mean.edu), by = 'country.code') %>% 
+  inner_join(gdp %>% 
+               dplyr::select(country.code, mean.gdp), by = 'country.code')
+
+# rm.usa <- dat[-13,]
+# below.5e10 <- dat %>% 
+#  filter(mean.gdp < 1e12)
+
+below.5e10 <- dat
+lrm <- lm(dep.mpi.diff.trans ~ log(I(mean.edu * mean.gdp)), data = below.5e10)
+summary(lrm)
+plot(log(below.5e10$mean.gdp * below.5e10$mean.edu), below.5e10$dep.mpi.diff.trans)
+abline(lrm)
+plot(lrm, which=1)
+hist(resid(lrm))

### 2022-10-18 src/data_cleaning.R

@@ -1,0 +1,8 @@
+library(e1071)
+library(MASS)
+library(ggplot2)
+library(dplyr)
+library(tidy)
+library(tibble)
+
+# Clean MPI Data

### 2022-10-18 src/gdp.R

@@ -1,0 +1,9 @@
+gdp <- read.csv('data/gdp.csv')
+
+gdp <- as.tibble(gdp) %>% 
+  dplyr::select(economy, country.code, gdp.2016 = X2012, gdp.2017 = X2013, gdp.2018 = X2014)
+
+gdp <- gdp %>% mutate(no.avail = 3 - rowSums(is.na(.)), mean.gdp = (gdp.2016 %+% gdp.2017 %+% gdp.2018) / no.avail) %>% 
+  filter(!is.nan(mean.gdp)) %>% 
+  dplyr::select(economy, country.code, mean.gdp)
+

### 2022-10-18 src/transform_mpi.R

@@ -1,0 +1,89 @@
+library(e1071)
+library(MASS)
+library(ggplot2)
+library(dplyr)
+
+df.povert <- read.csv('data/mpi.14.19.csv')
+
+str(df.povert)
+
+no.avail <- df.povert %>%
+  select(country.code, economy, year) %>% 
+  group_by(year) %>% 
+  summarise(no.avail = n())
+
+hist(df.povert$year, breaks = 2011:2021)
+
+getYear <- function(yr) {
+    df.povert %>% 
+      select(country.code, economy, year, dep.monetary, dep.mpi) %>% 
+      filter(year == yr)
+}
+
+seg <- lapply(2012:2021, getYear)
+names(seg) <- 2012:2021
+
+noOverlap <- function(s1, s2) {
+  s1 %>%
+    inner_join(s2, by = "country.code") %>% 
+    filter(year.x != year.y) %>% 
+    summarise(count = n()) %>% 
+    pull()
+}
+
+cross <- diag(10)
+rownames(cross) <- 2012:2021
+colnames(cross) <- 2012:2021
+
+for (i in 1:10) {
+  cross[i,] = sapply(seg[1:10], noOverlap, s2 = seg[[i]])
+}
+
+cross
+
+max.cr <- max(cross)
+
+df.cross <- as.data.frame.table(cross)
+
+df.cross %>% 
+  filter(Freq == max.cr)
+
+fin.df <- seg[["2012"]] %>%
+  inner_join(seg[["2018"]], by = "country.code") %>% 
+  filter(year.x != year.y) %>% 
+  dplyr::select(country.code,
+         economy = economy.x,
+         dep.mpi.2016 = dep.mpi.x,
+         dep.mpi.2018 = dep.mpi.y) %>% 
+  mutate(dep.mpi.diff = dep.mpi.2018 - dep.mpi.2016)
+
+fin.df %>% 
+  summarise(pos = sum(dep.mpi.diff > 0),
+            neg = sum(dep.mpi.diff < 0),
+            unchg = sum(dep.mpi.diff == 0))
+
+summary(fin.df)
+
+hist(fin.df$dep.mpi.diff, breaks = 30)
+
+# transform
+y <- fin.df$dep.mpi.diff
+y <- y - min(y-0.01)
+
+skewness(y)
+shapiro.test(y)
+
+b <- boxcox(lm(y~1), lambda = seq(-3, 3, 1/1000))
[…]

### 2022-10-28 src/Draft_Docs.Rmd
first commit, although plenty are done

@@
 
-## R Markdown
+## Setting up
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+```{r, message = FALSE}
+library(e1071)
+library(MASS)
+library(ggplot2)
+library(dplyr)
+library(tidyverse)
+library(tibble)
+library(conflicted)
+library(knitr)
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+# everytime `select` is called, it will prefer dplyr function instead of MASS
+conflict_prefer(name = "select", winner = "dplyr")
+conflict_prefer(name = "filter", winner = "dplyr")
 ```
 
-## Including Plots
+## Data Cleaning
+### MPI
 
-You can also embed plots, for example:
+```{r}
+df.povert <- read.csv('../data/mpi.14.19.csv')
+str(df.povert)
+```
+How many data point we have each year?
+```{r}
+no.avail <- df.povert %>%
+  select(country.code, economy, year) %>% 
+  group_by(year) %>% 
+  summarise(`data points` = n())
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+kable(no.avail)
+
+hist(df.povert$year, breaks = 2010:2022)
+```
+We want to find difference of MPI in 2 separate year (i.e. y1 and y2) in each country. It's useful to know how many countries which were investigated in y1, were also investigated in y2.
+```{r}
+# Extract data from a specific year
+getYear <- function(yr) {
+    df.povert %>% 
+      select(country.code, economy, year, dep.monetary, dep.mpi) %>% 
+      filter(year == yr)
+}
+
+# segmented data by year (2012-2021)
+seg <- lapply(2012:2021, getYear)
+names(seg) <- 2012:2021
+str(seg)
+```
+Find overlapping countries
+```{r}
+# Calulate number of countries overlapping between s1 and s2
+noOverlap <- function(y1, y2) {
+  y1 %>%
+    inner_join(y2, by = "country.code") %>% 
+    filter(year.x != year.y) %>% 
+    summarise(count = n()) %>% 
+    pull()
+}
+
+# A 10x10 matrix to represent the no. of overlapping countries
+cross <- diag(10)
+rownames(cross) <- 2012:2021
+colnames(cross) <- 2012:2021
+
+# find no.of overlapping countries for each pair of years
+for (i in 1:10) {
+  cross[i,] = sapply(seg[1:10], noOverlap, y2 = seg[[i]])
+}
+
+cross
+```
+We find that each pair of years from 2012-2019 are fairly overlapping (mostly > 45). Therefore, taking any two years in this range will make a reasonable dataset for analysis.
+
+Next, we make a function to return the difference in MPI in any two year
+```{r}
+mpiDiff <- function(y1, y2) {
+  y1 <- as.character(y1)
+  y2 <- as.character(y2)
+
+  fin.df <- seg[[y1]] %>%
+    inner_join(seg[[y2]], by = "country.code") %>% 
+    filter(year.x != year.y) %>% 
+    select(country.code,
+           economy = economy.x,
+           mpi.y1 = dep.mpi.x,
+           mpi.y2 = dep.mpi.y) %>% 
+    mutate(mpi.diff = mpi.y1 - mpi.y2)
+
+  fin.df
+}
+
+# E.g. difference of mpi in 2012 and 2018
+mpi.12.18 <- mpiDiff(2012, 2018)
+head(mpi.12.18)
+```
+### Education
+This is a very rough processing and not very rigorous. For reference only
+
+First, do some summary on gdp (2012-2014)
+```{r}
+gdp <- read.csv('../data/gdp.csv')
+
+gdp <- as.tibble(gdp) %>% 
+  dplyr::select(economy, country.code, gdp.2012 = X2012, gdp.2013 = X2013, gdp.2014 = X2014)
+
+gdp <- gdp %>% mutate(no.avail = 3 - rowSums(is.na(.)), mean.gdp = (gdp.2012 %+% gdp.2013 %+% gdp.2014) / no.avail) %>% 
+  filter(!is.nan(mean.gdp)) %>% 
+  dplyr::select(economy, country.code, mean.gdp)
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+```{r}
+#custom plus operator to ignore NAs
+#e.g. 1 + NA = 1.
+`%+%` <- function(x, y) {
+  mapply(sum, x, y, MoreArgs = list(na.rm = TRUE))
+}
+
+#import + clean data
+edu.expense <- read.csv('../data/edu.percent.gdp.csv')
+edu.expense <- as_tibble(edu.expense) %>% 
+  dplyr::select(economy, country.code, edu2012 = X2012, edu2013 = X2013, edu2014 = X2014)
+
+#get mean %expenditure on education 2012-2014
+edu.expense <- edu.expense %>% mutate(no.avail = 3 - rowSums(is.na(.)), mean.edu = (edu2012 %+% edu2013 %+% edu2014) / no.avail) %>% 
+  filter(!is.nan(mean.edu)) %>% 
+  dplyr::select(economy, country.code, mean.edu)
+
+# MPI with edu col
+dat <- mpi.12.18 %>%
+  dplyr::select(country.code, mpi.diff) %>% 
+  inner_join(edu.expense %>% 
+               dplyr::select(country.code, mean.edu), by = 'country.code') %>% 
+  inner_join(gdp %>% 
+               dplyr::select(country.code, mean.gdp), by = 'country.code')
+head(dat)
+
+lrm <- lm(mpi.diff ~ log(I(mean.edu * mean.gdp)), data = dat)
+summary(lrm)
+plot(dat$mean.gdp * dat$mean.edu, dat$mpi.diff)
+plot(lrm, which=1)
+hist(resid(lrm))
+```

### 2022-10-28 src/outline.Rmd
first commit, although plenty are done

---
title: "Miscellanous"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Structure
1. We find the impact of gov expenditure effect on poverty using conventional definition of poverty
2. We redo the analysis using MPI as indicator of poverty
3. We conclude form the results obtained.
4. Some surprising correlation we experimented with, and how these can be a product of different reasons
  1. Gender equality and poverty
  2.

## Reference

## Data source
[MPM](https://www.worldbank.org/en/topic/poverty/brief/multidimensional-poverty-measure)
[mpi](https://data.worldbank.org/indicator/SI.POV.MDIM)
[pov215](https://data.worldbank.org/indicator/SI.POV.DDAY)

[Country.Class](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)
[GDP.PC](https://data.worldbank.org/indicator/NY.GDP.PCAP.PP.CD)
[Pri.edu](https://data.worldbank.org/indicator/SE.XPD.PRIM.ZS)
[Sec.edu](https://data.worldbank.org/indicator/SE.XPD.SECO.ZS)
[ter.edu](https://data.worldbank.org/indicator/SE.XPD.TERT.ZS)
[total.edu](https://data.worldbank.org/indicator/SE.XPD.TOTL.GD.ZS)
[health.exp](https://data.worldbank.org/indicator/SH.XPD.GHED.GD.ZS)
[milis.exp](https://data.worldbank.org/indicator/MS.MIL.XPND.GD.ZS)

[fdi](https://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD)
[unemp](https://data.worldbank.org/indicator/SL.UEM.TOTL.NE.ZS)
[labor.frc.part](https://data.worldbank.org/indicator/SL.TLF.CACT.NE.ZS)
[gender.equal](https://data.worldbank.org/indicator/IQ.CPA.GNDR.XQ)
[pop.growth](https://data.worldbank.org/indicator/SP.POP.GROW)
[urb.grw](https://data.worldbank.org/indicator/SP.URB.GROW)
[rural.grw](https://data.worldbank.org/indicator/SP.RUR.TOTL.ZG)
[inflation](https://data.worldbank.org/indicator/NY.GDP.DEFL.KD.ZG)
[gross capital formation](https://data.worldbank.org/indicator/NE.GDI.TOTL.ZS)
[trade](https://data.worldbank.org/indicator/NE.TRD.GNFS.ZS)
[reg.class](https://datatopics.worldbank.org/world-development-indicators/the-world-by-income-and-region.html)
[income.class](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)

### 2022-10-28 src/clean_gdp.R

@@ -1,0 +1,108 @@
+# gdp
+gdp <- read.csv('data/gdppc.csv')
+gdp <- gdp %>% 
+  select(economy, country.code, X2007:X2021) %>% 
+  pivot_longer(cols = X2007:X2021, names_to = 'year', values_to = 'gdp.per.capita') %>% 
+  mutate(year = as.numeric(substring(year, 2)))
+
+# edu
+edu.expense <- read.csv('data/education.csv')
+edu.expense <- edu.expense %>% 
+  select(economy, country.code, X2007:X2021) %>% 
+  pivot_longer(cols = X2007:X2021, names_to = 'year', values_to = 'edu.percent') %>% 
+  mutate(year = as.numeric(substring(year, 2)))
+
+# healthcare
+health.expense <- read.csv('data/healthcare.csv')
+health.expense <- health.expense %>% 
+  select(economy, country.code, X2007:X2021) %>% 
+  pivot_longer(cols = X2007:X2021, names_to = 'year', values_to = 'health.percent') %>% 
+  mutate(year = as.numeric(substring(year, 2)))
+
+# military
+mili.expense <- read.csv('data/military.csv')
+mili.expense <- mili.expense %>% 
+  select(economy, country.code, X2007:X2021) %>% 
+  pivot_longer(cols = X2007:X2021, names_to = 'year', values_to = 'mili.percent') %>% 
+  mutate(year = as.numeric(substring(year, 2)))
+
+# join data
+df.expenditure <- gdp %>% 
+  inner_join(edu.expense, by = c('economy', 'country.code', 'year')) %>%
+  inner_join(health.expense, by = c('economy', 'country.code', 'year')) %>%
+  inner_join(mili.expense, by = c('economy', 'country.code', 'year')) %>%
+  mutate(edu.expenditure = edu.percent * gdp.per.capita / 100,
+         health.expenditure = health.percent * gdp.per.capita / 100, 
+         mili.expenditure = mili.percent * gdp.per.capita / 100) 
+
+
+# functions
+eduNYear <- function(y1, y2) {
+  df.expenditure %>% 
+    filter(year >= y1 & year <= y2) %>%
+    filter(!is.na(edu.expenditure)) %>% 
+    group_by(economy, country.code) %>% 
[…]

### 2022-10-28 src/clean_health.R

@@ -1,0 +1,21 @@
+healthNYear <- function(y1, y2) {
+  df.expenditure %>% 
+    filter(year >= y1 & year <= y2) %>%
+    filter(!is.na(health.expenditure)) %>% 
+    group_by(economy, country.code) %>% 
+    summarise(mean.health.expenditure = mean(health.expenditure, rm.na = TRUE)) %>% 
+    mutate(year = y1, distance = y2 - y1)
+}
+
+healthNYear2 <- function(noYear) {
+  y.min = min(df.expenditure$year)
+  y.max = max(df.expenditure$year) - noYear
+  
+  if (y.min < y.max){
+    list <- lapply(y.min:y.max, \(start.year) {
+      healthNYear(start.year, start.year + noYear)
+    })
+  }
+  
+  Reduce(rbind, list)
+}

### 2022-10-28 src/clean_mili.R

@@ -1,12 +1,4 @@
-library(dplyr)
-library(tidyverse)
-library(tidyr)
-library(tibble)
-
-options(dplyr.summarise.inform = FALSE)
-# edu
-
-eduNYear <- function(y1, y2) {
+miliNYear <- function(y1, y2) {
   df.expenditure %>% 
     filter(year >= y1 & year <= y2) %>%
     filter(!is.na(edu.expenditure)) %>% 
@@ -27,4 +19,3 @@ eduNYear2 <- function(noYear) {
   
   Reduce(rbind, list)
 }
-

### 2022-10-28 src/clean_mpi.R

@@ -1,0 +1,50 @@
+library(e1071)
+library(MASS)
+library(ggplot2)
+library(dplyr)
+library(tidyverse)
+library(tibble)
+library(conflicted)
+library(knitr)
+
+# everytime `select` is called, it will prefer dplyr function instead of MASS
+conflict_prefer(name = "select", winner = "dplyr")
+conflict_prefer(name = "filter", winner = "dplyr")
+
+df.povert <- tibble(read.csv('data/mpi.14.19.csv'))
+
+getYear <- function(yr) {
+  df.povert %>% 
+    filter(year == yr)
+}
+
+mpiDiff <- function(y1, y2) {
+  df.y1 <- getYear(y1)
+  df.y2 <- getYear(y2)
+  
+  fin.df <- df.y1 %>%
+    inner_join(df.y2, by = c("country.code", "economy")) %>% 
+    mutate(mpi.diff = dep.mpi.y - dep.mpi.x,
+           distance.mpi = year.y - year.x,
+           mpi.diff.percent = mpi.diff/dep.mpi.x*100,
+           edu.att.diff = edu.attainment.y - edu.attainment.x,
+           edu.enr.diff = edu.enrollment.y - edu.enrollment.x,
+           mon.diff = dep.monetary.y - dep.monetary.x) %>% 
+    select(economy, country.code, year = year.x, distance.mpi, mpi.diff, edu.att.diff, edu.enr.diff, mpi.diff.percent, mon.diff)
+  
+  fin.df
+}
+
+mpiDiff2 <- function(distance) {
+  y.min = 2012
+  y.max = 2022 - distance
+  
+  if (y.min <= y.max) {
+    list <- lapply(y.min:y.max, \(start.year){
+      mpiDiff(start.year, start.year + distance)
+    })
+  }
+  
+  Reduce(rbind, list)
+}
+

### 2022-10-28 src/model.R

@@ -1,0 +1,55 @@
+mpi6 <- mpiDiff2(6)
+edu3 <- eduNYear2(1)
+health3 <- healthNYear2(1)
+mili3 <- miliNYear2(3)
+
+fin.df <- mpi6 %>% 
+  inner_join(edu3, by = c('economy', 'country.code', 'year')) %>% 
+  inner_join(health3, by = c('economy', 'country.code', 'year')) %>% 
+  inner_join(mili3, by = c('economy', 'country.code', 'year'))
+
+lrm <- lm(mpi.diff ~ log(mean.edu.expenditure), data = fin.df)
+summary(lrm)
+
+lrm <- lm(mpi.diff ~ sqrt(mean.health.expenditure) + mean.health.expenditure, data = fin.df)
+summary(lrm)
+
+lrm <- lm(mpi.diff ~ sqrt(mean.mili.expenditure), data = fin.df)
+summary(lrm)
+
+lrm <- lm(mpi.diff ~ mean.edu.expenditure + log(mean.edu.expenditure), data = fin.df)
+summary(lrm)
+
+ggplot(fin.df, aes(-0.40588 * log(mean.edu.expenditure)^2 + 6.26488 * log(mean.edu.expenditure), mpi.diff)) +
+  geom_point() +
+  geom_smooth(method = 'lm')
+
+car::vif(lrm)
+
+plot(lrm, which = 1)
+
+# getRSquared <- function(mpiYear, eduYear, healthYear) {
+#   mpi <- mpiDiff2(mpiYear)
+#   edu <- eduNYear2(eduYear)
+#   health <- healthNYear2(healthYear)
+#   
+#   fin.df <- mpi %>% 
+#     inner_join(edu, by = c('economy', 'country.code', 'year'), suffix = c('.mpi', '.edu')) %>% 
+#     inner_join(health3, by = c('economy', 'country.code', 'year'), suffix = c('', '.health'))
+#   
+#   lrm <- lm(mpi.diff ~ mean.health.expenditure + log(mean.edu.expenditure), data = fin.df)
+#   
+#   summary(lrm)$adj.r.squared
+# }
+# 
+# assessment <- data.frame(mpi = numeric(), edu = numeric(), health = numeric(), r.sqr = numeric())
+# 
+# 
+# m <- 6
+# for (e in 1:4) {
+#   for (h in 1:4) {
+#     rq <- getRSquared(m, e, h)
+#     print(rq)
[…]

### 2022-10-28 src/transform_mpi.R

@@ -1,89 +1,0 @@
-library(e1071)
-library(MASS)
-library(ggplot2)
-library(dplyr)
-
-df.povert <- read.csv('data/mpi.14.19.csv')
-
-str(df.povert)
-
-no.avail <- df.povert %>%
-  select(country.code, economy, year) %>% 
-  group_by(year) %>% 
-  summarise(no.avail = n())
-
-hist(df.povert$year, breaks = 2011:2021)
-
-getYear <- function(yr) {
-    df.povert %>% 
-      select(country.code, economy, year, dep.monetary, dep.mpi) %>% 
-      filter(year == yr)
-}
-
-seg <- lapply(2012:2021, getYear)
-names(seg) <- 2012:2021
-
-noOverlap <- function(s1, s2) {
-  s1 %>%
-    inner_join(s2, by = "country.code") %>% 
-    filter(year.x != year.y) %>% 
-    summarise(count = n()) %>% 
-    pull()
-}
-
-cross <- diag(10)
-rownames(cross) <- 2012:2021
-colnames(cross) <- 2012:2021
-
-for (i in 1:10) {
-  cross[i,] = sapply(seg[1:10], noOverlap, s2 = seg[[i]])
-}
-
-cross
-
-max.cr <- max(cross)
-
-df.cross <- as.data.frame.table(cross)
-
-df.cross %>% 
-  filter(Freq == max.cr)
-
-fin.df <- seg[["2012"]] %>%
-  inner_join(seg[["2018"]], by = "country.code") %>% 
-  filter(year.x != year.y) %>% 
-  dplyr::select(country.code,
-         economy = economy.x,
-         dep.mpi.2016 = dep.mpi.x,
-         dep.mpi.2018 = dep.mpi.y) %>% 
-  mutate(dep.mpi.diff = dep.mpi.2018 - dep.mpi.2016)
-
-fin.df %>% 
-  summarise(pos = sum(dep.mpi.diff > 0),
-            neg = sum(dep.mpi.diff < 0),
-            unchg = sum(dep.mpi.diff == 0))
-
-summary(fin.df)
-
-hist(fin.df$dep.mpi.diff, breaks = 30)
-
-# transform
-y <- fin.df$dep.mpi.diff
-y <- y - min(y-0.01)
-
-skewness(y)
-shapiro.test(y)
-
-b <- boxcox(lm(y~1), lambda = seq(-3, 3, 1/1000))
[…]

### 2022-10-30 src/draft.Rmd
draft + citation file added

---
title: "Draft Documents"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: citations.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2. Data Characteristic

## 2.1. Nature of Data

The data set is collection The World Bank Data, the variables of interest are extracted from the raw data files and combined into a single data frame for analysis. The final data set includes:

1.  **country.code**: Country code\
2.  **country.name**: Country name\
3.  **year**: Year\
4.  **income**: Income class
    -   Low income (L)\
    -   Lower middle income (LM)\
    -   Upper middle income (UM)\
    -   High income (H)\
5.  **reg**: Region\
6.  **pov**: Poverty headcount ratio\
7.  **mpi**: Multidimensional Poverty Index\
8.  **edu.total**: Total expenditure on education (% of GDP)\
9.  **edu.pri**: Total expenditure on primary education (% of total education expenditure)\
10. **edu.sec**: Total expenditure on secondary education (% of total education expenditure)\
11. **edu.ter**: Total expenditure on tertiary education (% of total education expenditure)\
12. **hlth**: Total expenditure on health (% of GDP)\
13. **mil**: Total expenditure on military (% of GDP)\
14. **fdi**: Foreign Direct Investment\
15. **lbr.part**: Labour force participation (% of population ages 15+)\
16. **unemp**: Unemployment rate\
17. **pop.gwth.total**: Total population growth rate\
18. **pop.gwth.rural**: Total rural population growth rate\
19. **pop.gwth.urban**: Total urban population growth rate\
20. **gdp.dflt**: GDP deflator\
21. **gdr.eql**: Gender equality rating\
22. **gcf**: Gross Capital Formation\
23. **trade**: Trade = import + export (% of GDP)

Data imports and combining:

```{r, message = F}
# required library
library(knitr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

```{r}
# helper functions
importWDI <- function(filepath, value_name) {
  df <- read_csv(filepath, skip = 4)

  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))

  df <- df %>% 
    pivot_longer(5:ncol(.), names_to = "year", values_to = "value") %>% 
    filter(!is.null(value) & !is.na(value)) %>% 
    mutate(country.code = factor(country.code), 
           country.name = factor(country.name),
           year = as.numeric(year)) %>% 
    select(country.code, country.name, year, value)

  colnames(df)[4] <- value_name

  df
}

importRegionClass <- function(filepath) {
  df <- read_csv(filepath, skip = 4)

  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))

  df %>% mutate(country.name = factor(country.name),
                region = factor(region)) %>% 
    select(country.name, reg = region)
}

importIncomeClass <- function(filepath) {
  df <- read_csv(filepath, skip = 4)

  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))

  df %>% 
    pivot_longer(3:ncol(.), names_to = "year", values_to = "income") %>% 
    filter(!is.null(income) & !is.na(income)) %>% 
    mutate(country.code = factor(country.code), 
           country.name = factor(country.name), 
           year = as.numeric(year), 
           income = factor(income)) %>% 
    select(country.code, country.name, year, income)
}
```

```{r, message = F}
# import and combine data
setwd("../data")

poverty.headcount <- importWDI("poverty.headcount.215dollar.csv", "pov")
mpi <- importWDI("mpi.csv", "mpi")
education.expenditure.total <- importWDI("total.education.expenditure.csv", "edu.total")
education.expenditure.primary <- importWDI("primary.education.expenditure.csv", "edu.pri")
education.expenditure.secondary <- importWDI("secondary.education.expenditure.csv", "edu.sec")
education.expenditure.tertiary <- importWDI("tertiary.education.expenditure.csv", "edu.ter")
health.expenditure <- importWDI("health.expenditure.csv", "hlth")
military.expenditure <- importWDI("military.expenditure.csv", "mil")
fdi <- importWDI("fdi.csv", "fdi")
labour.force.participation <- importWDI("labour.force.participation.csv", "lbr.part")
unemployment.rate <- importWDI("unemployment.csv", "unemp")
population.growth <- importWDI("population.growth.csv", "pop.gwth.total")
rural.population.growth <- importWDI("rural.population.growth.csv", "pop.gwth.rural")
urban.population.growth <- importWDI("urban.population.growth.csv", "pop.gwth.urban")
gdp.deflator <- importWDI("gdp.deflator.csv", "gdp.dflt")
gender.equality <- importWDI("gender.equality.csv", "gdr.eql")
gross.capital.formation <- importWDI("gross.capital.formation.csv", "gcf")
trade <- importWDI("trade.csv", "trade")
region.class <- importRegionClass("region.class.csv")
income.class <- importIncomeClass("income.class.csv")

setwd("../src")

countries <- income.class %>% 
  full_join(region.class, by = "country.name") %>% 
  right_join(poverty.headcount, by = c("country.name", "country.code", "year")) %>%
  full_join(mpi, by = c("country.name", "country.code", "year")) %>%
  left_join(education.expenditure.total, by = c("country.name", "country.code", "year")) %>%
  left_join(education.expenditure.primary, by = c("country.name", "country.code", "year")) %>%
  left_join(education.expenditure.secondary, by = c("country.name", "country.code", "year")) %>%
  left_join(education.expenditure.tertiary, by = c("country.name", "country.code", "year")) %>%
  left_join(health.expenditure, by = c("country.name", "country.code", "year")) %>%
  left_join(military.expenditure, by = c("country.name", "country.code", "year")) %>%
  left_join(fdi, by = c("country.name", "country.code", "year")) %>%
  left_join(labour.force.participation, by = c("country.name", "country.code", "year")) %>%
  left_join(unemployment.rate, by = c("country.name", "country.code", "year")) %>%
  left_join(population.growth, by = c("country.name", "country.code", "year")) %>%
  left_join(rural.population.growth, by = c("country.name", "country.code", "year")) %>%
  left_join(urban.population.growth, by = c("country.name", "country.code", "year")) %>%
  left_join(gdp.deflator, by = c("country.name", "country.code", "year")) %>%
  left_join(gender.equality, by = c("country.name", "country.code", "year")) %>%
  left_join(gross.capital.formation, by = c("country.name", "country.code", "year")) %>%
  left_join(trade, by = c("country.name", "country.code", "year"))
```

Data preview

```{r}
head(countries)
str(countries)
summary(countries)
```

## 2.2. Missing values

As observed from the summary above, the data set contains a lot of missing values in some of the variables.

```{r}
nCompleteObs <- sum(complete.cases(countries))
print(paste("No. of complete cases:", nCompleteObs))
```

There are only 3 complete cases where all the variable is available. This is nowhere near acceptable to conduct any meaningful analysis. Therefore, we need to eliminate some variables for a more balance data set.

```{r}
mean(is.na(countries))
```

About 23% of the data set is missing.

```{r}
missings <- colMeans(is.na(countries))
ggplot(mapping = aes(x = names(missings), y = missings, fill = missings < 0.35)) +
  geom_bar(stat = "identity") +
  ggtitle("Missing rate") +
  xlab("variables") +
  ylab("% missing") +
  theme(axis.text.x = element_text(size=9, angle=90))
```

There are 5 variables with missing rate \>35%.

```{r}
missings[missings > 0.35]
```

These can be very useful and relevant information [@pubspend&pov]. However, we would like to exclude these variables from some first analyses to make use of the richer set of data. We can conduct a separate analysis with these variable to gain more insight.

```{r}
# variables with high missing rate
hMiss <- names(missings[missings > 0.35])
countries1 <- countries %>% select(!hMiss)
str(countries1)
```

Re-evaluate the `countries1` set.

```{r}
mean(is.na(countries1))
sum(complete.cases(countries1))
mean(complete.cases(countries1))
```

On average, each column has 10% missing rate, that results in 916 complete data point (i.e. 38%).\
This can be a sufficient number for the analysis. However, the missing data can induce loss of power due to the reduced sample size, and some other biases depending on which variables is missing. 
```{r}
countries1 %>% 
  group_by(reg) %>% 
  summarise(mean.complete = mean(complete.cases(.)))
```
The majority of missing values might be attribute to the inaccessibility of some economies, where data was not easy to be collected (e.g. Middle East & North Africa). Ironically, these countries with

### 2022-10-30 src/outline.Rmd
draft + citation file added

@@
 ```
-
-## Structure
+## Scrap
 1. We find the impact of gov expenditure effect on poverty using conventional definition of poverty
@@
   2.
+
+Refer to [guide to data analysis report](https://instruction.bus.wisc.edu/jfrees/jfreesbooks/Regression%20Modeling/BookWebDec2010/Writing/Chap20FreesRegBook03May2009.pdf)
+
+## Structure
+1. Introduction
+2. Data Characteristic
+  6. Nature of data
+  2. Dictionary
+  3. Missing values
+    - Imbalance data, removing will be costly since a large portion is missing -> Imputation
+  4. Descriptive Analytics
+  5. Correlation
+  1. Source
+3. Model selection and interpretation
+  1. reference to other studies
+  2. Nature of data -> choice of model
+  3. plotting, foreshadowing the relationship
+4. Results
+5. Conclusion
+6. Appendix
 

### 2022-10-30 src/DataCleaning.R

@@ -1,0 +1,79 @@
+library(corrplot)
+library(PerformanceAnalytics)
+library(Hmisc)
+library(ggplot2)
+
+# Import and combine
+setwd("data")
+
+poverty.headcount <- importWDI("poverty.headcount.215dollar.csv", "pov")
+mpi <- importWDI("mpi.csv", "mpi")
+education.expenditure.total <- importWDI("total.education.expenditure.csv", "edu.total")
+education.expenditure.primary <- importWDI("primary.education.expenditure.csv", "edu.pri")
+education.expenditure.secondary <- importWDI("secondary.education.expenditure.csv", "edu.sec")
+education.expenditure.tertiary <- importWDI("tertiary.education.expenditure.csv", "edu.ter")
+health.expenditure <- importWDI("health.expenditure.csv", "hlth")
+military.expenditure <- importWDI("military.expenditure.csv", "mil")
+fdi <- importWDI("fdi.csv", "fdi")
+labour.force.participation <- importWDI("labour.force.participation.csv", "lbr.part")
+unemployment.rate <- importWDI("unemployment.csv", "unemp")
+population.growth <- importWDI("population.growth.csv", "pop.gwth.total")
+rural.population.growth <- importWDI("rural.population.growth.csv", "pop.gwth.rural")
+urban.population.growth <- importWDI("urban.population.growth.csv", "pop.gwth.urban")
+gdp.deflator <- importWDI("gdp.deflator.csv", "gdp.dflt")
+gender.equality <- importWDI("gender.equality.csv", "gdr.eql")
+gross.capital.formation <- importWDI("gross.capital.formation.csv", "gcf")
+trade <- importWDI("trade.csv", "trade")
+region.class <- importRegionClass("region.class.csv")
+income.class <- importIncomeClass("income.class.csv")
+
+setwd("..")
+
+countries <- income.class %>% 
+  full_join(region.class, by = "country.name") %>% 
+  right_join(poverty.headcount, by = c("country.name", "country.code", "year")) %>%
+  full_join(mpi, by = c("country.name", "country.code", "year")) %>%
+  left_join(education.expenditure.total, by = c("country.name", "country.code", "year")) %>%
+  left_join(education.expenditure.primary, by = c("country.name", "country.code", "year")) %>%
+  left_join(education.expenditure.secondary, by = c("country.name", "country.code", "year")) %>%
[…]

### 2022-10-30 src/helperfunctions.R

@@ -1,0 +1,47 @@
+library(readr)
+library(tidyr)
+library(dplyr)
+
+importWDI <- function(filepath, value_name) {
+  df <- read_csv(filepath, skip = 4) 
+  
+  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))
+  
+  df <- df %>% 
+    pivot_longer(5:ncol(.), names_to = "year", values_to = "value") %>% 
+    filter(!is.null(value) & !is.na(value)) %>% 
+    mutate(country.code = factor(country.code), 
+           country.name = factor(country.name),
+           year = as.numeric(year)) %>% 
+    select(country.code, country.name, year, value)
+  
+  colnames(df)[4] <- value_name
+  
+  df
+}
+
+importRegionClass <- function(filepath) {
+  df <- read_csv(filepath, skip = 4) 
+  
+  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))
+  
+  df %>% mutate(country.name = factor(country.name),
+                region = factor(region)) %>% 
+    select(country.name, reg = region)
+}
+
+importIncomeClass <- function(filepath) {
+  df <- read_csv(filepath, skip = 4) 
+  
+  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))
+  
+  df %>% 
+    pivot_longer(3:ncol(.), names_to = "year", values_to = "income") %>% 
+    filter(!is.null(income) & !is.na(income)) %>% 
+    mutate(country.code = factor(country.code), 
+           country.name = factor(country.name), 
+           year = as.numeric(year), 
+           income = factor(income)) %>% 
+    select(country.code, country.name, year, income)
+}
+

### 2022-10-30 src/draft.Rmd
NA

@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
+```
+
+```{r, message = F}
+# required library
+library(knitr)
+library(readr)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
 ```
@@
 Data imports and combining:
-
-```{r, message = F}
-# required library
-library(knitr)
-library(readr)
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-```
 
@@
 ```{r, message = F}
-# import and combine data
+# import data
 setwd("../data")
@@
 setwd("../src")
+```
 
-countries <- income.class %>% 
-  full_join(region.class, by = "country.name") %>% 
-  right_join(poverty.headcount, by = c("country.name", "country.code", "year")) %>%
+We found that the data sets collected from [World Bank's Data helpdesk](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups) and [The World Bank's Data](https://data.worldbank.org/) have different naming convention for certain countries (e.g.
+"Czechia" vs.
+"Czechnia Republic").
+So we need to rename these countries to avoid some error when joining.
+
+Furthermore, WDI's data sets rate also account for non-country (e.g.
+country.name = "Low income" or "South Asia").
+These special groups are not in our scope of interest, which is national, so we eliminate them.
+
+```{r}
+# using poverty.headcount as a naming standard (as other data from WDI also use this convention)
+# join a subset of data to process the names
+d <- poverty.headcount %>%
+  select(country.name) %>%
+  mutate(inPov = T) %>%  
+  full_join(income.class %>% 
+              select(country.name) %>% 
+              mutate(inIncome = T), by = "country.name") %>%  
+  full_join(region.class %>%
+              select(country.name) %>% 
+              mutate(inReg = T), by = "country.name") %>% 
+  mutate(inPov = !is.na(inPov), inIncome = !is.na(inIncome), inReg = !is.na(inReg))
+
+d
+```
+
+First, remove special economic groups from `poverty.headcount`.
+We figured these regions will not appear in `income.class` or `region.class`, so we might find something from looking at the countries **only** appear in `poverty.headcount`.
+
+```{r}
+d %>% filter(inPov & (!inIncome | !inReg)) %>% distinct(country.name)
+```
+
+Lucky! We can look through these 18 results and compose a list of special regions.
+
+```{r}
+spec.reg <- c("Fragile and conflict affected situations", "IDA total", "World", "East Asia & Pacific", "Europe & Central Asia", "Latin America & Caribbean", "Middle East & North Africa", "South Asia", "Sub-Saharan Africa", "Low income", "Low & middle income", "Lower middle income", "Upper middle income", "High income")
+```
+
+Then, we rename those countries with inconsistent naming convention. Since we should only care about countries whose poverty headcount is available, reusing the list generated above, we can identify:
+
+1.  Cote d'Ivoire (also Côte d'Ivoire)
+2.  Czechia (also Czechoslovakia or Czech Republic)
+3.  Curacao (also Curaçao)
+4.  Turkiye (formerly known as Turkey, also Türkiye)
+5.  Sao Tome and Principe (also São Tomé and Príncipe)
+
+```{r}
+# mapping standard name and variation
+nameMap <- tibble(standard = c("Cote d'Ivoire", "Czechia", "Czechia", "Curacao", "Turkiye", "Turkiye", "Sao Tome and Principe"), variation = c("Côte d'Ivoire", "Czechoslovakia", "Czech Republic", "Curaçao", "Turkey", "Türkiye", "São Tomé and Príncipe"))
+
+correctName <- function(name) {
+  tibble(name = name) %>% 
+    left_join(nameMap, by = c("name" = "variation")) %>% 
+    mutate(standard = ifelse(is.na(standard), name, standard)) %>% 
+    select(standard) %>% 
+    pull()
+}
+
+orig.name <- c("Vietnam", "China", "Turkey", "Czechia Republic")
+correctName(orig.name)
+```
+
+Let's test this out!
+
+```{r}
+d <- poverty.headcount %>%
+  # correct name here
+  mutate(country.name = correctName(country.name)) %>% 
+  select(country.name) %>%
+  mutate(inPov = T) %>%
+  full_join(income.class %>% 
+              # correct name here
+              mutate(country.name = correctName(country.name)) %>% 
+              select(country.name) %>% 
+              mutate(inIncome = T), by = "country.name") %>%  
+  full_join(region.class %>%
+              # correct name here
+              mutate(country.name = correctName(country.name)) %>% 
+              select(country.name) %>% 
+              mutate(inReg = T), by = "country.name") %>% 
+  filter(!(country.name %in% spec.reg)) %>% 
+  mutate(inPov = !is.na(inPov), inIncome = !is.na(inIncome), inReg = !is.na(inReg))
+
+# countries not in region list, but is in Pov list
+d %>% 
+  filter(!inReg & inPov) %>% 
+  distinct(country.name) %>% 
+  nrow()
+# countries not in income list, but is in Pov list
+d %>% 
+  filter(!inIncome & inPov) %>% 
+  distinct(country.name) %>% 
+  nrow()
+```
+
+We are *pretty* confident that there's no inconsistent naming left unprocessed in the data sets.
+
+```{r}
+# Rename countries in all data sets.
+poverty.headcount <- poverty.headcount %>%
+  mutate(country.name = correctName(country.name))
+mpi <- mpi %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.total <- education.expenditure.total %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.primary <- education.expenditure.primary %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.secondary <- education.expenditure.secondary %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.tertiary <- education.expenditure.tertiary %>%
+  mutate(country.name = correctName(country.name))
+health.expenditure <- health.expenditure %>%
+  mutate(country.name = correctName(country.name))
+military.expenditure <- military.expenditure %>%
+  mutate(country.name = correctName(country.name))
+fdi <- fdi %>%
+  mutate(country.name = correctName(country.name))
+labour.force.participation <- labour.force.participation %>%
+  mutate(country.name = correctName(country.name))
+unemployment.rate <- unemployment.rate %>%
+  mutate(country.name = correctName(country.name))
+population.growth <- population.growth %>%
+  mutate(country.name = correctName(country.name))
+rural.population.growth <- rural.population.growth %>%
+  mutate(country.name = correctName(country.name))
+urban.population.growth <- urban.population.growth %>%
+  mutate(country.name = correctName(country.name))
+gdp.deflator <- gdp.deflator %>%
+  mutate(country.name = correctName(country.name))
+gender.equality <- gender.equality %>%
+  mutate(country.name = correctName(country.name))
+gross.capital.formation <- gross.capital.formation %>%
+  mutate(country.name = correctName(country.name))
+trade <- trade %>%
+  mutate(country.name = correctName(country.name))
+region.class <- region.class %>%
+  mutate(country.name = correctName(country.name))
+income.class <- income.class %>%
+  mutate(country.name = correctName(country.name))
+```
+
+Join the data
+
+```{r, message = F}
+countries <- poverty.headcount %>% 
+  # We used a full join here so we can conduct a separate analysis on mpi later
   full_join(mpi, by = c("country.name", "country.code", "year")) %>%
+  left_join(income.class, c("country.name", "country.code", "year")) %>% 
+  left_join(region.class, by = "country.name") %>%
   left_join(education.expenditure.total, by = c("country.name", "country.code", "year")) %>%
@@
   left_join(gross.capital.formation, by = c("country.name", "country.code", "year")) %>%
-  left_join(trade, by = c("country.name", "country.code", "year"))
+  left_join(trade, by = c("country.name", "country.code", "year")) %>% 
+  # filter special groups
+  filter(!(country.name %in% spec.reg))
 ```
@@
 
+There's no NA in `reg`, which is a sign that all naming in the data is remedied.
+There's some expected NAs in `income` and `pov`, as these data are collected by year.
+There's a substantial amount of missing data in `mpi`, as this is a relative new concept.
+We will address the nature, and processing of missing data in the next sections.
+
 ## 2.2. Missing values
@@
 As observed from the summary above, the data set contains a lot of missing values in some of the variables.
+
+```{r}
+mean(is.na(countries))
+```
+
+About 19% of the data set is missing.
 
@@
 ```{r}
-mean(is.na(countries))
-```
-
-About 23% of the data set is missing.
-
-```{r}
 missings <- colMeans(is.na(countries))
@@
   theme(axis.text.x = element_text(size=9, angle=90))
+
+missings[missings > 0.35]
 ```
@@
 There are 5 variables with missing rate \>35%.
-
-```{r}
-missings[missings > 0.35]
-```
 
@@
 hMiss <- names(missings[missings > 0.35])
+# exclude these variables in countries1
 countries1 <- countries %>% select(!hMiss)
@@
 
-On average, each column has 10% missing rate, that results in 916 complete data point (i.e. 38%).\
-This can be a sufficient number for the analysis. However, the missing data can induce loss of power due to the reduced sample size, and some other biases depending on which variables is missing. 
+On average, each column has 6% missing rate, results in 937 complete data point (i.e.
+49%).
+This can be a sufficient number for the analysis.
+However, the missing data can induce loss of power due to the reduced sample size, and some other biases depending on which variables is missing.
+
 ```{r}
+# complete rate of data by regions
 countries1 %>% 
+  mutate(isComplete = complete.cases(.)) %>% 
   group_by(reg) %>% 
-  summarise(mean.complete = mean(complete.cases(.)))
+  summarise(complete.rate = mean(isComplete)) %>% 
+  arrange(desc(complete.rate))
 ```
-The majority of missing values might be attribute to the inaccessibility of some economies, where data was not easy to be collected (e.g. Middle East & North Africa). Ironically, these countries with
+
+Countries from North America, Sub-Saharan Africa, and South Asia have the highest rate of missing data.
+We suspect that Sub-Saharan Africa, and South Asia are comparably less accessible regions.
+We also know that Americans don't like filling out forms, so their high rate of missing data is understandable as well.
+
+Still, we need to find a way to address this issue. we propose several approaches:
+
+1.  **Use complete cases**: Only use the complete cases for the analysis. This is a straightforward approach, but doesn't resolve the bias resulted from the mass loss of data.
+2.  **Step-by-step remove frequently missing variables**: The same as we did before, but this process should be carried out carefully as we run the chance of dropping an important variable.
+3.
+**Update the data set as we select variables**: As we drop insignificant variables (in backward selection), the number of NAs are changed as well.
+We can utilize the extra complete cases to build the next model in the steps.
+4.
+**Imputation**: The idea is to replace the missing observations on the response or the predictors with artificial values that try to preserve the data set structure.
+This is a quite complex topic of its own, but we think why not.
+You can read more at from @imputation.
+
+## 2.3. Descriptive Analytics
+
+## 2.4. Data Source
+
+- [poverty.headcount](https://data.worldbank.org/indicator/SI.POV.DDAY)
+- [mpi](https://data.worldbank.org/indicator/SI.POV.MDIM.XQ)
+- [education.expenditure.primary](https://data.worldbank.org/indicator/SE.XPD.PRIM.ZS)
+- [education.expenditure.secondary](https://data.worldbank.org/indicator/SE.XPD.SECO.ZS)
+- [education.expenditure.tertiary](https://data.worldbank.org/indicator/SE.XPD.TERT.ZS)
+- [education.expenditure.total](https://data.worldbank.org/indicator/SE.XPD.TOTL.GD.ZS)
+- [health.expenditure](https://data.worldbank.org/indicator/SH.XPD.GHED.GD.ZS)
+- [military.expenditure](https://data.worldbank.org/indicator/MS.MIL.XPND.GD.ZS)
+- [fdi](https://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD)
+- [unemployment.rate](https://data.worldbank.org/indicator/SL.UEM.TOTL.NE.ZS)
+- [labour.force.participation](https://data.worldbank.org/indicator/SL.TLF.CACT.NE.ZS)
+- [gender.equality](https://data.worldbank.org/indicator/IQ.CPA.GNDR.XQ)
+- [population.growth](https://data.worldbank.org/indicator/SP.POP.GROW)
+- [urban.population.growth](https://data.worldbank.org/indicator/SP.URB.GROW)
+- [rural.population.growth](https://data.worldbank.org/indicator/SP.RUR.TOTL.ZG)
+- [gdp.deflator](https://data.worldbank.org/indicator/NY.GDP.DEFL.KD.ZG)
+- [gross capital formation](https://data.worldbank.org/indicator/NE.GDI.TOTL.ZS)
+- [trade](https://data.worldbank.org/indicator/NE.TRD.GNFS.ZS)
+- [region.class](https://datatopics.worldbank.org/world-development-indicators/the-world-by-income-and-region.html)
+- [income.class](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)
+- [gross.capital.formation](https://data.worldbank.org/indicator/NE.GDI.TOTL.ZS)

### 2022-10-30 src/DataCleaning.R

@@ -29,9 +29,9 @@ income.class <- importIncomeClass("income.class.csv")
 
 setwd("..")
 
-countries <- income.class %>% 
-  full_join(region.class, by = "country.name") %>% 
-  right_join(poverty.headcount, by = c("country.name", "country.code", "year")) %>%
+countries <- poverty.headcount %>% 
+  left_join(income.class, c("country.name", "country.code", "year")) %>% 
+  left_join(region.class, by = "country.name") %>%
   full_join(mpi, by = c("country.name", "country.code", "year")) %>%
   left_join(education.expenditure.total, by = c("country.name", "country.code", "year")) %>%
   left_join(education.expenditure.primary, by = c("country.name", "country.code", "year")) %>%
@@ -49,6 +49,12 @@ countries <- income.class %>%
   left_join(gender.equality, by = c("country.name", "country.code", "year")) %>%
   left_join(gross.capital.formation, by = c("country.name", "country.code", "year")) %>%
   left_join(trade, by = c("country.name", "country.code", "year"))
+
+d <- poverty.headcount %>% select(country.name, country.code) %>% mutate(isPov = T) %>%  
+  full_join(income.class %>% select(country.name, country.code) %>% mutate(isIncome = T), by = "country.name") %>%  
+  full_join(region.class %>% select(country.name) %>% mutate(isReg = T), by = "country.name") 
+
+d %>% filter(is.na(isIncome) | is.na(isIncome)) %>% distinct(country.name)
 
 cor_matrix <- cor(countries[, 6:23], use = "complete.obs")
 print(cor_matrix, 2)
@@ -76,4 +82,46 @@ sum(complete.cases(dataset))
 lrm <- lm(pov ~ ., dataset)
 summary(lrm)
 plot(lrm)
+
+
+poverty.headcount <- poverty.headcount %>%
+  mutate(country.name = correctName(country.name))
+mpi <- mpi %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.total <- education.expenditure.total %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.primary <- education.expenditure.primary %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.secondary <- education.expenditure.secondary %>%
+  mutate(country.name = correctName(country.name))
+education.expenditure.tertiary <- education.expenditure.tertiary %>%
+  mutate(country.name = correctName(country.name))
+health.expenditure <- health.expenditure %>%
+  mutate(country.name = correctName(country.name))
[…]

### 2022-10-30 src/draft.Rmd
Layouts

@@
 1.  **Use complete cases**: Only use the complete cases for the analysis. This is a straightforward approach, but doesn't resolve the bias resulted from the mass loss of data.
-2.  **Step-by-step remove frequently missing variables**: The same as we did before, but this process should be carried out carefully as we run the chance of dropping an important variable.
+2.  **Selectively remove variables with high missing rate**: The same as we did before, but this process should be carried out carefully as we run the chance of dropping an important variable.
 3.
@@
 ## 2.3. Descriptive Analytics
-
+(To be done)
 ## 2.4. Data Source
@@
 - [gross.capital.formation](https://data.worldbank.org/indicator/NE.GDI.TOTL.ZS)
+
+# 3. Model Selection and Interpretation
+## 3.1. Assumption Check (To be done)
+
+## 3.2. Ordinary Multiple Linear Regression
+
+We conduct a normal linear regression, following the approaches mentioned above to address missing values issues.
+
+### 3.2.1. Use Complete Cases (To be done)
+#### 3.2.1.1. Model Fitting
+#### 3.2.1.2. Assessment
+#### 3.2.1.3. Interpretation
+
+### 3.2.2. Selectively remove variables with high missing rate (To be done)
+#### 3.2.2.1. Model Fitting
+#### 3.2.2.2. Assessment
+#### 3.2.2.3. Interpretation
+
+### 3.2.3. Update the data set as we select variables (To be done)
+#### 3.2.3.1. Model Fitting
+#### 3.2.3.2. Assessment
+#### 3.2.3.3. Interpretation
+
+### 3.2.4. Imputation (To be done)
+#### 3.2.4.1. Model Fitting
+#### 3.2.4.2. Assessment
+#### 3.2.4.3. Interpretation
+
+## 3.3. Panel Data Analysis (To be done)
+
+# 4. Conclusion
+# 5. Appendix
+# 6. References

### 2022-10-30 src/draft.Rmd
changes in draft structure
export countries data

@@
 date: "`r Sys.Date()`"
-output: pdf_document
+output:
+  pdf_document: default
+  html_document:
+    df_print: paged
 bibliography: citations.bib
@@
 ```
+
+# 1. Introduction (To be done)
 
@@
 
-These can be very useful and relevant information [@pubspend&pov].
+expenditure in primary, secondary, and tertiary edication can be very useful and relevant information to predict poverty reduction [@pubspend&pov].
 However, we would like to exclude these variables from some first analyses to make use of the richer set of data.
@@
 
-## 2.3. Descriptive Analytics
-(To be done)
+## 2.3. Descriptive Analytics (To be done)
+
 ## 2.4. Data Source

### 2022-11-03 src/All variables.Rmd
I created the linear regression model with all the variables

---
title: "Imtiyaaz"
output: html_document
date: "2022-11-03"
---
```{r}
library(tidyverse) # for data manipulations and visualisations
library(ggpubr) # for correlation coefficient on plot
library(gridExtra) # for arranging plots 
library(e1071) # for skewness
```

```{r}
df <-read.csv('data/countries.csv' , ,sep = "", header = TRUE)
df
subset1 <- df %>% select("income","edu.total","edu.pri","edu.sec","edu.ter","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","pop.gwth.urban","gdp.dflt","gdr.eql","gcf","trade")

```

```{r}
set.seed(10)
index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
train <- df[index, ]
test <- df[-index, ]
```

```{r}
lm.slr <- lm(pov ~ income+edu.total+edu.pri+edu.sec+edu.ter+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gdr.eql+gcf+trade,data = train)
summary(lm.slr)

```

### 2022-11-03 src/draft.Rmd
add gdp.pc (gdp per capita) to countries.csv and draft docs

@@
 23. **trade**: Trade = import + export (% of GDP)
+24. **gdp.pc**: GDP per capita (current US$)
 
@@
 income.class <- importIncomeClass("income.class.csv")
+gdp.pc <- importWDI("gdp.pc.csv", "gdp.pc")
 
@@
   mutate(country.name = correctName(country.name))
+gdp.pc <- gdp.pc %>%
+  mutate(country.name = correctName(country.name))
 ```
@@
   left_join(trade, by = c("country.name", "country.code", "year")) %>% 
+  left_join(gdp.pc, by = c("country.name", "country.code", "year")) %>% 
   # filter special groups
@@
 
-## 2.3. Descriptive Analytics (To be done)
+## 2.3. Descriptive Analytics
+Distribution of the predicted variable `pov`
+```{r}
+ggplot(countries, aes(x = pov)) +
+  geom_bar()
+```
 

### 2022-11-06 src/draft.Rmd
time series graph + intergpret

@@
 5.  **reg**: Region\
-6.  **pov**: Poverty headcount ratio\
+6.  **pov**: Poverty headcount ratio based on cut-off value of $2.15 per day\
 7.  **mpi**: Multidimensional Poverty Index\
@@
 ggplot(countries, aes(x = pov)) +
-  geom_bar()
+  geom_histogram(aes(y = ..density..)) +
+  geom_density(alpha = .2, fill = "blue")
 ```
+The graph displays a decreasing rate as poverty indicator increasing.
+This might not be representative of the current state of poverty in the world, but of the number presented in our data.
+For example, more recent data is likely to be more inclusive than ancient data, when poverty is more prevalent.
+We should look at data from the same period.
+
+```{r}
+# pov data from 1998 and 2018
+pov.98.18 <- poverty.headcount %>%
+  filter(year == 1998 | year == 2018)
+pov.98.18 %>% 
+  group_by(year) %>% 
+  summarise(sum = n())
+
+ggplot(pov.98.18, aes(x = pov)) +
+  geom_histogram(aes(y = ..density..)) +
+  geom_density(alpha = .2, fill = "blue") +
+  facet_grid(cols = vars(year))
+```
+
+The graph for 1998 has a much gentler slope, meaning poverty was more popular during that time, as predicted from out intuition. What about the general progress of the world?
+
+```{r}
+# Re-import pov and only take special regions
+# geographic
+geo.regs <- c("EAS", "ECS", "LCN", "MEA", "SAS", "SSF", "WLD")
+# economics
+eco.regs <- c("HIC", "LIC", "LMC", "LMY", "UMC")
+
+pov.reg <- importWDI("../data/poverty.headcount.215dollar.csv", "pov") %>% 
+  filter(country.code %in% c(geo.regs, eco.regs)) %>% 
+  mutate(type = ifelse(country.code %in% geo.regs, "Geographic", "Economics"))
+
+pov.reg %>%
+  distinct(country.code, country.name, type) %>% 
+  arrange(type)
+
+# World
+ggplot(pov.reg %>% filter(country.code == "WLD"),
+       aes(x = year, y = pov)) +
+  geom_point() +
+  geom_smooth()
+```
+
+An overall very steady decrease of poverty. How about each region?
+
+```{r}
+ggplot(pov.reg,
+       aes(x = year, y = pov, color = country.name)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(rows = vars(type))
+```
+
+There's a general steady, but distinct decline of poverty over time in each type of region.
 

### 2022-11-06 src/draft.Rmd
descriptive analysis

@@
 library(ggplot2)
+library(e1071)
+library(moments)
 ```
@@
 ```{r}
-ggplot(pov.reg,
+ggplot(pov.reg %>% filter(country.code != "WLD"),
        aes(x = year, y = pov, color = country.name)) +
@@
 
-There's a general steady, but distinct decline of poverty over time in each type of region.
+There's a general steady, but distinct decline of poverty over time in each type region of respective type.
+Latin America & Caribbean, Europe & Central Asia, Middle East & North Africa, and High Income group has a more gradual decline as they are not very poor to begin with.
+
+Among the income groups, Low & Middle Income, Lower & Middle Income, and Upper Middle Income have quite similar in term of poverty indicator and slope over the year.
+While these values vary greatly among different geographical regions.
+
+Let's see some important statistics
+
+```{r}
+stats <- poverty.headcount %>% 
+  summarise(count = n(), skewness = skewness(pov), kurtosis = kurtosis(pov), std.deviation = sd(pov))
+
+kable(stats)
+```
 

### 2022-11-07 src/draft.Rmd
preliminary assumption check

@@
 bibliography: citations.bib
+urlcolor: blue
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = F)
 knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
@@
 library(moments)
+library("PerformanceAnalytics")
 ```
@@
 ```{r}
+# number of data point available from 1967 to 2021
+ggplot(poverty.headcount, aes(x = year)) +
+  geom_histogram()
+
 # pov data from 1998 and 2018
@@
 # 3. Model Selection and Interpretation
-## 3.1. Assumption Check (To be done)
+## 3.1. Assumption Check
+We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data.
+[Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
+**1. Linear relationship:**  
+Use correlation matrix to check linearity
+```{r, warning = FALSE}
+# select only numeric data
+countries.num <- countries %>% select(where(is.numeric))
+chart.Correlation(countries.num, histogram = TRUE, pch = 19)
+```
+Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
+```{r}
+# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
+countries.num2 <- countries.num %>% 
+  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
+  filter(!is.na(value) & !is.na(pov))
+
+drawGraph <- function(indvar, data) {
+  ggplot(data %>% 
+           filter(variable == indvar),
+         aes(x = value, y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste("Relationship pov ~", indvar),
+         x = indvar,
+         y = "pov")
+}
+
+ivs <- distinct(countries.num2, variable)[[1]]
+
+for (indvar in ivs) {
+  print(
+    ggplot(countries.num2 %>% 
+           filter(variable == indvar),
+         aes(x = value, y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste("Relationship pov ~", indvar),
+         x = indvar,
+         y = "pov")
+  )
+}
+```
+
+Most variables display linear relationship with some exceptions, which are more appropriate to assume a logarithmic relationship. Gender equality should be viewed as a categorical data.
+```{r}
+logIvs <- c("fdi", "gdp.dflt", "gdp.pc")
+
+for (indvar in logIvs) {
+  print(
+    ggplot(countries.num2 %>% 
+           filter(variable == indvar),
+         aes(x = log(value), y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste0("Relationship pov ~ log(", indvar, ")"),
+         x = paste0("log(", indvar, ")"),
+         y = "pov")
+  )
+}
+```
+
+The relationship are more linear after the transformation.
+
+```{r}
+gdr.eql <- countries.num2 %>% 
+  filter(variable == "gdr.eql") %>% 
+  mutate(value = factor(value))
+
+ggplot(gdr.eql,
+       aes(x = value, y = pov)) +
+  geom_boxplot() +
+  geom_smooth(method = lm) +
+  labs(title = paste("Relationship pov ~ gdr.eql"),
+       x = "gdr.eql",
+       y = "pov")
+```
+There's a significant correlation between gender equality and poverty.  
+Correlation coefficients between `pov` and predictors.
+
+```{r}
+# transform some variables
+countries.num3 <- countries.num %>% 
+  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
+  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
+
+# name of the independent variables
+cn <- colnames(countries.num3)[-c(1,2)]
+
+corWithPov <- function(indevar, data) {
+  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
+}
+
+cor.pov <- sapply(cn, corWithPov, data = countries.num3)
+kable(cor.pov)
+```
+
+`edu.ter`, `mil`, `fdi`, `lbr.part`, `unemp`, `gdp.dflt`, `gcf` have negligible correlation with `pov`.
+`lgdp.pc`, `lfdi`, and `lgdp.dflt` have stronger linear relationship with `pov` than their un-transformed counterparts.
+
+**2. Predictors are independent**
+There should be no correlation/multicolinearity between each pair of predictors.
+
+```{r}
+countries.arr <- simplify2array(countries.num %>% 
+                                  select(-c("year", "pov")))
+
+cor.mtx <- rcorr(countries.arr, type = "spearman")
+
+round(cor.mtx$r, 2)
+
+corrplot(cor.mtx$r, type = "upper", order = "hclust", 
+         tl.col = "black", tl.srt = 45)
+```
+Some significant correlations can be found between `gdr.eql` with `edu.pri`, `pop.gwth.total`, and `fdi` with `mil`, etc.  
+We expect these variables to be eliminated in the VIF test.  
+Log transform some variables.  
+```{r}
+countries1 <- countries1 %>% 
+  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
+  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi)) # for lack of better measure, Inf and NaN value of lfdi will be ignored
+```
 

### 2022-11-07 src/Selectively_remove_variables.Rmd
Create Selectively_remove_variables.Rmd
Updated my part on "Selectively removin g variables", tried to do ridge regression but realise that heteroskedascity still exists, will try to consult prof tmr

---
title: "Selectively Remove Variables"
author: "Callista Stephine Yu"
date: "2022-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
library(knitr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stats)
library(tidyverse)
library(e1071)
library(car)
```

```{r}
countries <- read.csv("../data/countries.csv")
```

```{r}
head(countries)
```

```{r}

#dataset without the red 
#removing reg as it is a bit too complex to deal for now, not relevant

#subset1 <- subset(countries, select = c("edu.total", "fdi", "gcf", "gdp.dflt", "hlth", "income", "lbr.part", "mil", "pop.gwth rural", "pop.gwth.urban", "pov", "reg", "trade", "unemp", "year"))
subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))

subset1 <- subset0 %>% mutate(lfdi = log(subset1$fdi)) %>% mutate(lgdp.pc = log(subset1$gdp.pc)) %>% mutate(lgdp.gdlt = log(subset1$gdp.dflt))
#dataset
```

```{r}
#converting non-numeric arguments into factors

subset1$income <- factor(subset1$income)
subset1$income <- as.numeric(subset1$income)

```

```{r}
#Perform standardisation

(df.countries0 <- subset1)
scaler <- apply(X = df.countries0, MARGIN = 2, FUN = sd)

(df.countries <- as.data.frame(apply(df.countries0, 2, function(x) x/sd(x, na.rm = TRUE))))

```

```{r}
#Divide each feature/target by its standard deviation

head(df.countries)
```

```{r}
#check for correlation

corrplot(cor(df.countries, use = "complete.obs"), 
         method = "number",
         type = "upper",
         tl.col = "black", 
         tl.srt = 30
         )
```

```{r}
set.seed(123)
index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
train_1 <- subset1[index,]
test_1 <- subset1[-index,]

```

```{r}
#Boxplot of the predictor variables
```

```{r}
```

```{r}
#Everything
lm_1 <- lm(data= train_1, formula = pov ~ .)
summary(lm_1)
```
```{r}
#remove hlth 
lm_2 <- lm(data= train_1, formula = pov ~ edu.total + gcf + fdi + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_2)
plot(lm_2, which = 5)
```

```{r}
#Remove fdi
lm_3 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_3)
plot(lm_3, which = 5)
```
```{r}
#remove year 
lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_4)
plot(lm_4, which = 5)
```
```{r}
#remove gdp.dflt 
lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_4)
plot(lm_4, which = 5)
```
```{r}
lm_5 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc)
summary(lm_5)
plot(lm_5, which = 5)
```
```{r}
#removing lfdi
lm_6 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc  + lgdp.pc)
summary(lm_6)
plot(lm_6, which = 4)
```
```{r}
#
data0 <- subset(df.countries, select = c("edu.total","gcf","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp","gdp.pc", 
                                        "lgdp.pc"))
data <- na.omit(data0)
train.x <- data.matrix(subset(data, select = -pov))
train.y = data.matrix(data$pov)

corrplot(cor(data, use = "complete.obs"), 
         method = "number",
         type = "upper",
         tl.col = "black", 
         tl.srt = 30
         )
```

```{r}
#Find the ridge regression model
set.seed(123)
rate <- 0.8
train.size <- round(nrow(data)*rate)
sample(sample(nrow(data), train.size))
training <- data[sample, ]
test <- data[-sample, ]

```

```{r}
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)

t(coef(glm_ridge))
```

```{r}
train.x <- model.matrix(pov ~., data = training)[, -1]
train.y <- training$pov
test.x <- model.matrix(pov ~., data = test)[,-1]
test.y <- test$pov
```

```{r}
model_ridge_trial1 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.1)
t(coef(model_ridge_trial1))
```

```{r}
model_ridge_trial2 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.5)
t(coef(model_ridge_trial2))
```

```{r}
model_ridge_trial3 <- glmnet(train.x, train.y, alpha = 0, lambda = 1)
t(coef(model_ridge_trial3))
```
```{r}
table_summary <- cbind(coef(model_ridge_trial1),
                       coef(model_ridge_trial2),
                       coef(model_ridge_trial3))

colnames(table_summary) <- c("lambda = 0.1",
                             "lambda = 0.5",
                             "lambda = 1")

table_summary
```

```{r}
#generate sequence of lambda values 
lambda <- 10^seq(-6, 2, length = 100)
ridge_model = glmnet(train.x, train.y, alpha = 0, lambda = lambda)

t(coef(ridge_model))
```

```{r}
plot(ridge_model, xvar = "lambda", label = TRUE)

#customise the plot with labels
add_lbs <- function(fit, offset_x = 2.5) {
  L <- length(fit$lambda[L]) + offset_x
  y <- fit$beta[,L]
  labs <- names(y)
  text(x, y , labels = labs, cex = 1.5)

  title(main = "Coefficients of Predictors vs Log(Lambda)",cex.main = 1.4, adj = 0, line = 3)
  title(sub = "Number of Predictors", cex.sub = 1.1, adj = 0.55, line = -21.5)
  add_lbs(ridge_model)
  legend("topright", lwd = 1, col = 1:6, legend = colnames(train.x), cex = .7)

}
```

```{r}
plot_glmnet(ridge_model)
```
```{r}
#cross - validation
set.seed(123)
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0, type.measure = "mse")

```

```{r}
cv_ridge
cv_ridge$lambda.min
cv_ridge$lambda.1se
```

```{r}
plot(cv_ridge)
abline(v=log(cv_ridge$lambda.min), col = "red", lty=5)
text(log(cv_ridge$lambda.min)+1, 1.4, 
     labels= paste0("lambda_min = ", 
                    round(cv_ridge$lambda.min, digits = 3)), 
     cex = 1.1)

# Label for lambda.1se
abline(v=log(cv_ridge$lambda.1se), 
       col = "red", lty=5)
text(log(cv_ridge$lambda.1se)+1, 1.25, 
     labels= paste0("lambda_1se = ", 
                    round(cv_ridge$lambda.1se, digits = 3)), cex = 1.1)

```

```{r}
#train the ridge regression model using lambda.min

glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
t(coef(glm_Ridge))
```
```{r}
eval_results <- function(fit, true) {
  actual <- data.matrix(true)
  SSE <- sum((actual - fit)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE/SST
  data.frame(
    MSE = MSE(fit, true),
    MAE = MAE(fit, true),
    RMSE = RMSE(fit, true),
    MAPE = MAPE(fit, true),
    R2  = R_square

  )
}
```

```{r}
fit <- predict(glm_Ridge, train.x)
true <- train.y
summary_Ridge_train <- eval_results(fit, true)
summary_Ridge_train
```

```{r}
fit <- predict(glm_Ridge, test.x)
true <- test.y
summary_Ridge_test <- eval_results(fit, true)
summary_Ridge_test
```

```{r}
summary_Ridge <- rbind(summary_Ridge_train[-5], summary_Ridge_test[-5])
rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
knitr::kable(summary_Ridge, digits = 3)

```

```{r}
#Obtain fitted and true values

fit <- predict(glm_Ridge, train.x)
true <- train.y

```

```{r}
#Plot residuals s
plot(fit, true - fit, 
     main = "Residual Plot of Ridge Regression Model on the Training dataset",
     xlab = "Fitted value",
     ylab = "Standardised Residuals")
abline(h = 0, col = "red", lwd = 2)
abline(h = 1, col = "red", lty = 2, lwd = 2)
abline(h = -1, col = "red", lty = 2, lwd = 2)

```
Constant varince assumption not satisfied. What to do? Most likely need to transform the variable

```{r}
mlr_adj <- lm(pov ~., data = training)

summary(mlr_adj)
```

### 2022-11-08 src/draft.Rmd
some imputation

@@
 knitr::opts_chunk$set(echo = TRUE, warning = F)
-knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
+knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
 ```
@@
 ```{r, message = F}
-# required library
-library(knitr)
-library(readr)
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(e1071)
-library(moments)
-library("PerformanceAnalytics")
+# Package names
+# packages <- c("MASS", "knitr", "readr", "tidyr", "dplyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "needs")
+# 
+# # Install packages not yet installed
+# installed_packages <- packages %in% rownames(installed.packages())
+# if (any(installed_packages == FALSE)) {
+#   install.packages(packages[!installed_packages])
+# }
+# 
+# # Packages loading
+# invisible(lapply(packages, library, character.only = TRUE))
+# 
+# prioritize(dplyr)
+
+library(needs)
+
+needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice")
+
+prioritize(dplyr)
 ```
@@
 # exclude these variables in countries1
-countries1 <- countries %>% select(!hMiss)
+countries1 <- countries %>% 
+  select(!hMiss) %>% 
+  filter(!is.na(pov))
 str(countries1)
@@
 This is a quite complex topic of its own, but we think why not.
-You can read more at from @imputation.
+You can read more from @imputation.
 
@@
 We expect these variables to be eliminated in the VIF test.
-Log transform some variables.  
-```{r}
-countries1 <- countries1 %>% 
-  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
-  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi)) # for lack of better measure, Inf and NaN value of lfdi will be ignored
-```
+This might be a good property for imputation (3.2.4)
 
@@
 ### 3.2.4. Imputation (To be done)
+What is Imputation?
+**1. What is imputation**  
+Imputation is a technique to handle missing values by replacing missing data with substitute values.
+In our study, we focus on "item imputation", which means substituting for a component of a data point (i.e.
+a variable).
+The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), or predict the value using other variables.
+
+We use **Multivariate Imputation By Chained Equations (MICE)** as our primary tool for this technique, because of its wide acceptance in scientific studies (@imputWorks).
+In order to use this technique, we have to assume that the "missingness" of a field can be explained by the values in other columns, (e.g., If the countries is in North America, it's more likely to be missing as we have discovered in 2.2).
+The general ideas of the algorithm is iteratively predict the missing values based on other variables, and improve it after each iterations until predicted value converse to a stable point.
+The algorithmic details of MICE is very concisely explained in @micealgo.
+
+Our parameter for this algorithm is fairly standard:
+
+- `m = 5` imputed data sets
+- `maxit = 30` iterations
+
+The imputation method is `cart` (Classification and Regression Trees).
+
+```{r}
+set.seed(1984)
+# split data
+idx <- sample(1:nrow(countries1), replace = F, 0.7 * nrow(countries1))
+countries.train.4 <- countries1[idx,]
+countries.test.4 <- countries[-idx,]
+
+# blank for no imputation, cart for variable requiring imputation
+meth <- c(rep("", 4), "cart", "", rep("cart", 16))
+names(meth) <- colnames(countries1)
+meth
+
+# run the algorithm
+# countries1.imputed <- mice(countries1, m = 5, maxit = 30, method = meth)
+countries1.imputed <- readRDS("../countries1.imputed.RData")
+summary(countries1.imputed)
+```
+We can take a look into one of the imputated data sets.
+```{r}
+countries1.imputed.1 <- complete(countries1.imputed, 1)
+summary(countries1.imputed.1)
+sum(!complete.cases(countries1.imputed.1))
+```
+We found no missing cases as expected.
+
 #### 3.2.4.1. Model Fitting
+We generated 5 sets of imputated (training) data.
+We can build our model on each of these data sets and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
+We remove `reg` and `country.name` from the model because they're perfectly predicted by `country.code`
+
+We can start with using one of the imputated sets.
+```{r}
+fit1 <- lm(pov ~ .-country.name, countries1.imputed.1)
+summary(fit1)
+```
+
+We might be more interested with large regions of countries rather than individual countries.
+```{r}
+fit1reg <- lm(pov ~ .-country.name-country.code, countries1.imputed.1)
+summary(fit1reg)
+```
+
+With all the generated data sets.
+
+```{r}
+fit2 <- with(data = countries1.imputed, exp = lm(pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc))
+fit2.combined <- pool(fit2)
+summary(fit2.combined)
+pool.r.squared(fit2.combined)
+```
+
+We can conduct a step-wise AIC variable selection.
+It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models.
+
+continue with fitting and displaying model [link](https://thomaselove.github.io/432-notes/multiple-imputation-and-linear-regression.html#fit-linear-regression-using-ols-and-fit.mult.impute)  
+-> AIC on one model [link](https://rforpoliticalscience.com/2020/10/23/choose-model-variables-by-aic-in-a-stepwise-algorithm-with-the-mass-package-in-r/)  
+-> stepwise selection on mice  [link](https://stefvanbuuren.name/fimd/sec-stepwise.html)
+
 #### 3.2.4.2. Assessment
+Check assumption
+outliers
 #### 3.2.4.3. Interpretation

### 2022-11-14 src/draft.Rmd

@@
 
-needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice")
+needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet")
 
@@
 
+# correlation of independent variable and pov
 corWithPov <- function(indevar, data) {
@@
 `lgdp.pc`, `lfdi`, and `lgdp.dflt` have stronger linear relationship with `pov` than their un-transformed counterparts.
+
+```{r}
+countries1 <- countries1 %>% 
+  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
+  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
+```
 
@@
 
-### 3.2.4. Imputation (To be done)
-What is Imputation?
+### 3.2.4. Imputation
 **1. What is imputation**  
@@
 # split data
-idx <- sample(1:nrow(countries1), replace = F, 0.7 * nrow(countries1))
-countries.train.4 <- countries1[idx,]
-countries.test.4 <- countries[-idx,]
+isComplete <- which(complete.cases(countries1))
+idx <- sample(isComplete, replace = F, 0.3 * nrow(countries1))
+countries.train.4 <- countries1[-idx,]
+countries.test.4 <- countries1[idx,]
 
@@
 # run the algorithm
-# countries1.imputed <- mice(countries1, m = 5, maxit = 30, method = meth)
-countries1.imputed <- readRDS("../countries1.imputed.RData")
+# countries1.imputed <- mice(countries1, m = 3, maxit = 20, method = meth)
+# saveRDS(countries1.imputed, 'countries1.imputed.RData')
+countries1.imputed <- readRDS('countries1.imputed.RData')
 summary(countries1.imputed)
@@
 #### 3.2.4.1. Model Fitting
+##### A. Ordinary Linear Regression
 We generated 5 sets of imputated (training) data.
 We can build our model on each of these data sets and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
-We remove `reg` and `country.name` from the model because they're perfectly predicted by `country.code`
-
-We can start with using one of the imputated sets.
-```{r}
-fit1 <- lm(pov ~ .-country.name, countries1.imputed.1)
-summary(fit1)
-```
-
-We might be more interested with large regions of countries rather than individual countries.
-```{r}
-fit1reg <- lm(pov ~ .-country.name-country.code, countries1.imputed.1)
-summary(fit1reg)
-```
-
 With all the generated data sets.
@@
 ```{r}
-fit2 <- with(data = countries1.imputed, exp = lm(pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc))
+formula.str <- 'pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'
+
+# Using only the first imputed data set
+fit1 <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
+summary(fit1)
+fit1.pred <- predict(fit1, countries.test.4)
+
+# Build models with all generated data set and pool the estimates
+fit2 <- with(data = countries1.imputed, exp = lm(as.formula(formula.str)))
 fit2.combined <- pool(fit2)
 summary(fit2.combined)
-pool.r.squared(fit2.combined)
+# dummy lm model
+fit2.dummy <- lm(as.formula(formula.str), data = countries1)
+# replace coefficients of dummy model & predict
+fit2.dummy$coefficients <- fit2.combined$pooled$estimate
+fit2.pred <- predict(fit2.dummy, countries.test.4)
+```
+Helper function to calculate R-Squared
+```{r}
+r2 <- function(pred, orig) {
+   RSS <- sum((pred - orig)^2)
+   TSS <- sum((pred - mean(orig))^2)
+   R2 <- 1 - RSS/TSS
+   return (R2)
+}
+
+adjR2 <- function(pred, orig, k) {
+  R2 <- r2(pred, orig)
+  n <- length(pred)
+  adjr2 <- 1 - (1-R2)*(n-1)/(n-k-1)
+  return (adjr2)
+}
+```
+Adjusted R-Squared on train and test
+```{r}
+k <- length(fit1$coefficients) - 1
+fit1.train.r2 <- summary(fit1)$adj.r.squared
+fit1.test.r2 <- adjR2(fit1.pred, countries.test.4$pov, k)
+fit2.train.r2 <- pool.r.squared(fit2, adjusted = T)[,'est']
+fit2.test.r2 <- adjR2(fit2.pred, countries.test.4$pov, k)
+res <- data.frame(model = c("Single", "Pooled"), train = c(fit1.train.r2, fit2.train.r2), test = c(fit1.test.r2, fit2.test.r2))
+kable(res)
 ```
 
+The models seem to be over-fitting. We should perform model simplification.
+
+##### B. Regularization
+
+Prepare data
+
+```{r}
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
+
+  return(sd(vct, na.rm = T))
+}
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+
+  return(vct/scl)
+}
+
+standardise <- function(train, test) {
+  data <- rbind(train, test)
+  scaler <- unlist(lapply(data, sd0))
+
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}
+```
+
+```{r}
+optimalModel <- function(train) {
+  #standardized
+}
+```
+
+##### B. Step-wise AIC
 We can conduct a step-wise AIC variable selection.
 It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models.
+The lower AIC is, the better the model fits.
+\\
+We have no idea how to pool coefficients for AIC model,
 
-continue with fitting and displaying model [link](https://thomaselove.github.io/432-notes/multiple-imputation-and-linear-regression.html#fit-linear-regression-using-ols-and-fit.mult.impute)  
+```{r}
+# helper
+fitAIC <- function(i) {
+  fit <- lm(as.formula(formula.str), complete(countries1.imputed, i))
+  aic.fit <- stepAIC(fit, trace = F, direction = 'backward')
+  return(aic.fit)
+}
+
+# build models
+aic.fits <- lapply(1:countries1.imputed$m, fitAIC)
+# predict with each model
+aic.pred <- lapply(aic.fits, predict, newdata = countries.test.4)
+# calculate adj r2
+aic.train.r2 <- unlist(lapply(aic.fits, \(model) summary(model)$adj.r.squared))
+
+k <- lapply(aic.fits, \(m) length(m$coefficients))
+aic.test.r2 <- unlist(mapply(adjR2, aic.pred, k = k, MoreArgs = list(orig = countries.test.4$pov)))
+
+res <- data.frame(`set no.` = 1:3, train = aic.train.r2, test = aic.test.r2)
+```
+
+Adjusted R-Squared
+
+```{r}
+kable(res)
+```
+
 -> AIC on one model [link](https://rforpoliticalscience.com/2020/10/23/choose-model-variables-by-aic-in-a-stepwise-algorithm-with-the-mass-package-in-r/)
+
 -> stepwise selection on mice  [link](https://stefvanbuuren.name/fimd/sec-stepwise.html)

### 2022-11-14 src/helperFunctions.R
helper functions, standardise

@@ -1,0 +1,29 @@
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
+  
+  return(sd(vct, na.rm = T))
+}
+
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+  
+  return(vct/scl)
+}
+
+scale <- function(data) {
+  unlist(lapply(data, sd0))
+}
+
+standardise <- function(data) {
+  scaler <- scale(data)
+  
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}

### 2022-11-14 data/forward.Rmd

---
title: "Untitled"
output: html_document
date: "2022-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

```{r cars}
summary(lm(formula = pov ~ hlth,data = countries1))
summary(lm(formula = pov ~ fdi,data = countries1))
summary(lm(formula = pov ~ reg,data = countries1))
summary(lm(formula = pov ~ mil,data = countries1))
summary(lm(formula = pov ~ lbr.part,data = countries1))
summary(lm(formula = pov ~ unemp,data = countries1))
summary(lm(formula = pov ~ trade,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural,data = countries1))
summary(lm(formula = pov ~ pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ gcf,data = countries1))
summary(lm(formula = pov ~ gdp.dflt,data = countries1))
summary(lm(formula = pov ~ trade,data = countries1))
```
```{r}
summary(lm(formula = pov ~ hlth+reg,data = countries1))
summary(lm(formula = pov ~ fdi+reg,data = countries1))
summary(lm(formula = pov ~ mil+reg,data = countries1))
summary(lm(formula = pov ~ lbr.part+reg,data = countries1))
summary(lm(formula = pov ~ unemp+reg,data = countries1))
summary(lm(formula = pov ~ trade+reg,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total+reg,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural+reg,data = countries1))
summary(lm(formula = pov ~ pop.gwth.urban+reg,data = countries1))
summary(lm(formula = pov ~ gcf+reg,data = countries1))
summary(lm(formula = pov ~ gdp.dflt+reg,data = countries1))
summary(lm(formula = pov ~ trade+reg,data = countries1))
summary(lm(formula = pov ~ income+reg,data = countries1))

```
```{r}
summary(lm(formula = pov ~ income+reg,data = countries1))
summary(lm(formula = pov ~ hlth+reg+income,data = countries1))
summary(lm(formula = pov ~ fdi+reg+income,data = countries1))
summary(lm(formula = pov ~ mil+reg+income,data = countries1))
summary(lm(formula = pov ~ lbr.part+reg+income,data = countries1))
summary(lm(formula = pov ~ unemp+reg+income,data = countries1))
summary(lm(formula = pov ~ trade+reg+income,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total+reg+income,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural+reg+income,data = countries1))
summary(lm(formula = pov ~ pop.gwth.urban+reg+income,data = countries1))
summary(lm(formula = pov ~ gcf+reg+income,data = countries1))
summary(lm(formula = pov ~ gdp.dflt+reg+income,data = countries1))
summary(lm(formula = pov ~ trade+reg+income,data = countries1))
```
```{r}
summary(lm(formula = pov ~ mil + reg + income, data = countries1))

```
```{r}
summary(lm(formula = pov ~ pop.gwth.urban + reg + income + mil, data = countries1))
summary(lm(formula = pov ~ hlth+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ fdi+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ lbr.part+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ unemp+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ pop.gwth.total+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ pop.gwth.rural+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ gcf+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ gdp.dflt+reg+income+mil+pop.gwth.urban,data = countries1))
summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
```
```{r}
countries9 <- countries1 %>% 
  select(Age, TotalWorkingYears, NumCourses, YearsAtCurrCompany, MonthlyIncome)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-11-14 src/Selectively_remove_variables.Rmd

@@
 library(car)
+library(glmnet)
+library(corrplot)
 ```
-
+We first read the 
 ```{r}
@@
 
-#subset1 <- subset(countries, select = c("edu.total", "fdi", "gcf", "gdp.dflt", "hlth", "income", "lbr.part", "mil", "pop.gwth rural", "pop.gwth.urban", "pov", "reg", "trade", "unemp", "year"))
+#subset1 <- subset(countries, select = c("edu.total", "fdi", "gcf", "gdp.dflt", "hlth", "income", "lbr.part", "mil", "pop.gwth.rural", "pop.gwth.urban", "pov", "reg", "trade", "unemp", "year"))
 subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
 
-subset1 <- subset0 %>% mutate(lfdi = log(subset1$fdi)) %>% mutate(lgdp.pc = log(subset1$gdp.pc)) %>% mutate(lgdp.gdlt = log(subset1$gdp.dflt))
+subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.gdlt = log(subset0$gdp.dflt))
 #dataset
@@
 Constant varince assumption not satisfied. What to do? Most likely need to transform the variable
+TO TRANSFORM TO BOXCOX
 
@@
 ```
+
+```{r}
+#Train a LASSO regression model
+set.seed(123)
+cv_lasso <- cv.glmnet(train.x, train.y, alpha = 1, type.measure = "mse")
+```
+
+```{r}
+cv_lasso
+```
+
+```{r}
+plot(cv_lasso)
+```
+
+```{r}
+glm_Lasso <- glmnet(train.x, train.y, alpha = 1, lambda = cv_lasso$lambda.min)
+t(coef(glm_Lasso))
+```
+
+```{r}
+fit2 <- predict(glm_Lasso, train.x)
+true2 <- train.y
+summary_Lasso_train <- eval_results(fit2, true2)
+summary_Lasso_train
+```
+
+```{r}
+fit3 <- predict(glm_Lasso, test.x)
+true3 <- test.y
+summary_Lasso_test <- eval_results(fit3, true3)
+summary_Lasso_test
+```

### 2022-11-14 data/forward.Rmd
forward2

@@
 ```{r}
+df_count<-read.csv("../data/countries1.csv")
+head(df_count)
+
+set.seed(10)
+index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
+train <- df_count[index, ]
+test <- df_count[-index, ]
+str(test)
+```
+
+income has highest adjusted r-square value =0.6105
+```{r cars}
+summary(lm(formula = pov ~ income,data = train))
+summary(lm(formula = pov ~ reg,data = train))
+summary(lm(formula = pov ~ edu.total,data = train))
+summary(lm(formula = pov ~ hlth,data = train))
+summary(lm(formula = pov ~ mil,data = train))
+summary(lm(formula = pov ~ fdi,data = train))
+summary(lm(formula = pov ~ lbr.part,data = train))
+summary(lm(formula = pov ~ unemp,data = train))
+summary(lm(formula = pov ~ trade,data = train))
+summary(lm(formula = pov ~ pop.gwth.total,data = train))
+summary(lm(formula = pov ~ pop.gwth.rural,data = train))
+summary(lm(formula = pov ~ pop.gwth.urban,data = train))
+summary(lm(formula = pov ~ gdp.dflt,data = train))
+summary(lm(formula = pov ~ gcf,data = train))
+summary(lm(formula = pov ~ trade,data = train))
+summary(lm(formula = pov ~ gdp.pc,data = train))
+summary(lm(formula = pov ~ lfdi,data = train))
+summary(lm(formula = pov ~ lgdp.dflt,data = train))
+summary(lm(formula = pov ~ lgdp.pc,data = train))
 
@@
 
-```{r cars}
-summary(lm(formula = pov ~ hlth,data = countries1))
-summary(lm(formula = pov ~ fdi,data = countries1))
-summary(lm(formula = pov ~ reg,data = countries1))
-summary(lm(formula = pov ~ mil,data = countries1))
-summary(lm(formula = pov ~ lbr.part,data = countries1))
-summary(lm(formula = pov ~ unemp,data = countries1))
-summary(lm(formula = pov ~ trade,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ gcf,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt,data = countries1))
-summary(lm(formula = pov ~ trade,data = countries1))
+#reg+income adjusted r-square=0.7031
+```{r}
+summary(lm(formula = pov ~ reg+income,data = train))
+summary(lm(formula = pov ~ edu.total+income,data = train))
+summary(lm(formula = pov ~ hlth+income,data = train))
+summary(lm(formula = pov ~ mil+income,data = train))
+summary(lm(formula = pov ~ fdi+income,data = train))
+summary(lm(formula = pov ~ lbr.part+income,data = train))
+summary(lm(formula = pov ~ unemp+income,data = train))
+summary(lm(formula = pov ~ trade+income,data = train))
+summary(lm(formula = pov ~ pop.gwth.total+income,data = train))
+summary(lm(formula = pov ~ pop.gwth.rural+income,data = train))
+summary(lm(formula = pov ~ pop.gwth.urban+income,data = train))
+summary(lm(formula = pov ~ gdp.dflt+income,data = train))
+summary(lm(formula = pov ~ gcf+income,data = train))
+summary(lm(formula = pov ~ trade+income,data = train))
+summary(lm(formula = pov ~ gdp.pc+income,data = train))
+summary(lm(formula = pov ~ lfdi+income,data = train))
+summary(lm(formula = pov ~ lgdp.dflt+income,data = train))
+summary(lm(formula = pov ~ lgdp.pc+income,data = train))
+
 ```
+#lgdp.pc+income+reg r=0.7364
 ```{r}
-summary(lm(formula = pov ~ hlth+reg,data = countries1))
-summary(lm(formula = pov ~ fdi+reg,data = countries1))
-summary(lm(formula = pov ~ mil+reg,data = countries1))
-summary(lm(formula = pov ~ lbr.part+reg,data = countries1))
-summary(lm(formula = pov ~ unemp+reg,data = countries1))
-summary(lm(formula = pov ~ trade+reg,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total+reg,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural+reg,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.urban+reg,data = countries1))
-summary(lm(formula = pov ~ gcf+reg,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt+reg,data = countries1))
-summary(lm(formula = pov ~ trade+reg,data = countries1))
-summary(lm(formula = pov ~ income+reg,data = countries1))
+summary(lm(formula = pov ~ edu.total+income +reg,data = train))
+summary(lm(formula = pov ~ hlth+income+reg,data = train))
+summary(lm(formula = pov ~ mil+income+reg,data = train))
+summary(lm(formula = pov ~ fdi+income+reg,data = train))
+summary(lm(formula = pov ~ lbr.part+income+reg,data = train))
+summary(lm(formula = pov ~ unemp+income+reg,data = train))
+summary(lm(formula = pov ~ trade+income+reg,data = train))
+summary(lm(formula = pov ~ pop.gwth.total+income+reg,data = train))
+summary(lm(formula = pov ~ pop.gwth.rural+income+reg,data = train))
+summary(lm(formula = pov ~ pop.gwth.urban+income+reg,data = train))
+summary(lm(formula = pov ~ gdp.dflt+income+reg,data = train))
+summary(lm(formula = pov ~ gcf+income+reg,data = train))
+summary(lm(formula = pov ~ trade+income+reg,data = train))
+summary(lm(formula = pov ~ gdp.pc+income+reg,data = train))
+summary(lm(formula = pov ~ lfdi+income+reg,data = train))
+summary(lm(formula = pov ~ lgdp.dflt+income+reg,data = train))
+summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train))
+```
+gdp.pc+income+reg+lgdp.pc r=0.7599
+```{r}
+summary(lm(formula = pov ~ edu.total+income +reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ hlth+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ mil+income+reg+lgdp.pc ,data = train))#0.7424
+summary(lm(formula = pov ~ fdi+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ lbr.part+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ unemp+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ trade+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ pop.gwth.total+income+reg+lgdp.pc,data = train)) #0.7433
+summary(lm(formula = pov ~ pop.gwth.rural+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ pop.gwth.urban+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ gdp.dflt+income+reg+lgdp.pc,data = train)) #0.7466
+summary(lm(formula = pov ~ gcf+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ trade+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7599
+summary(lm(formula = pov ~ lfdi+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ lgdp.dflt+income+reg+lgdp.pc,data = train))
 
@@
 ```{r}
-summary(lm(formula = pov ~ income+reg,data = countries1))
-summary(lm(formula = pov ~ hlth+reg+income,data = countries1))
-summary(lm(formula = pov ~ fdi+reg+income,data = countries1))
-summary(lm(formula = pov ~ mil+reg+income,data = countries1))
-summary(lm(formula = pov ~ lbr.part+reg+income,data = countries1))
-summary(lm(formula = pov ~ unemp+reg+income,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total+reg+income,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural+reg+income,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.urban+reg+income,data = countries1))
-summary(lm(formula = pov ~ gcf+reg+income,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt+reg+income,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income,data = countries1))
+summary(lm(formula = pov ~ edu.total+income +reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ hlth+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ mil+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ fdi+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ lbr.part+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ unemp+income+reg+lgdp.pc ,data = train))
+summary(lm(formula = pov ~ trade+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ pop.gwth.total+income+reg+lgdp.pc,data = train)) 
+summary(lm(formula = pov ~ pop.gwth.rural+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ pop.gwth.urban+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ gdp.dflt+income+reg+lgdp.pc,data = train)) 
+summary(lm(formula = pov ~ gcf+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ trade+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ lfdi+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ lgdp.dflt+income+reg+lgdp.pc,data = train))
+#highest r-square=0.7466
 ```
 ```{r}
-summary(lm(formula = pov ~ mil + reg + income, data = countries1))
-
+lm1<-(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train))
+summary(lm1)
 ```
 ```{r}
-summary(lm(formula = pov ~ pop.gwth.urban + reg + income + mil, data = countries1))
-summary(lm(formula = pov ~ hlth+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ fdi+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ lbr.part+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ unemp+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ gcf+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
+eval.metrics.linreg <- function(actual, predicted) {
+  residual <- actual - predicted
+  mse <- mean(residual ^ 2)
+  mae <- mean(abs(residual))
+  rmse <-  sqrt(mse)
+  mape <- mean(abs(residual / actual)) * 100
+
+  data.frame(
+    MSE = mse,
+    MAE = mae,
+    RMSE = rmse,
+    MAPE = mape
+  )
+}
+predicted<-predict(lm1,newdata=test)
+actual<-test$pov
+eval.metrics.linreg(actual,predicted)
+actual
+predicted
 ```
-```{r}
-countries9 <- countries1 %>% 
-  select(Age, TotalWorkingYears, NumCourses, YearsAtCurrCompany, MonthlyIncome)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-11-15 src/All variables.Rmd
I did regularisation n lasso but have some issues i feel

@@
 library(e1071) # for skewness
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(corrplot)
+library(car)
+library(DescTools)
+library(glmnet)
 ```
@@
 ```{r}
-df <-read.csv('data/countries.csv' , ,sep = "", header = TRUE)
+eval_results <- function(true, predict) {
+  actual <- data.matrix(true)
+  SSE <- sum((predict - actual)^2)
+  SST <- sum((actual - mean(actual))^2)
+  R_square <- 1 - SSE / SST
+  data.frame(
+    MSE = MSE(predict, true),
+    MAE = MAE(predict, true),
+    RMSE = RMSE(predict, true),
+    MAPE = MAPE(predict, true),
+    Rsquare = R_square
+  )
+}
+```
+
+```{r}
+df <-read.csv('../data/countries1.csv')
 df
-subset1 <- df %>% select("income","edu.total","edu.pri","edu.sec","edu.ter","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","pop.gwth.urban","gdp.dflt","gdr.eql","gcf","trade")
-
+subset1 <- df %>% select("pov","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
+subset1
 ```
@@
 ```{r}
-lm.slr <- lm(pov ~ income+edu.total+edu.pri+edu.sec+edu.ter+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gdr.eql+gcf+trade,data = train)
+lm.slr <- lm(pov ~ income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
 summary(lm.slr)
@@
 ```
+```{r}
+subset1$income <- factor(subset1$income)
+subset1$income <- as.numeric(subset1$income)
+
+```
+
+```{r}
+str(subset1)
+df0 <- subset1
+scaler <- apply(X = df0, MARGIN = 2, FUN = sd)
+subset1 <- as.data.frame(apply(df0, 2, function(x) x/sd(x, na.rm = TRUE)))
+subset1
+apply(subset1, 2, sd)
+
+```
+```{r}
+model1 <- lm(pov ~ ., data = subset1)
+summary(model1)
+vif(model1)
+
+```
+```{r}
+
+```
+```{r}
+subset2 <- subset(subset1,select =c("pov","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
+subset2 <- na.omit(subset2)
+train.x <- data.matrix(subset(subset2, select = -pov))
+train.y = data.matrix(subset2$pov)
+```
+
+```{r}
+set.seed(123)
+rate <- 0.8
+train.size <- round(nrow(subset2)*rate)
+sample(sample(nrow(subset2), train.size))
+training <- subset2[sample, ]
+test <- subset2[-sample, ]
+
+```
+```{r}
+corrplot(cor(subset2, use = "complete.obs"), 
+         method = "number",
+         type = "upper",
+         tl.col = "black", 
+         tl.srt = 30
+         )
+```
+
+```{r}
+set.seed(123)
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x,
+                    train.y,
+                    alpha = 0,
+                    lambda = cv_ridge$lambda.min)
+t(coef(glm_ridge))
+
+```
+
+```{r}
+train.x <- model.matrix(pov ~., data = training)[, -1]
+train.y <- training$pov
+test.x <- model.matrix(pov ~., data = test)[,-1]
+test.y <- test$pov
+```
+
+```{r}
+model_ridge_trial1 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.1)
+t(coef(model_ridge_trial1))
+```
+
+```{r}
+model_ridge_trial2 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.5)
+t(coef(model_ridge_trial2))
+```
+
+```{r}
+model_ridge_trial3 <- glmnet(train.x, train.y, alpha = 0, lambda = 1)
+t(coef(model_ridge_trial3))
+```
+
+```{r}
+table_summary <- cbind(coef(model_ridge_trial1),
+                       coef(model_ridge_trial2),
+                       coef(model_ridge_trial3))
+
+colnames(table_summary) <- c("lambda = 0.1",
+                             "lambda = 0.5",
+                             "lambda = 1")
+
+table_summary
+```
+
+```{r}
+set.seed(123)
+cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
+glm_lasso <- glmnet(train.x,train.y, 
+                    alpha = 1, lambda = cv_lasso$lambda.min)
+t(coef(glm_lasso))
+```
+
+```{r}
+# Obtain performance of MLR model
+#fit <- model1$fitted.values
+#true <- subset1$Price
+#summary_model1 <- eval_results(true, fit)
+
+# Obtain performance of ridge model
+fit <- predict(glm_ridge, newx = train.x)
+true <- train.y
+summary_ridge <- eval_results(true, fit)
+
+# Obtain performance of LASSO model
+fit <- predict(glm_lasso, newx = train.x)
+true <- train.y
+summary_lasso <- eval_results(true, fit)
+
+# Combine all performances into a single output
+summary <- rbind(summary_ridge, summary_lasso)
+rownames(summary) <- c( "Ridge_model", "LASSO_model" )
+colnames(summary)[5] <- "R^2"
+knitr::kable(summary, digits = 2)
+```

### 2022-11-15 src/draft.Rmd
imputation + regularization

@@
 output:
-  pdf_document: default
   html_document:
     df_print: paged
+  pdf_document: default
 bibliography: citations.bib
@@
 ```{r, message = F}
-# Package names
-# packages <- c("MASS", "knitr", "readr", "tidyr", "dplyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "needs")
-# 
-# # Install packages not yet installed
-# installed_packages <- packages %in% rownames(installed.packages())
-# if (any(installed_packages == FALSE)) {
-#   install.packages(packages[!installed_packages])
-# }
-# 
-# # Packages loading
-# invisible(lapply(packages, library, character.only = TRUE))
-# 
-# prioritize(dplyr)
-
 library(needs)
 
-needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet")
+needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce")
 
@@
 ```{r}
-formula.str <- 'pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'
+formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'
 
@@
 # dummy lm model
-fit2.dummy <- lm(as.formula(formula.str), data = countries1)
+fit2.dummy <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
 # replace coefficients of dummy model & predict
@@
    RSS <- sum((pred - orig)^2)
-   TSS <- sum((pred - mean(orig))^2)
+   TSS <- sum((orig - mean(orig))^2)
    R2 <- 1 - RSS/TSS
@@
 
-Prepare data
+Helper functions to standardise data.
 
@@
 standardise <- function(train, test) {
+  # save number of train and combine both data set
+  trainCases <- 1:nrow(train)
   data <- rbind(train, test)
+  # calc scaler
   scaler <- unlist(lapply(data, sd0))
-
+  # get the numeric columns
   numCols <- which(unlist(lapply(data, is.numeric)))
+  # divide numeric columns by scalers, and combine with factor columns
   num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
   fct <- data[,-numCols]
-  return(cbind(fct, num))
+  final <- cbind(fct, num)
+  # split to train and test
+  trainSet <- final[trainCases,]
+  testSet <- final[-trainCases,]
+
+  res <- list(final = final, train = trainSet, test = testSet, scaler = scaler)
+  return(res)
 }
@@
 
+Helper function to build the optimal model.
+
 ```{r}
-optimalModel <- function(train) {
-  #standardized
+optimalModel <- function(formula, train, test) {
+  # standise and save number of train cases (to split later)
+  std.data <- standardise(train, test)
+  trainCases <- 1:nrow(train)
+  # matrix-ise
+  xs <- model.matrix(formula, std.data$final)[,-1]
+  ys <- std.data$final$pov
+  # split data
+  train.x <- xs[trainCases,]
+  train.y <- ys[trainCases]
+  test.x <- xs[-trainCases,]
+  test.y <- ys[-trainCases]
+  # list of tested alpha, resolution = 0.1
+  alphas <- seq(0, 1, 0.1)
+  # build a bunch of models
+  models <- lapply(alphas, \(a) cv.glmnet(train.x, train.y, type.measure = "mse", alpha = a))
+  cv.error <- unlist(lapply(models, \(model) model$cvm[model$lambda == model$lambda.min]))
+  # best model
+  best.model.idx <- which.min(cv.error)
+  # optimal alpha and lambda
+  alpha.opt <- alphas[best.model.idx]
+  lambda.opt <- models[[best.model.idx]]$lambda.min
+  best.model <- glmnet(train.x, train.y, alpha = alpha.opt, lambda = lambda.opt)
+  # predict for test data
+  train.fit <- predict(best.model, train.x)
+  test.fit <- predict(best.model, test.x)
+  # evaluation - train
+  k <- best.model$df
+  train.r2 <- best.model$dev.ratio
+  train.adjR2 <- adjR2(train.fit, train.y, k)
+  train.resid <- train.fit - train.y
+  train.mse <- mean(train.resid^2)
+  train.rmse <- sqrt(train.mse)
+  train.mae <- mean(abs(train.resid))
+  train.mape <- mean(abs(train.resid / train.y))
+  # evaluation - test
+  test.r2 <- r2(test.fit, test.y)
+  test.adjR2 <- adjR2(test.fit, test.y, k)
+  test.resid <- test.fit - test.y
+  test.mse <- mean(test.resid^2)
+  test.rmse <- sqrt(test.mse)
+  test.mae <- mean(abs(test.resid))
+  test.mape <- mean(abs(test.resid / test.y))
+  # final results
+  results <- list(
+    train = list(
+      data = train,
+      r2 = train.r2,
+      adj.r2 = train.adjR2,
+      residuals = train.resid,
+      mse  = train.mse,
+      rmse = train.rmse,
+      mae  = train.mae,
+      mape = train.mape
+    ),
+    test = list(
+      data = test,
+      r2 = test.r2,
+      adj.r2 = test.adjR2,
+      residuals = test.resid,
+      mse  = test.mse,
+      rmse = test.rmse,
+      mae  = test.mae,
+      mape = test.mape
+    ),
+    scaler = scaler,
+    alpha = alpha.opt,
+    lambda = lambda.opt,
+    model = best.model
+  )
+
+  return(results)
 }
 ```
+
+Build the optimal models using each imputated data set.
+
+```{r}
+# helper function to build from a specific imputated data
+buildWith <- function(set, imputed, test, formula, fun) {
+  train <- complete(imputed, set)
+  model <- fun(formula, train, test)
+  return(model)
+}
+
+set.seed(1984)
+
+models <- lapply(1:countries1.imputed$m, buildWith, imputed = countries1.imputed, test = countries.test.4, formula = as.formula(formula.str), fun = optimalModel)
+```
+
+Display parameters and evaluation.
+
+```{r}
+result.list <- lapply(models, \(mod) {
+  data.frame(
+    alpha = mod$alpha,
+    lambda = mod$lambda,
+    `Train MSE` = mod$train$mse,
+    `Test MSE` = mod$test$mse,
+    `Train MAE` = mod$train$mae,
+    `Test MAE` = mod$test$mae,
+    `Train R2` = mod$train$r2,
+    `Test R2` = mod$test$r2,
+    `Train Adj.R2` = mod$train$adj.r2,
+    `Test Adj.R2` = mod$test$adj.r2
+  )
+})
+
+result.df <- do.call(rbind.data.frame, result.list)
+result.df <- cbind(data.frame(Set = 1:countries1.imputed$m), result.df)
+kable(result.df)
+```
+
+It seems to have very little impact on the adjusted R-Squared of train data and test data.
 

### 2022-11-15 src/draft.Rmd
imputation (regularization, vif and aic) git pull

@@
 
-needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce")
+needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet")
 
@@
 ggplot(countries, aes(x = pov)) +
-  geom_histogram(aes(y = ..density..)) +
+  geom_histogram(aes(y = ..density..), bins = 30) +
   geom_density(alpha = .2, fill = "blue")
@@
 ggplot(poverty.headcount, aes(x = year)) +
-  geom_histogram()
+  geom_histogram(bins = 30)
 
@@
 ggplot(pov.98.18, aes(x = pov)) +
-  geom_histogram(aes(y = ..density..)) +
+  geom_histogram(aes(y = ..density..), bins = 30) +
   geom_density(alpha = .2, fill = "blue") +
@@
   geom_point() +
-  geom_smooth()
+  geom_smooth(formula = y~x, method = loess)
 ```
@@
   geom_point() +
-  geom_smooth() +
+  geom_smooth(formula = y~x, method = loess) +
   facet_grid(rows = vars(type))
@@
     geom_point() +
-    geom_smooth(method = lm) +
+    geom_smooth(formula = y~x, method = lm) +
     labs(title = paste("Relationship pov ~", indvar),
@@
 a variable).
-The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), or predict the value using other variables.
-
-We use **Multivariate Imputation By Chained Equations (MICE)** as our primary tool for this technique, because of its wide acceptance in scientific studies (@imputWorks).
+The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), and use other variables to predict the missing values (much like regression).
+Since we observed some correlations during descriptive analysis, we expect the imputation algorithm to work well.
+
+We use **Multivariate Imputation By Chained Equations (MICE)** as primary tool for this technique, because of its wide acceptance in scientific studies (@imputWorks).
 In order to use this technique, we have to assume that the "missingness" of a field can be explained by the values in other columns, (e.g., If the countries is in North America, it's more likely to be missing as we have discovered in 2.2).
-The general ideas of the algorithm is iteratively predict the missing values based on other variables, and improve it after each iterations until predicted value converse to a stable point.
-The algorithmic details of MICE is very concisely explained in @micealgo.
-
-Our parameter for this algorithm is fairly standard:
+The general ideas of the algorithm is to fill in the missing values, and improve it iteratively until predicted values converse to a stable point.
+The algorithmic details of MICE is very concisely (and enthrallingly) explained in @micealgo.
+
+We use relatively small parameters in the interest of time.
 
@@
 ##### A. Ordinary Linear Regression
-We generated 5 sets of imputated (training) data.
-We can build our model on each of these data sets and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
+We generated 3 sets of imputated (training) data.
+We can build separate models using each data set, and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
 With all the generated data sets.
@@
 
-# Using only the first imputed data set
-fit1 <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
-summary(fit1)
-fit1.pred <- predict(fit1, countries.test.4)
+# Using single imputed data set
+fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
 
@@
 fit2.combined <- pool(fit2)
-summary(fit2.combined)
 # dummy lm model
 fit2.dummy <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
-# replace coefficients of dummy model & predict
+# replace coefficients of dummy model
+name.coef <- names(fit2.dummy$coefficients)
 fit2.dummy$coefficients <- fit2.combined$pooled$estimate
-fit2.pred <- predict(fit2.dummy, countries.test.4)
+names(fit2.dummy$coefficients) <- name.coef
 ```
@@
 ```{r}
-k <- length(fit1$coefficients) - 1
-fit1.train.r2 <- summary(fit1)$adj.r.squared
-fit1.test.r2 <- adjR2(fit1.pred, countries.test.4$pov, k)
-fit2.train.r2 <- pool.r.squared(fit2, adjusted = T)[,'est']
-fit2.test.r2 <- adjR2(fit2.pred, countries.test.4$pov, k)
-res <- data.frame(model = c("Single", "Pooled"), train = c(fit1.train.r2, fit2.train.r2), test = c(fit1.test.r2, fit2.test.r2))
+fit1.summ <- lapply(fit1, \(f) {
+  # number of variables
+  k <- length(f$coefficients) - 1
+  # predict value of test
+  pred <- predict(f, countries.test.4)
+  resid <- pred - countries.test.4$pov
+  data.frame(
+    `Train MSE` = mean(summary(f)$residuals^2),
+    `Test MSE` = mean(resid^2),
+    `Train MAE` = mean(abs(summary(f)$residuals)),
+    `Test MAE` = mean(abs(test.resid)),
+    `Train Adj.R2` = summary(f)$adj.r.squared,
+    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
+  )
+})
+
+# number of variables
+k <- length(fit2.dummy$coefficients) - 1
+# number of variables
+fit2.pred <- predict(fit2.dummy, countries.test.4)
+resid <- fit2.pred - countries.test.4$pov
+fit2.summ <- data.frame(
+  `Train MSE` = mean(summary(fit2.dummy)$residuals^2),
+  `Test MSE` = mean(resid^2),
+  `Train MAE` = mean(abs(summary(fit2.dummy)$residuals)),
+  `Test MAE` = mean(abs(test.resid)),
+  `Train Adj.R2` = pool.r.squared(fit2, adjusted = T)[,'est'],
+  `Test Adj.R2` = adjR2(fit2.pred, countries.test.4$pov, k)
+)
+
+res <- do.call(rbind.data.frame, fit1.summ)
+res <- rbind(res, fit2.summ)
+res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)
+
 kable(res)
@@
 
-The models seem to be over-fitting. We should perform model simplification.
-
-##### B. Regularization
+Pooled model slightly improve performance. Test data fitting has better MSE and MAE than train data. However, Adjusted R-Squared is lower in test data. There might be some over-fitting in our models.
+
+```{r}
+# find coef of variable other than country.code
+var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","pop.gwth.urban","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
+
+res <- do.call(rbind, lapply(fit1, \(f) f$coefficients[var]))
+res <- rbind(res, fit2.dummy$coefficients[var])
+res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)
+kable(res)
+```
+
+In contradiction to expectation, `hlth` is positively correlated with `pov`, and lower-middle income countries (incomeLM) are less poor than high income countries (incomeH).
+`pop.gwth.total` is negatively related to `pov`, while `pop.gwth.rural` and `pop.gwth.urban` are positively related.
+They might be indications of over-fitting model.
+
+```{r}
+# fit model on imputated set no. 1
+summary(fit1[[1]])
+```
+
+There are also some weak variable, we should perform some variable selection.
+
+##### B. VIF
+
+Our data set contains a lots of variables. We can perform some variable selection to reduce over-fitting.
+
+```{r}
+gvif <- lapply(fit1, vif)
+gvif
+```
+
+[Some forum](https://stats.stackexchange.com/questions/70679/which-variance-inflation-factor-should-i-be-using-textgvif-or-textgvif) suggested that we should use the standard $GVIF^{1/(2 \cdot df)} < 2$ as equivalent to $GVIF < 4$ to account for high degree of freedom of some variables.
+
+```{r}
+# adjusted gvif higher than 3
+lapply(gvif, \(g) {
+  g[g[,3] > 3, 3]
+})
+```
+
+Remove the highest-ranked variables (`pop.gwth.total`, `pop.gwth.urban`, `pop.gwth.rural`, `gdp.pc`, `lgdp.pc`) and rebuild the models.
+
+```{r}
+formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+trade+lfdi+lgdp.dflt'
+
+# Using single imputed data set
+fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
+```
+
+Re-run GVIF analysis
+
+```{r}
+gvif <- lapply(fit1, vif)
+
+lapply(gvif, \(g) {
+  g[g[,3] > 2, 3]
+})
+```
+
+There're still some variable with adjusted GVIF > 2.
+We are interested in the effects of `hlth` (expenditure in Healthcare), and `mil` (expenditure in Military), so we won't remove those variables.
+We can remove `trade`.
+
+```{r}
+# remove trade
+formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+lfdi+lgdp.dflt'
+
+# Using single imputed data set
+fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
+
+# gvif
+gvif <- lapply(fit1, vif)
+
+lapply(gvif, \(g) {
+  g[g[,3] > 2, 3]
+})
+```
+
+The adjusted GVIF are acceptably low.
+
+Coefficients overview.
+
+```{r}
+# find coef of variable other than country.code
+var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","fdi","lbr.part","unemp","gdp.dflt","gcf","lfdi","lgdp.dflt")
+
+res <- do.call(rbind, lapply(fit1, \(f) f$coefficients[var]))
+kable(res)
+```
+
+There are several sign-switching in `income` coefficients, indicating the effect of removing some multicollinearity.
+
+```{r}
+fit1.summ <- lapply(fit1, \(f) {
+  k <- length(f$coefficients) - 1
+  pred <- predict(f, countries.test.4)
+  data.frame(
+    `Train Adj.R2` = summary(f)$adj.r.squared,
+    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
+  )
+})
+
+res <- do.call(rbind.data.frame, fit1.summ)
+res <- cbind(data.frame(Set = 1:3), res)
+kable(res)
+```
+
+Directly removing variables seems to be penalising our test results.
+As our interest is to investigate the effect of certain variables on poverty, the yielded R-Squared is within acceptable range.
+We can continue variable selection on the new models.
+
+```{r}
+formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+lfdi+lgdp.dflt'
+```
+
+##### B. Step-wise AIC
+We can conduct a step-wise AIC variable selection.
+It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models.
+The lower AIC is, the better the model fits.
+\\
+
+```{r}
+# helper
+fitAIC <- function(i) {
+  set.seed(1984)
+  fit <- lm(as.formula(formula.str), complete(countries1.imputed, i))
+  aic.fit <- stepAIC(fit, trace = T, direction = 'backward')
+  return(aic.fit)
+}
+
+# build models
+aic.fits <- lapply(1:countries1.imputed$m, fitAIC)
+
+# check final terms of each model
+lapply(aic.fits, \(mod) formula(mod$terms))
+```
+
+The most commonly removed variables are:
+
+- `gdp.dflt`
+- `ldgp.dflt`
+- `unemp`
+- `fdi`
+
+```{r}
+# predict with each model
+aic.pred <- lapply(aic.fits, predict, newdata = countries.test.4)
+# calculate adj r2
+aic.train.r2 <- unlist(lapply(aic.fits, \(model) summary(model)$adj.r.squared))
+
+k <- lapply(aic.fits, \(m) length(m$coefficients))
+aic.test.r2 <- unlist(mapply(adjR2, aic.pred, k = k, MoreArgs = list(orig = countries.test.4$pov)))
+
+res <- data.frame(`set no.` = 1:3, train = aic.train.r2, test = aic.test.r2)
+```
+
+We can compare Adjusted R-Squared to estimate relative over-fitting in the new models.
+
+```{r}
+kable(res)
+```
+
+Performance on train set and test set are not affected. However, we have simplified the model by removing some unnecessary terms. We can update our models.
+
+```{r}
+formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+lbr.part+gcf+lfdi'
+
+fit3 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
+
+# Build models with all generated data set and pool the estimates
+fit4 <- with(data = countries1.imputed, exp = lm(as.formula(formula.str)))
+fit4.combined <- pool(fit4)
+# dummy lm model
+fit4.dummy <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
+# replace coefficients of dummy model & predict
+fit4.dummy$coefficients <- fit4.combined$pooled$estimate
+
+fit3.summ <- lapply(fit3, \(f) {
+  k <- length(f$coefficients) - 1
+  pred <- predict(f, countries.test.4)
+  resid <- pred - countries.test.4$pov
+  data.frame(
+    `Train MSE` = mean(summary(f)$residuals^2),
+    `Test MSE` = mean(resid^2),
+    `Train MAE` = mean(abs(summary(f)$residuals)),
+    `Test MAE` = mean(abs(test.resid)),
+    `Train Adj.R2` = summary(f)$adj.r.squared,
+    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
+  )
+})
+
+k <- length(fit4.dummy$coefficients) - 1
+fit4.pred <- predict(fit4.dummy, countries.test.4)
+resid <- fit4.pred - countries.test.4$pov
+fit4.summ <- data.frame(
+  `Train MSE` = mean(summary(fit4.dummy)$residuals^2),
+  `Test MSE` = mean(resid^2),
+  `Train MAE` = mean(abs(summary(fit4.dummy)$residuals)),
+  `Test MAE` = mean(abs(test.resid)),
+  `Train Adj.R2` = pool.r.squared(fit4, adjusted = T)[,'est'],
+  `Test Adj.R2` = adjR2(fit4.pred, countries.test.4$pov, k)
+)
+
+res <- do.call(rbind.data.frame, fit3.summ)
+res <- rbind(res, fit4.summ)
+res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)
+
+kable(res)
+```
+
+Coefficients overview.
+
+```{r}
+var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","lbr.part","gcf","lfdi")
+
+res <- do.call(rbind, lapply(fit3, \(f) f$coefficients[var]))
+kable(res)
+```
+
+We can further treat over-fitting with regularization.
+
+##### C. Regularization
 
@@
 optimalModel <- function(formula, train, test) {
+  set.seed(1984)
   # standise and save number of train cases (to split later)
@@
     ),
-    scaler = scaler,
+    scaler = std.data$scaler,
     alpha = alpha.opt,
@@
 
-It seems to have very little impact on the adjusted R-Squared of train data and test data.
-
-##### B. Step-wise AIC
-We can conduct a step-wise AIC variable selection.
-It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models.
-The lower AIC is, the better the model fits.
-\\
-We have no idea how to pool coefficients for AIC model,
-
-```{r}
-# helper
-fitAIC <- function(i) {
-  fit <- lm(as.formula(formula.str), complete(countries1.imputed, i))
-  aic.fit <- stepAIC(fit, trace = F, direction = 'backward')
-  return(aic.fit)
-}
-
-# build models
-aic.fits <- lapply(1:countries1.imputed$m, fitAIC)
-# predict with each model
-aic.pred <- lapply(aic.fits, predict, newdata = countries.test.4)
-# calculate adj r2
-aic.train.r2 <- unlist(lapply(aic.fits, \(model) summary(model)$adj.r.squared))
-
-k <- lapply(aic.fits, \(m) length(m$coefficients))
-aic.test.r2 <- unlist(mapply(adjR2, aic.pred, k = k, MoreArgs = list(orig = countries.test.4$pov)))
-
-res <- data.frame(`set no.` = 1:3, train = aic.train.r2, test = aic.test.r2)
-```
-
-Adjusted R-Squared
-
-```{r}
+There are some improvement on the test data performance based on adjusted R-Squared. MSE and MAE remain low on both data sets.
+
+Coefficients overview.
+
+```{r}
+var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","lbr.part","gcf","lfdi")
+
+res <- do.call(rbind, lapply(models, \(f) coefficients(f$model)[var,]))
 kable(res)
@@
 
--> AIC on one model [link](https://rforpoliticalscience.com/2020/10/23/choose-model-variables-by-aic-in-a-stepwise-algorithm-with-the-mass-package-in-r/)
-
--> stepwise selection on mice  [link](https://stefvanbuuren.name/fimd/sec-stepwise.html)
+All non-country.code variables are non-zero. We can observe that the regularisation models just remove the effect of `country.code`, for countries that have similar "baseline".
 

### 2022-11-15 data/forward.Rmd

--- 2022-11-14 data/forward.Rmd
@@
 ```{r}
+df_count<-read.csv("../data/countries1.csv")
+head(df_count)
 
+set.seed(10)
+index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
+train <- df_count[index, ]
+test <- df_count[-index, ]
 ```
 
+Through forward selection, we will do feature selection, through iteration of different models, a linear model consisting of income as a predictor has the highest adjusted r-square value=0.6105.
+
+Below are some of the linear models that have been explored.
 ```{r cars}
-summary(lm(formula = pov ~ hlth,data = countries1))
-summary(lm(formula = pov ~ fdi,data = countries1))
-summary(lm(formula = pov ~ reg,data = countries1))
-summary(lm(formula = pov ~ mil,data = countries1))
-summary(lm(formula = pov ~ lbr.part,data = countries1))
-summary(lm(formula = pov ~ unemp,data = countries1))
-summary(lm(formula = pov ~ trade,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ gcf,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt,data = countries1))
-summary(lm(formula = pov ~ trade,data = countries1))
+summary(lm(formula = pov ~ income,data = train)) #0.6105
+summary(lm(formula = pov ~ reg,data = train)) #0.5287
+summary(lm(formula = pov ~ edu.total,data = train)) #0.09506
+summary(lm(formula = pov ~ hlth,data = train)) #0.09806
+```
+Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
+reg as a predictor has the highest adjusted r-value=0.7031
+
+```{r}
+summary(lm(formula = pov ~ reg+income,data = train))
+summary(lm(formula = pov ~ edu.total+income,data = train))
+summary(lm(formula = pov ~ hlth+income,data = train))
+summary(lm(formula = pov ~ mil+income,data = train))
+summary(lm(formula = pov ~ fdi+income,data = train))
+```
+
+The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7364
+```{r}
+summary(lm(formula = pov ~ edu.total+income +reg,data = train))
+summary(lm(formula = pov ~ hlth+income+reg,data = train))
+summary(lm(formula = pov ~ lgdp.dflt+income+reg,data = train))
+summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train))
+```
+gdp.pc+income+reg+lgdp.pc r=0.7599
+```{r}
+summary(lm(formula = pov ~ mil+income+reg+lgdp.pc ,data = train))#0.7424
+summary(lm(formula = pov ~ pop.gwth.total+income+reg+lgdp.pc,data = train)) #0.7433
+summary(lm(formula = pov ~ gdp.dflt+income+reg+lgdp.pc,data = train)) #0.7466
+summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7599
+```
+Highest adjusted r-square=0.7466 which is lower than our previous model
+```{r}
+summary(lm(formula = pov ~ trade+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ pop.gwth.total+income+reg+lgdp.pc,data = train)) 
+summary(lm(formula = pov ~ pop.gwth.rural+income+reg+lgdp.pc,data = train))
+summary(lm(formula = pov ~ pop.gwth.urban+income+reg+lgdp.pc,data = train)) #0.7466
 ```
 ```{r}
-summary(lm(formula = pov ~ hlth+reg,data = countries1))
-summary(lm(formula = pov ~ fdi+reg,data = countries1))
-summary(lm(formula = pov ~ mil+reg,data = countries1))
-summary(lm(formula = pov ~ lbr.part+reg,data = countries1))
-summary(lm(formula = pov ~ unemp+reg,data = countries1))
-summary(lm(formula = pov ~ trade+reg,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total+reg,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural+reg,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.urban+reg,data = countries1))
-summary(lm(formula = pov ~ gcf+reg,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt+reg,data = countries1))
-summary(lm(formula = pov ~ trade+reg,data = countries1))
-summary(lm(formula = pov ~ income+reg,data = countries1))
+lm1<-(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train))
+summary(lm1)
+```
+Assessment
+```{r}
+eval.metrics.linreg <- function(actual, predicted) {
+  residual <- actual - predicted
+  mse <- mean(residual ^ 2)
+  mae <- mean(abs(residual))
+  rmse <-  sqrt(mse)
+  mape <- mean(abs(residual / actual)) * 100
 
+  data.frame(
+    MSE = mse,
+    MAE = mae,
+    RMSE = rmse,
+    MAPE = mape
+  )
+}
+predicted<-predict(lm1,newdata=test)
+actual<-test$pov
+eval.metrics.linreg(actual,predicted)
+actual
+predicted
 ```
-```{r}
-summary(lm(formula = pov ~ income+reg,data = countries1))
-summary(lm(formula = pov ~ hlth+reg+income,data = countries1))
-summary(lm(formula = pov ~ fdi+reg+income,data = countries1))
-summary(lm(formula = pov ~ mil+reg+income,data = countries1))
-summary(lm(formula = pov ~ lbr.part+reg+income,data = countries1))
-summary(lm(formula = pov ~ unemp+reg+income,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total+reg+income,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural+reg+income,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.urban+reg+income,data = countries1))
-summary(lm(formula = pov ~ gcf+reg+income,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt+reg+income,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income,data = countries1))
-```
-```{r}
-summary(lm(formula = pov ~ mil + reg + income, data = countries1))
-
-```
-```{r}
-summary(lm(formula = pov ~ pop.gwth.urban + reg + income + mil, data = countries1))
-summary(lm(formula = pov ~ hlth+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ fdi+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ lbr.part+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ unemp+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.total+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ pop.gwth.rural+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ gcf+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ gdp.dflt+reg+income+mil+pop.gwth.urban,data = countries1))
-summary(lm(formula = pov ~ trade+reg+income+mil+pop.gwth.urban,data = countries1))
-```
-```{r}
-countries9 <- countries1 %>% 
-  select(Age, TotalWorkingYears, NumCourses, YearsAtCurrCompany, MonthlyIncome)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-11-16 src/All variables.Rmd
Update All variables.Rmd

@@
 ```{r}
-
-```
-```{r}
 subset2 <- subset(subset1,select =c("pov","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
@@
 ```
+
 ```{r}
@@
 # Obtain performance of MLR model
-#fit <- model1$fitted.values
-#true <- subset1$Price
-#summary_model1 <- eval_results(true, fit)
+fit <- model1$fitted.values
+true <- subset2$pov
+summary_model1 <- eval_results(true, fit)
 
@@
 # Combine all performances into a single output
-summary <- rbind(summary_ridge, summary_lasso)
-rownames(summary) <- c( "Ridge_model", "LASSO_model" )
+summary <- rbind(summary_model1, summary_ridge, summary_lasso)
+rownames(summary) <- c("MLR", "Ridge_model", "LASSO_model" )
 colnames(summary)[5] <- "R^2"
-knitr::kable(summary, digits = 2)
+knitr::kable(summary, digits = 3)
 ```

### 2022-11-16 src/draft.Rmd
imputation evaluation, docs restruct

@@
 output:
+  pdf_document: default
   html_document:
     df_print: paged
-  pdf_document: default
+  word_document: default
 bibliography: citations.bib
@@
 ```{r setup, include=FALSE}
+library(knitr)
 knitr::opts_chunk$set(echo = TRUE, warning = F)
 knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
+# inline hook to 2 digits
+inline_hook <- function(x) {
+    if(is.numeric(x)) x <- round(x, 2)
+    paste(as.character(x), collapse=", ")
+}
+knit_hooks$set(inline = inline_hook)
+# code chunk out put to 
+options(digits = 5)
 ```
@@
 
-needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet")
+needs("dplyr", "MASS", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce", "lmtest")
 
@@
 
-## 2.4. Data Source
+## 2.4. Assumption Check
+We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data.
+[Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
+**1. Linear relationship:**  
+Use correlation matrix to check linearity
+```{r, warning = FALSE}
+# select only numeric data
+countries.num <- countries %>% select(where(is.numeric))
+chart.Correlation(countries.num, histogram = TRUE, pch = 19)
+```
+Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
+```{r}
+# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
+countries.num2 <- countries.num %>% 
+  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
+  filter(!is.na(value) & !is.na(pov))
+
+drawGraph <- function(indvar, data) {
+  ggplot(data %>% 
+           filter(variable == indvar),
+         aes(x = value, y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste("Relationship pov ~", indvar),
+         x = indvar,
+         y = "pov")
+}
+
+ivs <- distinct(countries.num2, variable)[[1]]
+
+# for (indvar in ivs) {
+#   print(
+#     ggplot(countries.num2 %>% 
+#            filter(variable == indvar),
+#          aes(x = value, y = pov)) +
+#     geom_point() +
+#     geom_smooth(formula = y~x, method = lm) +
+#     labs(title = paste("Relationship pov ~", indvar),
+#          x = indvar,
+#          y = "pov")
+#   )
+# }
+
+p <- ggplot(countries.num2, 
+     aes(x = value, y = pov)) +
+geom_point() +
+geom_smooth(formula = y~x, method = lm) +
+labs(title = paste("pov ~ variable"),
+     x = "variable",
+     y = "pov") +
+facet_wrap_paginate(~ variable,
+                    ncol = 3,
+                    nrow = 1,
+                    scales = "free")
+
+requiredPages <- n_pages(p)
+
+for (i in 1:requiredPages) {
+  ggplot(countries.num2,
+         aes(x = value, y = pov)) +
+  geom_point() +
+  geom_smooth(formula = y~x, method = lm) +
+  labs(title = paste("pov ~ variable"),
+       x = "variable",
+       y = "pov") +
+  facet_wrap_paginate(~ variable,
+                      ncol = 3,
+                      nrow = 1,
+                      scales = "free",
+                      page = i) -> p
+  print(p)
+}
+```
+
+Most variables display linear relationship with some exceptions, which are more appropriate to assume a logarithmic relationship. Gender equality should be viewed as a categorical data.
+```{r}
+logIvs <- c("fdi", "gdp.dflt", "gdp.pc")
+
+for (indvar in logIvs) {
+  print(
+    ggplot(countries.num2 %>% 
+           filter(variable == indvar),
+         aes(x = log(value), y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste0("pov ~ log(", indvar, ")"),
+         x = paste0("log(", indvar, ")"),
+         y = "pov")
+  )
+}
+```
+
+The relationship are more linear after the transformation.
+
+```{r}
+gdr.eql <- countries.num2 %>% 
+  filter(variable == "gdr.eql") %>% 
+  mutate(value = factor(value))
+
+ggplot(gdr.eql,
+       aes(x = value, y = pov)) +
+  geom_boxplot() +
+  labs(title = paste("pov ~ gdr.eql"),
+       x = "gdr.eql",
+       y = "pov")
+```
+There's a significant correlation between gender equality and poverty.  
+Correlation coefficients between `pov` and predictors.
+
+```{r}
+# transform some variables
+countries.num3 <- countries.num %>% 
+  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
+  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
+
+# name of the independent variables
+cn <- colnames(countries.num3)[-c(1,2)]
+
+# correlation of independent variable and pov
+corWithPov <- function(indevar, data) {
+  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
+}
+
+cor.pov <- sapply(cn, corWithPov, data = countries.num3)
+kable(cor.pov)
+```
+
+`edu.ter`, `mil`, `fdi`, `lbr.part`, `unemp`, `gdp.dflt`, `gcf` have negligible correlation with `pov`.
+`lgdp.pc`, `lfdi`, and `lgdp.dflt` have stronger linear relationship with `pov` than their un-transformed counterparts.
+
+```{r}
+countries1 <- countries1 %>% 
+  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
+  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
+```
+
+**2. Predictors are independent**
+There should be no correlation/multicolinearity between each pair of predictors.
+
+```{r}
+countries.arr <- simplify2array(countries.num %>% 
+                                  select(-c("year", "pov")))
+
+cor.mtx <- rcorr(countries.arr, type = "spearman")
+
+round(cor.mtx$r, 2)
+
+corrplot(cor.mtx$r, type = "upper", order = "hclust", 
+         tl.col = "black", tl.srt = 45)
+```
+Some significant correlations can be found between `gdr.eql` with `edu.pri`, `pop.gwth.total`, and `fdi` with `mil`, etc.
+We expect these variables to be eliminated in the VIF test.
+This might be a good property for imputation (3.2.4)
+
+## 2.5. Data Source
 
@@
 # 3. Model Selection and Interpretation
-## 3.1. Assumption Check
-We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data.
-[Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
-**1. Linear relationship:**  
-Use correlation matrix to check linearity
-```{r, warning = FALSE}
-# select only numeric data
-countries.num <- countries %>% select(where(is.numeric))
-chart.Correlation(countries.num, histogram = TRUE, pch = 19)
-```
-Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
-```{r}
-# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
-countries.num2 <- countries.num %>% 
-  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
-  filter(!is.na(value) & !is.na(pov))
-
-drawGraph <- function(indvar, data) {
-  ggplot(data %>% 
-           filter(variable == indvar),
-         aes(x = value, y = pov)) +
-    geom_point() +
-    geom_smooth(method = lm) +
-    labs(title = paste("Relationship pov ~", indvar),
-         x = indvar,
-         y = "pov")
-}
-
-ivs <- distinct(countries.num2, variable)[[1]]
-
-for (indvar in ivs) {
-  print(
-    ggplot(countries.num2 %>% 
-           filter(variable == indvar),
-         aes(x = value, y = pov)) +
-    geom_point() +
-    geom_smooth(formula = y~x, method = lm) +
-    labs(title = paste("Relationship pov ~", indvar),
-         x = indvar,
-         y = "pov")
-  )
-}
-```
-
-Most variables display linear relationship with some exceptions, which are more appropriate to assume a logarithmic relationship. Gender equality should be viewed as a categorical data.
-```{r}
-logIvs <- c("fdi", "gdp.dflt", "gdp.pc")
-
-for (indvar in logIvs) {
-  print(
-    ggplot(countries.num2 %>% 
-           filter(variable == indvar),
-         aes(x = log(value), y = pov)) +
-    geom_point() +
-    geom_smooth(method = lm) +
-    labs(title = paste0("Relationship pov ~ log(", indvar, ")"),
-         x = paste0("log(", indvar, ")"),
-         y = "pov")
-  )
-}
-```
-
-The relationship are more linear after the transformation.
-
-```{r}
-gdr.eql <- countries.num2 %>% 
-  filter(variable == "gdr.eql") %>% 
-  mutate(value = factor(value))
-
-ggplot(gdr.eql,
-       aes(x = value, y = pov)) +
-  geom_boxplot() +
-  geom_smooth(method = lm) +
-  labs(title = paste("Relationship pov ~ gdr.eql"),
-       x = "gdr.eql",
-       y = "pov")
-```
-There's a significant correlation between gender equality and poverty.  
-Correlation coefficients between `pov` and predictors.
-
-```{r}
-# transform some variables
-countries.num3 <- countries.num %>% 
-  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
-  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
-
-# name of the independent variables
-cn <- colnames(countries.num3)[-c(1,2)]
-
-# correlation of independent variable and pov
-corWithPov <- function(indevar, data) {
-  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
-}
-
-cor.pov <- sapply(cn, corWithPov, data = countries.num3)
-kable(cor.pov)
-```
-
-`edu.ter`, `mil`, `fdi`, `lbr.part`, `unemp`, `gdp.dflt`, `gcf` have negligible correlation with `pov`.
-`lgdp.pc`, `lfdi`, and `lgdp.dflt` have stronger linear relationship with `pov` than their un-transformed counterparts.
-
-```{r}
-countries1 <- countries1 %>% 
-  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
-  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
-```
-
-**2. Predictors are independent**
-There should be no correlation/multicolinearity between each pair of predictors.
-
-```{r}
-countries.arr <- simplify2array(countries.num %>% 
-                                  select(-c("year", "pov")))
-
-cor.mtx <- rcorr(countries.arr, type = "spearman")
-
-round(cor.mtx$r, 2)
-
-corrplot(cor.mtx$r, type = "upper", order = "hclust", 
-         tl.col = "black", tl.srt = 45)
-```
-Some significant correlations can be found between `gdr.eql` with `edu.pri`, `pop.gwth.total`, and `fdi` with `mil`, etc.
-We expect these variables to be eliminated in the VIF test.
-This might be a good property for imputation (3.2.4)
-
-## 3.2. Ordinary Multiple Linear Regression
-
-We conduct a normal linear regression, following the approaches mentioned above to address missing values issues.
-
-### 3.2.1. Use Complete Cases (To be done)
-#### 3.2.1.1. Model Fitting
-#### 3.2.1.2. Assessment
-#### 3.2.1.3. Interpretation
-
-### 3.2.2. Selectively remove variables with high missing rate (To be done)
-#### 3.2.2.1. Model Fitting
-#### 3.2.2.2. Assessment
-#### 3.2.2.3. Interpretation
-
-### 3.2.3. Update the data set as we select variables (To be done)
-#### 3.2.3.1. Model Fitting
-#### 3.2.3.2. Assessment
-#### 3.2.3.3. Interpretation
-
-### 3.2.4. Imputation
+
+We can now conduct linear regressions following the approaches mentioned above to address missing values issues.
+
+## 3.1. Use Complete Cases (To be done)
+###  3.1.1. Model Fitting
+###  3.1.2. Assessment
+###  3.1.3. Interpretation
+
+## 3.2. Selectively remove variables with high missing rate (To be done)
+###  3.2.1. Model Fitting
+###  3.2.2. Assessment
+###  3.2.3. Interpretation
+
+## 3.3. Update the data set as we select variables (To be done)
+###  3.3.1. Model Fitting
+###  3.3.2. Assessment
+###  3.3.3. Interpretation
+
+## 3.4. Imputation
 **1. What is imputation**  
@@
 
-#### 3.2.4.1. Model Fitting
+###  3.4.1. Model Fitting
 ##### A. Ordinary Linear Regression
@@
   pred <- predict(f, countries.test.4)
-  resid <- pred - countries.test.4$pov
+  test.resid <- pred - countries.test.4$pov
   data.frame(
     `Train MSE` = mean(summary(f)$residuals^2),
-    `Test MSE` = mean(resid^2),
+    `Test MSE` = mean(test.resid^2),
     `Train MAE` = mean(abs(summary(f)$residuals)),
@@
 fit2.pred <- predict(fit2.dummy, countries.test.4)
-resid <- fit2.pred - countries.test.4$pov
+test.resid <- fit2.pred - countries.test.4$pov
 fit2.summ <- data.frame(
   `Train MSE` = mean(summary(fit2.dummy)$residuals^2),
-  `Test MSE` = mean(resid^2),
+  `Test MSE` = mean(test.resid^2),
   `Train MAE` = mean(abs(summary(fit2.dummy)$residuals)),
@@
   fit <- lm(as.formula(formula.str), complete(countries1.imputed, i))
-  aic.fit <- stepAIC(fit, trace = T, direction = 'backward')
+  aic.fit <- stepAIC(fit, trace = F, direction = 'backward')
   return(aic.fit)
@@
 fit4.pred <- predict(fit4.dummy, countries.test.4)
-resid <- fit4.pred - countries.test.4$pov
+test.resid <- fit4.pred - countries.test.4$pov
 fit4.summ <- data.frame(
   `Train MSE` = mean(summary(fit4.dummy)$residuals^2),
-  `Test MSE` = mean(resid^2),
+  `Test MSE` = mean(test.resid^2),
   `Train MAE` = mean(abs(summary(fit4.dummy)$residuals)),
@@
 
-#### 3.2.4.2. Assessment
-Check assumption
-outliers
-#### 3.2.4.3. Interpretation
-
-## 3.3. Panel Data Analysis (To be done)
+###  3.4.2. Assessment
+
+Now we are evaluating some performance metrics, checking assumptions, and remedy some potential problems.
+
+##### A. Homoscedasticity
+
+Plot the residuals ~ pov
+
+```{r}
+for (mod in models) {
+  resid <- mod$train$residuals
+  fitted <- mod$train$data$pov + mod$train$residuals
+  # plot(resid ~ pov, main = paste("Model", i))
+  print(ggplot(mapping = aes(y = resid, x = fitted)) +
+    geom_point() +
+    geom_smooth(formula = y ~ x, method = loess))
+}
+```
+
+There seems to be some larger variance with higher values of `pov`. We can conduct a statistical test to confirm the present of heteroscedasticity.
+
+The Goldfeld-Quandt test is performed by eliminating a certain number of observations from the dataset’s center, then comparing the spread of residuals between the two datasets on either side of the central observations.
+
+The Goldfeld-Quandt test examines two submodels’ variances divided by a defined breakpoint and rejects if the variances disagree.
+
+Under H0, the Goldfeld-Quandt test’s test statistic follows an F distribution with degrees of freedom as specified in the parameter.
+
+- **Null (H0):** Heteroscedasticity is not present.
+- **Alternative (H1):** Heteroscedasticity is present.
+
+```{r}
+lapply(models, \(m) {
+  resid <- m$train$residuals
+  # apply Goldfeld-Quandt test
+  gqtest(formula = resid ~ 1, order.by = m$train$data$pov)
+})
+```
+
+There is heteroscedasticity in our models.
+[Generalized Least Squares With Unknown For of Variance](https://rpubs.com/cyobero/187387) is a possible remedy of this problem, but at the cost of model interpretation.
+The high variance in higher end of `pov` can be explained by the relative volatile economics and politic climate in highly poor countries ($pov > 0.25$), leading to unstable effect of predictors.
+
+##### B. Independence
+
+Some possible source of dependency are poverty measured on the same country, or in the same year. These are accounted in our models by controlling those variables.
+
+Some countries have been engaging in international wars, being under the influences of foreign forces such as Iraq, and Afghanistan.
+They possess similar (and different) political climates.
+Some other groups of economical alliance, or enjoying similar natural resources: OPEC, EU, ASEAN, etc.
+might have similar characteristics that make them dependent.
+
+A solution to resolve this is to conduct the study on countries with similar geographical characteristics.
+In our study, however, we are interested in the general effects of certain variable across the globe.
+Although there are some dependence, the large number of entities with various features, we hypothesize, will cancel each other out, resulting in a net effect that gives an overview of the influences of variables.
+
+##### C. Normally Distributed Residuals
+
+```{r}
+for (m in models) {
+  p <- ggplot(mapping = aes(x = m$train$residuals)) +
+    geom_histogram(aes(y = ..density..), bins = 30) +
+    geom_density(alpha = .2, fill = "blue")
+
+  print(p)
+}
+```
+
+Although quite negatively skewed, the residuals are visibly normal.
+
+##### D. Outliers
+
+As elastic net model produced in
+
+###  3.4.3. Interpretation
 

### 2022-11-16 data/forward.Rmd

@@
 
-Through forward selection, we will do feature selection, through iteration of different models, a linear model consisting of income as a predictor has the highest adjusted r-square value=0.6105.
+Through forward selection, we will do feature selection. We will stop the forward selection when the adjusted r-value>0.75.
+
+Through iteration of different models, a linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201.
 
@@
 ```{r cars}
-summary(lm(formula = pov ~ income,data = train)) #0.6105
-summary(lm(formula = pov ~ reg,data = train)) #0.5287
-summary(lm(formula = pov ~ edu.total,data = train)) #0.09506
-summary(lm(formula = pov ~ hlth,data = train)) #0.09806
+summary(lm(formula = pov ~ income,data = train)) #0.6201
+summary(lm(formula = pov ~ reg,data = train)) #0.5569
+summary(lm(formula = pov ~ edu.total,data = train)) #0.09577
+summary(lm(formula = pov ~ hlth,data = train)) #0.0921
 ```
 Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
-reg as a predictor has the highest adjusted r-value=0.7031
+reg as a predictor has the highest adjusted r-value=0.7181
 
 ```{r}
-summary(lm(formula = pov ~ reg+income,data = train))
-summary(lm(formula = pov ~ edu.total+income,data = train))
-summary(lm(formula = pov ~ hlth+income,data = train))
-summary(lm(formula = pov ~ mil+income,data = train))
-summary(lm(formula = pov ~ fdi+income,data = train))
+summary(lm(formula = pov ~ reg+income,data = train)) #0.7181
+summary(lm(formula = pov ~ edu.total+income,data = train)) #0.6033
+summary(lm(formula = pov ~ hlth+income,data = train)) #0.6201
+summary(lm(formula = pov ~ mil+income,data = train)) #0.6282
 ```
 
-The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7364
+The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7492
 ```{r}
-summary(lm(formula = pov ~ edu.total+income +reg,data = train))
-summary(lm(formula = pov ~ hlth+income+reg,data = train))
-summary(lm(formula = pov ~ lgdp.dflt+income+reg,data = train))
-summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train))
+summary(lm(formula = pov ~ edu.total+income+reg,data = train)) #0.7347
+summary(lm(formula = pov ~ hlth+income+reg,data = train)) #0.6201
+summary(lm(formula = pov ~ mil+income+reg,data = train)) #0.6282
+summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train)) #0.7492
 ```
-gdp.pc+income+reg+lgdp.pc r=0.7599
+The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
 ```{r}
-summary(lm(formula = pov ~ mil+income+reg+lgdp.pc ,data = train))#0.7424
-summary(lm(formula = pov ~ pop.gwth.total+income+reg+lgdp.pc,data = train)) #0.7433
-summary(lm(formula = pov ~ gdp.dflt+income+reg+lgdp.pc,data = train)) #0.7466
-summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7599
+summary(lm(formula = pov ~ edu.total+income+reg+lgdp.pc,data = train)) #0.7495
+summary(lm(formula = pov ~ mil+income+reg+lgdp.pc,data = train)) #0.7571
+summary(lm(formula = pov ~ pop.gwth.urban+income+reg+lgdp.pc,data = train)) #0.7606
+summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7703
 ```
-Highest adjusted r-square=0.7466 which is lower than our previous model
+
+Fitting of multiple linear regression with predictor gdp.pc, income, reg and lgd.pc.
 ```{r}
-summary(lm(formula = pov ~ trade+income+reg+lgdp.pc,data = train))
-summary(lm(formula = pov ~ pop.gwth.total+income+reg+lgdp.pc,data = train)) 
-summary(lm(formula = pov ~ pop.gwth.rural+income+reg+lgdp.pc,data = train))
-summary(lm(formula = pov ~ pop.gwth.urban+income+reg+lgdp.pc,data = train)) #0.7466
-```
-```{r}
-lm1<-(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train))
+lm1<-lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)
 summary(lm1)
 ```
-Assessment
+
+Assessment of linear model lm1: 
+the predicted model has some NA values which are removed. The error rates are too high, therefore, this method of variable selection may not be suitable
 ```{r}
@@
 actual<-test$pov
-eval.metrics.linreg(actual,predicted)
-actual
-predicted
+df23<-na.omit(data.frame(actual,predicted))
+ac<-df23$actual
+pr<-df23$predicted
+eval.metrics.linreg(ac,pr)
 ```
+
+Intepretation: gdp per capital, income class and region of the country can provide a model to predict the poverty headcount ratio in a country.

### 2022-11-16 src/All variables.Rmd
Update All variables.Rmd

@@
 knitr::kable(summary, digits = 3)
+
 ```
+
+```{r}
+knitr::kable(summary_MLR, digits = 3)
+knitr::kable(summary_ridge[,-5], digits = 3)
+```

### 2022-11-16 data/forward.Rmd
forward5

@@
 df_count<-read.csv("../data/countries1.csv")
-head(df_count)
-
 set.seed(10)
@@
 
-Through iteration of different models, a linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201.
+Through iteration of different models. A linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201, thus it is selected.
 
-Below are some of the linear models that have been explored.
+Only some of the linear models that have been explored are showed.
 ```{r cars}
-summary(lm(formula = pov ~ income,data = train)) #0.6201
+sum1<-summary(lm(formula = pov ~ income,data = train)) #0.6201
 summary(lm(formula = pov ~ reg,data = train)) #0.5569

### 2022-11-16 src/draft.Rmd
forward5

--- 2022-11-15 src/draft.Rmd
@@
 #### 3.2.2.1. Model Fitting
+Through forward selection, we will do feature selection. We will stop the forward selection when the adjusted r-value>0.75.
+
+Through iteration of different models. A linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201, thus it is selected.
+
+Only some of the linear models that have been explored are showed.
+```{r}
+df_count<-read.csv("../data/countries1.csv")
+set.seed(10)
+index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
+train <- df_count[index, ]
+test <- df_count[-index, ]
+```
+
+Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
+reg as a predictor has the highest adjusted r-value=0.7181
+
+```{r}
+summary(lm(formula = pov ~ reg+income,data = train)) #0.7181
+```
+
+The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7492
+```{r}
+summary(lm(formula = pov ~ edu.total+income+reg,data = train)) #0.7347
+```
+
+The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
+```{r}
+summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7703
+```
+
+Fitting of multiple linear regression with predictor gdp.pc, income, reg and lgd.pc.
+```{r}
+lm1<-lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)
+summary(lm1)
+```
+
 #### 3.2.2.2. Assessment
+Assessment of linear model lm1: 
+the predicted model has some NA values which are removed. The error rates are too high, therefore, this method of variable selection may not be suitable
+```{r}
+eval.metrics.linreg <- function(actual, predicted) {
+  residual <- actual - predicted
+  mse <- mean(residual ^ 2)
+  mae <- mean(abs(residual))
+  rmse <-  sqrt(mse)
+  mape <- mean(abs(residual / actual)) * 100
+
+  data.frame(
+    MSE = mse,
+    MAE = mae,
+    RMSE = rmse,
+    MAPE = mape
+  )
+}
+predicted<-predict(lm1,newdata=test)
+actual<-test$pov
+df23<-na.omit(data.frame(actual,predicted))
+ac<-df23$actual
+pr<-df23$predicted
+eval.metrics.linreg(ac,pr)
+```
+
 #### 3.2.2.3. Interpretation
+gdp per capital, income class and region of the country can provide a model to predict the poverty headcount ratio in a country.
 

### 2022-11-16 src/All variables.Rmd
Update All variables.Rmd

@@
 ```{r}
+corrplot(cor(subset2, use = "complete.obs"), 
+         method = "number",
+         type = "upper",
+         tl.col = "black", 
+         tl.srt = 30
+         )
+```
+
+```{r}
 set.seed(123)
@@
 ```
+```{r}
+model3 <- lm(pov ~., data = training)
+summary(model3)
+
+```
 
 ```{r}
-corrplot(cor(subset2, use = "complete.obs"), 
-         method = "number",
-         type = "upper",
-         tl.col = "black", 
-         tl.srt = 30
-         )
+fit <- model3$fitted.values
+true <- training$pov
+summary_MLR3_train <- eval_results(true, fit)
+```
+
+```{r}
+fit <- predict(model3, newdata=test)
+true <- test$pov
+summary_MLR3_test <- eval_results(true, fit)
+```
+
+```{r}
+summary_MLR <- rbind(summary_MLR3_train[-5], 
+                     summary_MLR3_test[-5])
+rownames(summary_MLR) <- c("Baseline MLR_Train", 
+                           "Baseline MLR_Test" )
+knitr::kable(summary_MLR, digits = 3)
+```
+
+```{r}
+# Training Data
+data <- training
+train.x <- data.matrix(data[,-ncol(data)])
+train.y <- data.matrix(data[,ncol(data)])
+
+# Test Data
+data <- test
+test.x <- data.matrix(data[,-ncol(data)])
+test.y <- data.matrix(data[,ncol(data)])
+```
+
+```{r}
+# Fit Ridge Regression Model
+set.seed(123)
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x, train.y, 
+                    alpha = 0, lambda = cv_ridge$lambda.min)
+
+# Evaluate Ridge Regression Model on Training Set
+fit <- predict(glm_ridge, newx = train.x, type = "response")
+true <- train.y
+summary_ridge_train <- eval_results(true, fit)
+
+# Evaluate Ridge Regression Model on Test Set
+fit <- predict(glm_ridge, newx = test.x, type = "response")
+true <- test.y
+summary_ridge_test <- eval_results(true, fit)
+
+# Summarise Ridge Regression Model Performance into Table
+summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
+rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
+```
+
+```{r}
+# Fit LASSO Model
+set.seed(123)
+cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
+glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
+                    lambda = cv_lasso$lambda.min)
+
+# Evaluate LASSO Model on Training Set
+fit <- predict(glm_lasso, newx = train.x)
+true <- train.y
+summary_lasso_train <- eval_results(true, fit)
+
+# Evaluate LASSO Model on Test Set
+fit <- predict(glm_lasso, newx = test.x)
+true <- test.y
+summary_lasso_test <- eval_results(true, fit)
+
+# Summarise LASSO Model Performance into Table
+summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
+rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
+```
+
+```{r}
+summary_3models <- rbind(summary_MLR3_train, 
+                         summary_ridge_train,                               
+                         summary_lasso_train, 
+                         summary_MLR3_test,                             
+                         summary_ridge_test, 
+                         summary_lasso_test)
+
+rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
+                               "LASSO Model_Train",  "Baseline MLR_Test",  
+                               "Ridge Model_Test", "LASSO Model_Test" )
+
+knitr::kable(summary_3models[1:3,-5], digits = 3)
+knitr::kable(summary_3models[4:6,-5], digits = 3)
+
 ```
@@
 ```
+
+```{r}
+# Fit Ridge Regression Model
+set.seed(123)
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x, train.y, 
+                    alpha = 0, lambda = cv_ridge$lambda.min)
+
+# Evaluate Ridge Regression Model on Training Set
+fit <- predict(glm_ridge, newx = train.x, type = "response")
+true <- train.y
+summary_ridge_train <- eval_results(true, fit)
+
+# Evaluate Ridge Regression Model on Test Set
+fit <- predict(glm_ridge, newx = test.x, type = "response")
+true <- test.y
+summary_ridge_test <- eval_results(true, fit)
+
+# Summarise Ridge Regression Model Performance into Table
+summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
+rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
+```
+
+```{r}
+# Fit LASSO Model
+set.seed(123)
+cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
+glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
+                    lambda = cv_lasso$lambda.min)
+
+# Evaluate LASSO Model on Training Set
+fit <- predict(glm_lasso, newx = train.x)
+true <- train.y
+summary_lasso_train <- eval_results(true, fit)
+
+# Evaluate LASSO Model on Test Set
+fit <- predict(glm_lasso, newx = test.x)
+true <- test.y
+summary_lasso_test <- eval_results(true, fit)
+
+# Summarise LASSO Model Performance into Table
+summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
+rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
+```
+
+```{r}
+summary_3models <- rbind(summary_MLR3_train, 
+                         summary_ridge_train,                               
+                         summary_lasso_train, 
+                         summary_MLR3_test,                             
+                         summary_ridge_test, 
+                         summary_lasso_test)
+
+rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
+                               "LASSO Model_Train",  "Baseline MLR_Test",  
+                               "Ridge Model_Test", "LASSO Model_Test" )
+
+knitr::kable(summary_3models[1:3,-5], digits = 3)
+knitr::kable(summary_3models[4:6,-5], digits = 3)
+```

### 2022-11-16 src/draft.Rmd

--- 2022-11-16 src/draft.Rmd
@@
     geom_point() +
-    geom_smooth(method = lm) +
+    geom_smooth(formula = y ~ x, method = lm) +
     labs(title = paste0("pov ~ log(", indvar, ")"),
@@
 
-## 3.1. Use Complete Cases (To be done)
+## 3.1. Use Complete Cases (Imtiyaaz
 ###  3.1.1. Model Fitting
@@
 
-## 3.2. Selectively remove variables with high missing rate (To be done)
+## 3.2. Selectively remove variables with high missing rate (Callista)
 ###  3.2.1. Model Fitting
@@
 
-## 3.3. Update the data set as we select variables (To be done)
+## 3.3. Update the data set as we select variables (Isaac)
 ###  3.3.1. Model Fitting
@@
 ## 3.4. Imputation
-**1. What is imputation**  
+**What is imputation?**  
 Imputation is a technique to handle missing values by replacing missing data with substitute values.
@@
 
-- `m = 5` imputed data sets
-- `maxit = 30` iterations
+- `m = 3` imputed data sets
+- `maxit = 20` iterations
 
@@
 # split data
-isComplete <- which(complete.cases(countries1))
-idx <- sample(isComplete, replace = F, 0.3 * nrow(countries1))
-countries.train.4 <- countries1[-idx,]
-countries.test.4 <- countries1[idx,]
-
-# blank for no imputation, cart for variable requiring imputation
-meth <- c(rep("", 4), "cart", "", rep("cart", 16))
-names(meth) <- colnames(countries1)
-meth
-
-# run the algorithm
-# countries1.imputed <- mice(countries1, m = 3, maxit = 20, method = meth)
+seeds <- countries1 %>% 
+  group_by(country.code) %>% 
+  filter(row_number() == 1)
+
+plants <- countries1 %>% 
+  group_by(country.code) %>% 
+  filter(row_number() != 1)
+
+isComplete <- which(complete.cases(plants))
+idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
+
+countries.train.4 <- plants[-idx,] %>% 
+  rbind(seeds)
+countries.test.4 <- plants[idx,]
+
+# run the algorithm, or read from the file we generated as this is a very time-consuming process
+# countries1.imputed <- mice(countries.train.4, m = 3, maxit = 20, method = meth)
 # saveRDS(countries1.imputed, 'countries1.imputed.RData')
@@
 ```
-We can take a look into one of the imputated data sets.
+We can take a look into one of the imputed data sets.
 ```{r}
@@
 ###  3.4.1. Model Fitting
-##### A. Ordinary Linear Regression
-We generated 3 sets of imputated (training) data.
+#### A. Ordinary Linear Regression
+We generated 3 sets of imputed (training) data.
 We can build separate models using each data set, and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
@@
 ```{r}
-# fit model on imputated set no. 1
+# fit model on imputed set no. 1
 summary(fit1[[1]])
@@
 
-##### B. VIF
+#### B. VIF
 
@@
 
-##### B. Step-wise AIC
+#### B. Step-wise AIC
 We can conduct a step-wise AIC variable selection.
@@
 
-Performance on train set and test set are not affected. However, we have simplified the model by removing some unnecessary terms. We can update our models.
+Performance on train set and test set are slightly improved. However, we have simplified the model by removing some unnecessary terms. We can update our models.
 
@@
 
-##### C. Regularization
+#### C. Regularization
 
@@
       adj.r2 = train.adjR2,
+      fitted = train.fit,
       residuals = train.resid,
@@
       adj.r2 = test.adjR2,
+      fitted = test.fit,
       residuals = test.resid,
@@
 
-Build the optimal models using each imputated data set.
-
-```{r}
-# helper function to build from a specific imputated data
+Build the optimal models using each imputed data set.
+
+```{r}
+# helper function to build from a specific imputed data
 buildWith <- function(set, imputed, test, formula, fun) {
@@
 
-##### A. Homoscedasticity
+#### A. Homoscedasticity
 
@@
   resid <- mod$train$residuals
-  fitted <- mod$train$data$pov + mod$train$residuals
+  fitted <- mod$train$fitted
   # plot(resid ~ pov, main = paste("Model", i))
@@
   # apply Goldfeld-Quandt test
-  gqtest(formula = resid ~ 1, order.by = m$train$data$pov)
+  gqtest(formula = resid ~ 1, order.by = m$train$fitted)
 })
@@
 [Generalized Least Squares With Unknown For of Variance](https://rpubs.com/cyobero/187387) is a possible remedy of this problem, but at the cost of model interpretation.
-The high variance in higher end of `pov` can be explained by the relative volatile economics and politic climate in highly poor countries ($pov > 0.25$), leading to unstable effect of predictors.
-
-##### B. Independence
+The high variance in higher end of `pov` can be explained by the relative volatile economical and political climate in highly poor countries (one standard deviation from mean), leading to unstable effect of predictors.
+
+#### B. Independence
 
@@
 
-##### C. Normally Distributed Residuals
+#### C. Normally Distributed Residuals
 
@@
 
-##### D. Outliers
-
-As elastic net model produced in
+#### D. Outliers
+
+As elastic net models produced in the last section are not very fit for detecting outliers using R-support function.
+As they just removed dummy variables in `country.code`, we can using OLS to detect some potential outliers.
+They are the same models we produced after the step-AIC selection (`fit3` and `fit4`)
+
+```{r}
+for (mod in fit3) {
+  plot(mod, which = 4)
+}
+```
+
+Some obvious outliers are point 1122, 1327, 1395, 1396, 1424, 1474.
+
+```{r}
+out <- c(1122, 1327, 1395, 1396, 1424, 1474)
+var <- c("country.code","country.name","year","pov","income","edu.total","hlth","mil","lbr.part","gcf","lfdi")
+# See the original data instead of imputed data
+countries.train.4[out,var]
+```
+
+These countries all appear very infrequently in our data set (only twice for South Sudan and St.Lucia).
+The gap between their appearance are quite large, too.
+Such reasons could have made the model struggle to predict their poverty states.
+
+Guinea is an interesting point. Because the outlier data (87.9% in 1991) is quite different to the next data point (40.8% in 1994), a huge jump in just 3 years.
 

### 2022-11-16 src/draft.Rmd
f2
some restructuring

@@
 
-Guinea is an interesting point. Because the outlier data (87.9% in 1991) is quite different to the next data point (40.8% in 1994), a huge jump in just 3 years.
+Yemen's data were collected in 1998, 2005, and 2014, a huge gap in time, by when it has doubled its poverty rate.
+
+Guinea is an interesting point.
+Because the outlier data (87.9% in 1991) is quite different to the next data point (40.8% in 1994), a huge jump in just 3 years.
+This might be a consequence of the political changes (new constitution, establishment of the Supreme Court, etc.) in this period.
+They are the changes not captured by the variables.
+
+Gambia has the issue with Guinea, when it saw a huge reduction of poverty of 25% from 1998-2003.
+
+Lucia only has two data points with large gap in time (1995-2016). It can be difficult to know much about someone if you only met them twice, with 21 years in between.
 

### 2022-11-16 src/draft.Rmd
f3

--- 2022-11-16 src/draft.Rmd
@@
 ###  3.3.1. Model Fitting 
+```{r}
+df_count<-read.csv("../data/countries1.csv")
+set.seed(10)
+index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
+train <- df_count[index, ]
+test <- df_count[-index, ]
+```
+
+Through forward selection, we will do feature selection. We will stop the forward selection when the adjusted r-value>0.75.
+
+Through iteration of different models. A linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201, thus it is selected.
+
+Only some of the linear models that have been explored are showed.
+```{r cars}
+summary(lm(formula = pov ~ income,data = train)) #0.6201
+```
+Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
+reg as a predictor has the highest adjusted r-value=0.7181
+
+```{r}
+summary(lm(formula = pov ~ reg+income,data = train)) #0.7181
+```
+
+The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7492
+```{r}
+summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train)) #0.7492
+```
+The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
+```{r}
+summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7703
+```
+
+Fitting of multiple linear regression with predictor gdp.pc, income, reg and lgd.pc.
+```{r}
+lm1<-lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)
+summary(lm1)
+```
+
 ###  3.3.2. Assessment

### 2022-11-16 src/draft.Rmd
f4

@@
 ###  3.3.2. Assessment
+Assessment of linear model lm1: 
+the predicted model has some NA values which are removed. The error rates are too high, therefore, this method of variable selection may not be suitable
+```{r}
+eval.metrics.linreg <- function(actual, predicted) {
+  residual <- actual - predicted
+  mse <- mean(residual ^ 2)
+  mae <- mean(abs(residual))
+  rmse <-  sqrt(mse)
+  mape <- mean(abs(residual / actual)) * 100
+
+  data.frame(
+    MSE = mse,
+    MAE = mae,
+    RMSE = rmse,
+    MAPE = mape
+  )
+}
+predicted<-predict(lm1,newdata=test)
+actual<-test$pov
+df23<-na.omit(data.frame(actual,predicted))
+ac<-df23$actual
+pr<-df23$predicted
+eval.metrics.linreg(ac,pr)
+```
 ###  3.3.3. Interpretation
+gdp per capital, income class and region of the country can provide a model to predict the poverty headcount ratio in a country.
 

### 2022-11-16 src/draft.Rmd

--- 2022-11-16 src/draft.Rmd
@@
 
-Only some of the linear models that have been explored are showed.
 ```{r cars}

### 2022-11-16 src/draft.Rmd
f6

--- 2022-11-16 src/draft.Rmd
@@
 ```
+
 Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
@@
 ```
+
 The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
@@
 ```
+
 ###  3.3.3. Interpretation

### 2022-11-16 src/Selectively_remove_variables.Rmd
updated my part (selectively removing variables)

@@
 ```
-
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
 
@@
 ```
-We first read the 
+
 ```{r}
@@
 ```
-
+This approach takes variables that contain less than 35% of missing data.The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models.
+Hence, we transformed the data with the mutate() function before building the machine learning model.
 ```{r}
 
-#dataset without the red 
-#removing reg as it is a bit too complex to deal for now, not relevant
-
-#subset1 <- subset(countries, select = c("edu.total", "fdi", "gcf", "gdp.dflt", "hlth", "income", "lbr.part", "mil", "pop.gwth.rural", "pop.gwth.urban", "pov", "reg", "trade", "unemp", "year"))
 subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
 
-subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.gdlt = log(subset0$gdp.dflt))
-#dataset
+subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
+
 ```
+subset1 contains the variables below 35% of missing values. 
+```{r}
+subset1 <- subset(subset1, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
 
+subset1 <- na.omit(subset1)
+```
+We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step. 
 ```{r}
-#converting non-numeric arguments into factors
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
 
-subset1$income <- factor(subset1$income)
-subset1$income <- as.numeric(subset1$income)
+  return(sd(vct, na.rm = T))
+}
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+
+  return(vct/scl)
+}
+
+scale <- function(data) {
+  unlist(lapply(data, sd0))
+}
+
+standardise <- function(data) {
+  scaler <- scale(data)
+
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}
 
@@
 ```{r}
-#Perform standardisation
 
-(df.countries0 <- subset1)
-scaler <- apply(X = df.countries0, MARGIN = 2, FUN = sd)
-
-(df.countries <- as.data.frame(apply(df.countries0, 2, function(x) x/sd(x, na.rm = TRUE))))
-
+df.countries0 <- standardise(subset1)
+df.countries <- df.countries0
 ```
@@
 
-```{r}
-#check for correlation
-
-corrplot(cor(df.countries, use = "complete.obs"), 
-         method = "number",
-         type = "upper",
-         tl.col = "black", 
-         tl.srt = 30
-         )
-```
-
+First, we split the data into training and testing data. 
 ```{r}
@@
 
-```{r}
-#Boxplot of the predictor variables
-```
-
-```{r}
-```
-
+Our first attempt 
 ```{r}
@@
 #remove hlth 
-lm_2 <- lm(data= train_1, formula = pov ~ edu.total + gcf + fdi + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
+lm_2 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
 summary(lm_2)
@@
 ```{r}
-#Remove fdi
-lm_3 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
+#Remove lgdp.dflt
+lm_3 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
 summary(lm_3)
@@
 ```{r}
-#remove year 
-lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
+#remove lfdi
+lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
 summary(lm_4)
@@
 ```
+Our model shows that among all the 13 predictor variables that we started with, only 10 of them showed significance in building the model.
+
+We then evaluated the results of the model. 
 ```{r}
-#remove gdp.dflt 
-lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
-summary(lm_4)
-plot(lm_4, which = 5)
+#evaluating lm_4
+eval.metrics.linreg <- function(actual, predicted) {
+  residual <- actual - predicted
+  mse <- mean(residual ^ 2)
+  mae <- mean(abs(residual))
+  rmse <-  sqrt(mse)
+  mape <- mean(abs(residual / actual)) * 100
+
+  data.frame(
+    MSE = mse,
+    MAE = mae,
+    RMSE = rmse,
+    MAPE = mape
+  )
+}
 ```
+
 ```{r}
-lm_5 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc)
-summary(lm_5)
-plot(lm_5, which = 5)
+actual <- train_1$pov
+predicted <- predict(lm_4, newdata = train_1)
+eval.metrics.linreg(actual, predicted)
 ```
+
 ```{r}
-#removing lfdi
-lm_6 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc  + lgdp.pc)
-summary(lm_6)
-plot(lm_6, which = 4)
-```
-```{r}
-#
-data0 <- subset(df.countries, select = c("edu.total","gcf","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp","gdp.pc", 
-                                        "lgdp.pc"))
+#subsetting df.countries to only contain variables present in lm_7
+
+data0 <- subset1
 data <- na.omit(data0)
 train.x <- data.matrix(subset(data, select = -pov))
-train.y = data.matrix(data$pov)
+train.y <- data.matrix(data$pov)
 
-corrplot(cor(data, use = "complete.obs"), 
-         method = "number",
-         type = "upper",
-         tl.col = "black", 
-         tl.srt = 30
-         )
 ```
@@
 set.seed(123)
-rate <- 0.8
+rate <- 0.6
 train.size <- round(nrow(data)*rate)
@@
 ```{r}
-cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
-glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
-
-t(coef(glm_ridge))
-```
-
-```{r}
-train.x <- model.matrix(pov ~., data = training)[, -1]
+training <- na.omit(training)
+train.x <- model.matrix(pov ~., data = training)[,-1]
 train.y <- training$pov
@@
 ```{r}
-model_ridge_trial1 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.1)
-t(coef(model_ridge_trial1))
-```
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
 
-```{r}
-model_ridge_trial2 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.5)
-t(coef(model_ridge_trial2))
-```
-
-```{r}
-model_ridge_trial3 <- glmnet(train.x, train.y, alpha = 0, lambda = 1)
-t(coef(model_ridge_trial3))
-```
-```{r}
-table_summary <- cbind(coef(model_ridge_trial1),
-                       coef(model_ridge_trial2),
-                       coef(model_ridge_trial3))
-
-colnames(table_summary) <- c("lambda = 0.1",
-                             "lambda = 0.5",
-                             "lambda = 1")
-
-table_summary
+t(coef(glm_ridge))
 ```
@@
 ```{r}
-mlr_adj <- lm(pov ~., data = training)
-
-summary(mlr_adj)
-```
-
-```{r}
 #Train a LASSO regression model

### 2022-11-16 src/draft.Rmd
updated my part (selectively removing variables)

--- 2022-11-16 src/draft.Rmd
@@
     geom_point() +
-    geom_smooth(formula = y ~ x, method = lm) +
+    geom_smooth(method = lm) +
     labs(title = paste0("pov ~ log(", indvar, ")"),
@@
 
-## 3.1. Use Complete Cases (Imtiyaaz
+## 3.1. Use Complete Cases (To be done)
 ###  3.1.1. Model Fitting
@@
 
-## 3.2. Selectively remove variables with high missing rate (Callista)
+## 3.2. Selectively remove variables with high missing rate (To be done)
+**1. About selectively removing variables with high missing rate**
+
+Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables.
+This section attempts to build a model on variables that have missing rates of below 35%.
+
+The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models.
+Hence, we transformed the data with the mutate() function before building the machine learning model.
+```{r}
+subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
+
+subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
+
+subset1 <- subset(subset1, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
+
+subset1 <- na.omit(subset1)
+```
+We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step.
+
+```{r}
+#helper function
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
+
+  return(sd(vct, na.rm = T))
+}
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+
+  return(vct/scl)
+}
+
+scale <- function(data) {
+  unlist(lapply(data, sd0))
+}
+
+standardise <- function(data) {
+  scaler <- scale(data)
+
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}
+
+df.countries0 <- standardise(subset1)
+df.countries <- df.countries0
+head(df.countries)
+```
+
 ###  3.2.1. Model Fitting
+##### A. Ordinary Linear Regression
+First, we split the data into training and testing data. The training dataset contains 80% of the initial dataset, and the testing data contains 20% of the whole dataset. 
+```{r}
+set.seed(123)
+index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
+train_1 <- subset1[index,]
+test_1 <- subset1[-index,]
+
+```
+Our first model includes all the variables with missing rate lower than 35%. This model is assigned to `lm_1`.
+```{r}
+lm_1 <- lm(data= train_1, formula = pov ~ .)
+summary(lm_1)
+```
+The summary of `lm_1` suggests that the model still contains predictor variables with pr(>|t|) > 0.05.
+We can imply that these variables do not provide a high significance to our model and are thus able to be removed.
+To avoid removing significant variables, we shall conduct the removal of the variables deliberately.
+In `lm_1`, we notice that the variable `lgdp.dflt` has a high Pr(>|t|).
+Hence, we remove `lgdp.dflt`.
+```{r}
+lm_2 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
+summary(lm_2)
+```
+Removing `lfdi`,
+```{r}
+lm_3 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
+summary(lm_3)
+```
+Further removing `hlth`,
+```{r}
+lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
+summary(lm_4)
+plot(lm_4, which = 5)
+```
+Our model shows that among all the 13 predictor variables that we started with, only 10 of them showed significance in building the model.
+
+##### B. VIF
+
+##### c. Regularization
+Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model. 
+Later on, we will compare the results of all the models we generated.
+```{r}
+data0 <- subset1
+data <- na.omit(data0)
+train.x <- data.matrix(subset(data, select = -pov))
+train.y <- data.matrix(data$pov)
+
+```
+Dividing into training and testing datasets.
+```{r}
+set.seed(123)
+rate <- 0.6
+train.size <- round(nrow(data)*rate)
+sample(sample(nrow(data), train.size))
+training <- data[sample, ]
+test <- data[-sample, ]
+```
+Further dividing into `train.x`, `train.y`, `test.x`, and `test.y`.
+```{r}
+training <- na.omit(training)
+train.x <- model.matrix(pov ~., data = training)[,-1]
+train.y <- training$pov
+test.x <- model.matrix(pov ~., data = test)[,-1]
+test.y <- test$pov
+```
+Using the lambda value that yields the lowest MSE value. 
+```{r}
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
+
+t(coef(glm_ridge))
+```
+
+```{r}
+#train the ridge regression model using lambda.min
+
+glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
+t(coef(glm_Ridge))
+```
+
+Function to evaluate the results 
+```{r}
+eval_results <- function(fit, true) {
+  actual <- data.matrix(true)
+  SSE <- sum((actual - fit)^2)
+  SST <- sum((actual - mean(actual))^2)
+  R_square <- 1 - SSE/SST
+  data.frame(
+    MSE = MSE(fit, true),
+    MAE = MAE(fit, true),
+    RMSE = RMSE(fit, true),
+    MAPE = MAPE(fit, true),
+    R2  = R_square
+
+  )
+}
+```
+
+```{r}
+fit <- predict(glm_Ridge, train.x)
+true <- train.y
+summary_Ridge_train <- eval_results(fit, true)
+
+```
+
+```{r}
+fit <- predict(glm_Ridge, test.x)
+true <- test.y
+summary_Ridge_test <- eval_results(fit, true)
+
+```
+
+```{r}
+summary_Ridge <- rbind(summary_Ridge_train[-5], summary_Ridge_test[-5])
+rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
+knitr::kable(summary_Ridge, digits = 3)
+
+```
+
 ###  3.2.2. Assessment
@@
 
-## 3.3. Update the data set as we select variables (Isaac)
+## 3.3. Update the data set as we select variables (To be done)
 ###  3.3.1. Model Fitting
-```{r}
-df_count<-read.csv("../data/countries1.csv")
-set.seed(10)
-index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
-train <- df_count[index, ]
-test <- df_count[-index, ]
-```
-
-Through forward selection, we will do feature selection. We will stop the forward selection when the adjusted r-value>0.75.
-
-Through iteration of different models. A linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201, thus it is selected.
-
-```{r cars}
-summary(lm(formula = pov ~ income,data = train)) #0.6201
-```
-
-Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
-reg as a predictor has the highest adjusted r-value=0.7181
-
-```{r}
-summary(lm(formula = pov ~ reg+income,data = train)) #0.7181
-```
-
-The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7492
-```{r}
-summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train)) #0.7492
-```
-
-The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
-```{r}
-summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7703
-```
-
-Fitting of multiple linear regression with predictor gdp.pc, income, reg and lgd.pc.
-```{r}
-lm1<-lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)
-summary(lm1)
-```
-
 ###  3.3.2. Assessment
-Assessment of linear model lm1: 
-the predicted model has some NA values which are removed. The error rates are too high, therefore, this method of variable selection may not be suitable
-```{r}
-eval.metrics.linreg <- function(actual, predicted) {
-  residual <- actual - predicted
-  mse <- mean(residual ^ 2)
-  mae <- mean(abs(residual))
-  rmse <-  sqrt(mse)
-  mape <- mean(abs(residual / actual)) * 100
-
-  data.frame(
-    MSE = mse,
-    MAE = mae,
-    RMSE = rmse,
-    MAPE = mape
-  )
-}
-predicted<-predict(lm1,newdata=test)
-actual<-test$pov
-df23<-na.omit(data.frame(actual,predicted))
-ac<-df23$actual
-pr<-df23$predicted
-eval.metrics.linreg(ac,pr)
-```
-
 ###  3.3.3. Interpretation
-gdp per capital, income class and region of the country can provide a model to predict the poverty headcount ratio in a country.
 
 ## 3.4. Imputation
-**What is imputation?**  
+**1. What is imputation**  
 Imputation is a technique to handle missing values by replacing missing data with substitute values.
@@
 
-- `m = 3` imputed data sets
-- `maxit = 20` iterations
+- `m = 5` imputed data sets
+- `maxit = 30` iterations
 
@@
 # split data
-seeds <- countries1 %>% 
-  group_by(country.code) %>% 
-  filter(row_number() == 1)
-
-plants <- countries1 %>% 
-  group_by(country.code) %>% 
-  filter(row_number() != 1)
-
-isComplete <- which(complete.cases(plants))
-idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
-
-countries.train.4 <- plants[-idx,] %>% 
-  rbind(seeds)
-countries.test.4 <- plants[idx,]
-
-# run the algorithm, or read from the file we generated as this is a very time-consuming process
-# countries1.imputed <- mice(countries.train.4, m = 3, maxit = 20, method = meth)
+isComplete <- which(complete.cases(countries1))
+idx <- sample(isComplete, replace = F, 0.3 * nrow(countries1))
+countries.train.4 <- countries1[-idx,]
+countries.test.4 <- countries1[idx,]
+
+# blank for no imputation, cart for variable requiring imputation
+meth <- c(rep("", 4), "cart", "", rep("cart", 16))
+names(meth) <- colnames(countries1)
+meth
+
+# run the algorithm
+# countries1.imputed <- mice(countries1, m = 3, maxit = 20, method = meth)
 # saveRDS(countries1.imputed, 'countries1.imputed.RData')
@@
 ```
-We can take a look into one of the imputed data sets.
+We can take a look into one of the imputated data sets.
 ```{r}
@@
 ###  3.4.1. Model Fitting
-#### A. Ordinary Linear Regression
-We generated 3 sets of imputed (training) data.
+##### A. Ordinary Linear Regression
+We generated 3 sets of imputated (training) data.
 We can build separate models using each data set, and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
@@
 ```{r}
-# fit model on imputed set no. 1
+# fit model on imputated set no. 1
 summary(fit1[[1]])
@@
 
-#### B. VIF
+##### B. VIF
 
@@
 
-#### B. Step-wise AIC
+##### B. Step-wise AIC
 We can conduct a step-wise AIC variable selection.
@@
 
-Performance on train set and test set are slightly improved. However, we have simplified the model by removing some unnecessary terms. We can update our models.
+Performance on train set and test set are not affected. However, we have simplified the model by removing some unnecessary terms. We can update our models.
 
@@
 
-#### C. Regularization
+##### C. Regularization
 
@@
       adj.r2 = train.adjR2,
-      fitted = train.fit,
       residuals = train.resid,
@@
       adj.r2 = test.adjR2,
-      fitted = test.fit,
       residuals = test.resid,
@@
 
-Build the optimal models using each imputed data set.
-
-```{r}
-# helper function to build from a specific imputed data
+Build the optimal models using each imputated data set.
+
+```{r}
+# helper function to build from a specific imputated data
 buildWith <- function(set, imputed, test, formula, fun) {
@@
 
-#### A. Homoscedasticity
+##### A. Homoscedasticity
 
@@
   resid <- mod$train$residuals
-  fitted <- mod$train$fitted
+  fitted <- mod$train$data$pov + mod$train$residuals
   # plot(resid ~ pov, main = paste("Model", i))
@@
   # apply Goldfeld-Quandt test
-  gqtest(formula = resid ~ 1, order.by = m$train$fitted)
+  gqtest(formula = resid ~ 1, order.by = m$train$data$pov)
 })
@@
 [Generalized Least Squares With Unknown For of Variance](https://rpubs.com/cyobero/187387) is a possible remedy of this problem, but at the cost of model interpretation.
-The high variance in higher end of `pov` can be explained by the relative volatile economical and political climate in highly poor countries (one standard deviation from mean), leading to unstable effect of predictors.
-
-#### B. Independence
+The high variance in higher end of `pov` can be explained by the relative volatile economics and politic climate in highly poor countries ($pov > 0.25$), leading to unstable effect of predictors.
+
+##### B. Independence
 
@@
 
-#### C. Normally Distributed Residuals
+##### C. Normally Distributed Residuals
 
@@
 
-#### D. Outliers
-
-As elastic net models produced in the last section are not very fit for detecting outliers using R-support function.
-As they just removed dummy variables in `country.code`, we can using OLS to detect some potential outliers.
-They are the same models we produced after the step-AIC selection (`fit3` and `fit4`)
-
-```{r}
-for (mod in fit3) {
-  plot(mod, which = 4)
-}
-```
-
-Some obvious outliers are point 1122, 1327, 1395, 1396, 1424, 1474.
-
-```{r}
-out <- c(1122, 1327, 1395, 1396, 1424, 1474)
-var <- c("country.code","country.name","year","pov","income","edu.total","hlth","mil","lbr.part","gcf","lfdi")
-# See the original data instead of imputed data
-countries.train.4[out,var]
-```
-
-These countries all appear very infrequently in our data set (only twice for South Sudan and St.Lucia).
-The gap between their appearance are quite large, too.
-Such reasons could have made the model struggle to predict their poverty states.
-
-Guinea is an interesting point. Because the outlier data (87.9% in 1991) is quite different to the next data point (40.8% in 1994), a huge jump in just 3 years.
+##### D. Outliers
+
+As elastic net model produced in
 

### 2022-11-16 src/Selectively_remove_variables.Rmd
Update Selectively_remove_variables.Rmd

@@
 
-subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
+subset0 <- subset(countries, select = c("country.code", "year","edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
 
-subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
+subset1_0 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
 
@@
 ```{r}
-subset1 <- subset(subset1, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
+subset1 <- subset(subset1_0, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
 
@@
 
-data0 <- subset1
+data0 <- subset(subset1_0, select = c("country.code","year","edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
 data <- na.omit(data0)
@@
 #Find the ridge regression model
-set.seed(123)
-rate <- 0.6
-train.size <- round(nrow(data)*rate)
-sample(sample(nrow(data), train.size))
-training <- data[sample, ]
-test <- data[-sample, ]
+# set.seed(123)
+# rate <- 0.8
+# train.size <- round(nrow(data)*rate)
+# sample(sample(nrow(data), train.size))
+# training <- data[sample, ]
+# test <- data[-sample, ]
+#
 
@@
 ```{r}
-training <- na.omit(training)
-train.x <- model.matrix(pov ~., data = training)[,-1]
+split_train_test <- function(data) {
+set.seed(1984)
+# split data
+seeds <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() == 1)
+
+plants <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() != 1)
+
+isComplete <- which(complete.cases(plants))
+idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
+
+training <- plants[-idx,] %>% 
+  rbind(seeds)
+test <- plants[idx,]
+
+return(list(training, test))
+}
+```
+
+```{r}
+training <- split_train_test(data)[[1]]
+test <- split_train_test(data)[[2]]
+```
+
+```{r}
+ntrain <- 1:nrow(training)
+combine <- rbind(training, test)
+combine.mtrx <- model.matrix(pov ~ ., data = combine)
+train.mtrx <- combine.mtrx[ntrain,]
+test.mtrx <- combine.mtrx[-ntrain,]
+
+train.x <- train.mtrx[,-1]
 train.y <- training$pov
-test.x <- model.matrix(pov ~., data = test)[,-1]
-test.y <- test$pov
+
+test.x <- test.mtrx[,-1]
+text.y <- test$pov
 ```

### 2022-11-16 src/All variables.Rmd
Update All variables.Rmd

@@
 
+First we will write the function "eval_results(true,predict) to evaluate the models
 ```{r}
@@
     RMSE = RMSE(predict, true),
-    MAPE = MAPE(predict, true),
     Rsquare = R_square
@@
 ```
-
+First we read the csv file for the countries containing the dataset. We will then put the variables we need in to subset1.
 ```{r}
@@
 df
-subset1 <- df %>% select("pov","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
+subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
 subset1
@@
 
+Next we will build a SLR model. We will put 80% of dataset to train and the rest to test.
 ```{r}
@@
 
+The model output will be placed in lm.slr and can be viewed in summary(lm.slr). THe r-squared value is 0.9619.
 ```{r}
-lm.slr <- lm(pov ~ income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
+lm.slr <- lm(pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
 summary(lm.slr)
@@
 ```
+Next we will be perfoming regularisation. We have to use this function as not all the varibales are numerical. Therefore to overcome it we build the function below.
 ```{r}
 subset1$income <- factor(subset1$income)
-subset1$income <- as.numeric(subset1$income)
+subset1$country.code <- factor(subset1$country.code)
 
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
+
+  return(sd(vct, na.rm = T))
+}
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+
+  return(vct/scl)
+}
+
+scale <- function(data) {
+  unlist(lapply(data, sd0))
+}
+
+standardise <- function(data) {
+  scaler <- scale(data)
+
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}
 ```
 
+Next we will perform standardistation to the dataset.
 ```{r}
@@
 df0 <- subset1
-scaler <- apply(X = df0, MARGIN = 2, FUN = sd)
-subset1 <- as.data.frame(apply(df0, 2, function(x) x/sd(x, na.rm = TRUE)))
 subset1
@@
 ```
+
+Next we will plot a pov vs income scatter plot to see the effect of income on pov.
+From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov.
+This is consistent with our hypothesis that countries that are richer do not have as much poverty issue as the poorer countries.
+```{r}
+ggplot(data = subset1, aes(x = income, y = pov)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  labs(title = "Pov vs. Income ",
+       x = "income",
+       y = "pov")
+
+```
+
 ```{r}
@@
 ```{r}
-subset2 <- subset(subset1,select =c("pov","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
+subset2 <- subset(subset1,select =c("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
 subset2 <- na.omit(subset2)
+subset2
 train.x <- data.matrix(subset(subset2, select = -pov))
@@
 ```{r}
-corrplot(cor(subset2, use = "complete.obs"), 
-         method = "number",
-         type = "upper",
-         tl.col = "black", 
-         tl.srt = 30
-         )
+split_train_test <- function(data) {
+set.seed(1984)
+# split data
+seeds <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() == 1)
+
+plants <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() != 1)
+
+isComplete <- which(complete.cases(plants))
+idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
+
+training <- plants[-idx,] %>% 
+  rbind(seeds)
+test <- plants[idx,]
+
+return(list(training, test))
+}
 ```
@@
 ```{r}
-set.seed(123)
-rate <- 0.8
-train.size <- round(nrow(subset2)*rate)
-sample(sample(nrow(subset2), train.size))
-training <- subset2[sample, ]
-test <- subset2[-sample, ]
 
+training <- split_train_test(subset2)[[1]]
+test <- split_train_test(subset2)[[2]]
 ```
+
+```{r}
+ntrain <- 1:nrow(training)
+combine <- rbind(training, test)
+combine.mtrx <- model.matrix(pov ~ ., data = combine)
+train.mtrx <- combine.mtrx[ntrain,]
+test.mtrx <- combine.mtrx[-ntrain,]
+
+train.x <- train.mtrx[,-1]
+train.y <- training$pov
+
+test.x <- test.mtrx[,-1]
+text.y <- test$pov
+```
+
 ```{r}
@@
 ```
-
-```{r}
-set.seed(123)
-cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
-glm_ridge <- glmnet(train.x,
-                    train.y,
-                    alpha = 0,
-                    lambda = cv_ridge$lambda.min)
-t(coef(glm_ridge))
-
-```
-
-```{r}
-train.x <- model.matrix(pov ~., data = training)[, -1]
-train.y <- training$pov
-test.x <- model.matrix(pov ~., data = test)[,-1]
-test.y <- test$pov
-```
-
-```{r}
-model_ridge_trial1 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.1)
-t(coef(model_ridge_trial1))
-```
-
-```{r}
-model_ridge_trial2 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.5)
-t(coef(model_ridge_trial2))
-```
-
-```{r}
-model_ridge_trial3 <- glmnet(train.x, train.y, alpha = 0, lambda = 1)
-t(coef(model_ridge_trial3))
-```
-
-```{r}
-table_summary <- cbind(coef(model_ridge_trial1),
-                       coef(model_ridge_trial2),
-                       coef(model_ridge_trial3))
-
-colnames(table_summary) <- c("lambda = 0.1",
-                             "lambda = 0.5",
-                             "lambda = 1")
-
-table_summary
-```
-
-```{r}
-set.seed(123)
-cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
-glm_lasso <- glmnet(train.x,train.y, 
-                    alpha = 1, lambda = cv_lasso$lambda.min)
-t(coef(glm_lasso))
-```
-
-```{r}
-# Obtain performance of MLR model
-fit <- model1$fitted.values
-true <- subset2$pov
-summary_model1 <- eval_results(true, fit)
-
-# Obtain performance of ridge model
-fit <- predict(glm_ridge, newx = train.x)
-true <- train.y
-summary_ridge <- eval_results(true, fit)
-
-# Obtain performance of LASSO model
-fit <- predict(glm_lasso, newx = train.x)
-true <- train.y
-summary_lasso <- eval_results(true, fit)
-
-# Combine all performances into a single output
-summary <- rbind(summary_model1, summary_ridge, summary_lasso)
-rownames(summary) <- c("MLR", "Ridge_model", "LASSO_model" )
-colnames(summary)[5] <- "R^2"
-knitr::kable(summary, digits = 3)
-
-```
-
-```{r}
-knitr::kable(summary_MLR, digits = 3)
-knitr::kable(summary_ridge[,-5], digits = 3)
-```
-
-```{r}
-# Fit Ridge Regression Model
-set.seed(123)
-cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
-glm_ridge <- glmnet(train.x, train.y, 
-                    alpha = 0, lambda = cv_ridge$lambda.min)
-
-# Evaluate Ridge Regression Model on Training Set
-fit <- predict(glm_ridge, newx = train.x, type = "response")
-true <- train.y
-summary_ridge_train <- eval_results(true, fit)
-
-# Evaluate Ridge Regression Model on Test Set
-fit <- predict(glm_ridge, newx = test.x, type = "response")
-true <- test.y
-summary_ridge_test <- eval_results(true, fit)
-
-# Summarise Ridge Regression Model Performance into Table
-summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
-rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
-```
-
-```{r}
-# Fit LASSO Model
-set.seed(123)
-cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
-glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
-                    lambda = cv_lasso$lambda.min)
-
-# Evaluate LASSO Model on Training Set
-fit <- predict(glm_lasso, newx = train.x)
-true <- train.y
-summary_lasso_train <- eval_results(true, fit)
-
-# Evaluate LASSO Model on Test Set
-fit <- predict(glm_lasso, newx = test.x)
-true <- test.y
-summary_lasso_test <- eval_results(true, fit)
-
-# Summarise LASSO Model Performance into Table
-summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
-rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
-```
-
-```{r}
-summary_3models <- rbind(summary_MLR3_train, 
-                         summary_ridge_train,                               
-                         summary_lasso_train, 
-                         summary_MLR3_test,                             
-                         summary_ridge_test, 
-                         summary_lasso_test)
-
-rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
-                               "LASSO Model_Train",  "Baseline MLR_Test",  
-                               "Ridge Model_Test", "LASSO Model_Test" )
-
-knitr::kable(summary_3models[1:3,-5], digits = 3)
-knitr::kable(summary_3models[4:6,-5], digits = 3)
-```

### 2022-11-16 src/draft.Rmd
I'm fucking done

--- 2022-11-16 src/draft.Rmd
@@
 ```
-
 Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
@@
 ```
-
 The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
@@
 ```
-
 ###  3.3.3. Interpretation
@@
 set.seed(1984)
-# split data
+# split data, make sure all the coutries
 seeds <- countries1 %>% 
@@
 res <- do.call(rbind, lapply(models, \(f) coefficients(f$model)[var,]))
+res <- cbind(data.frame(Set = 1:3), res)
 kable(res)
@@
 
+These data seems valid and there is no convincing reasons to remove them from the model (in the case of Guinea and Gambia, maybe there was some false reports because the jump were quite suspicious, but we need more evidence to conclude that they are false data).
+
 ###  3.4.3. Interpretation
+
+Final model coefficients (exclude `country.code`):
+
+```{r}
+var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","lbr.part","gcf","lfdi")
+
+res <- do.call(rbind, lapply(models, \(f) coefficients(f$model)[var,]))
+res <- cbind(data.frame(Set = 1:3), res)
+kable(res)
+```
+
+We should convert them to un-standardised value for easier interepretation.
+
+```{r}
+var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","lbr.part","gcf","lfdi")
+
+res <- do.call(rbind, lapply(models, \(f) {
+  # get scaler saved in the model
+  scaler <- f$scaler
+  # categorical has scaler = 1
+  scaler[c("incomeL","incomeLM","incomeUM")] <- 1
+  # beta = std.beta * sigma_y / sigma_x
+  coefficients(f$model)[var,] * scaler['pov'] / scaler[var]
+}))
+kable(res)
+```
 

### 2022-11-17 src/All variables.Rmd
Update All variables.Rmd

@@
 ```
-Next we will be perfoming regularisation. We have to use this function as not all the varibales are numerical. Therefore to overcome it we build the function below.
+Next we will be performing standardisation.However not all the variables are numerical.
+Therefore to overcome it we build the function below.
+This will allow us to perform standardisation to the dataset.
 ```{r}
@@
 From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov.
-This is consistent with our hypothesis that countries that are richer do not have as much poverty issue as the poorer countries.
+This is consistent with our intuition that countries that are richer do not have as much poverty issue as the poorer countries.
 ```{r}
@@
 ```
+Next we will build a MLR model with data from subset 1 to see the relationship between dependent variable:pov and the other variables
 
@@
 ```
+Next we will perform regularisation. We will split the data to predictor variable "train.x" and response variables "train.y"
 ```{r}
@@
 
+Next we will perform splitting of data to train and test.
 ```{r}
@@
 }
-```
-
-```{r}
 
@@
 test <- split_train_test(subset2)[[2]]
-```
 
-```{r}
 ntrain <- 1:nrow(training)
@@
 
+We then build a MLR model with the training model to train it.
+
 ```{r}
@@
 
+We then evaluate the model between train and test
 ```{r}
@@
 ```
+
+We can see the evaluation of the model from the function below. From the train model, the R^squared value will be 0.965 and for the test model it will be 0.881.
 
@@
 ```
-
+Next we will be performing ridge and lasso regression model using the following code. We will split the data to train and test.
 ```{r}
@@
 
+Next we will perform the Ridge regression model.
 ```{r}
@@
 ```
+
+Next we will perform Lasso Regression model
 
@@
 
+We will compare the 3 models that we have build, Multi linear ,Ridge and Lasso regression models.
+We evaluate the 3 models together and compare both train and test models of each of the 3 regression model.
+
 ```{r}
@@
 ```
+From the 3 models we can see that the original multi linear regression model is the best one with the highest R^squared value.
+However the Lasso Regression Model has the lowest MSE, MAE and RMSE.
+We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting.
+Next we will build other models to see if they work better.

### 2022-11-17 src/Selectively_remove_variables.Rmd
Update Selectively_remove_variables.Rmd

@@
 ```
-This approach takes variables that contain less than 35% of missing data.The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models.
+###  3.2.1. Model Fitting
+##### A. Ordinary Linear Regression
+Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables.
+This section attempts to build a model on variables that have missing rates of below 35%.
+
+The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models.
 Hence, we transformed the data with the mutate() function before building the machine learning model.
@@
 ```
+
+The summary of `lm_1` suggests that the model still contains predictor variables with pr(>|t|) > 0.05.
+We can imply that these variables do not provide a high significance to our model and are thus able to be removed.
+To avoid removing significant variables, we shall conduct the removal of the variables deliberately.
+In `lm_1`, we notice that the variable `lgdp.dflt` has a high Pr(>|t|).
+Hence, we remove `lgdp.dflt`.
+
 ```{r}
@@
 ```
-Our model shows that among all the 13 predictor variables that we started with, only 10 of them showed significance in building the model.
+Our model shows that among all the 13 predictor variables that we started with, only 11 of them showed significance in building the model.
 
@@
 ```
+The MAPE shows `Inf` due to the presence of 0 values in the Y variable.
 
+##### c. Regularization
+Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model.
+
+Next, we will perform Ridge regression and LASSO regression in a dataset that contains `country.code` and `year` in addition to the previous dataset that we used for the previous linear regression model.
 ```{r}
-#subsetting df.countries to only contain variables present in lm_7
+#Subsetting the data
 
@@
 
-```{r}
-#Find the ridge regression model
-# set.seed(123)
-# rate <- 0.8
-# train.size <- round(nrow(data)*rate)
-# sample(sample(nrow(data), train.size))
-# training <- data[sample, ]
-# test <- data[-sample, ]
-#
-
-```
+Splitting the aforementioned dataset into training and testing data.
 
@@
 ```
-
+Further splitting the training and testing data into  `train.x` and `train.y`
 ```{r}
@@
 
+Building a model with the lambda that outputs the smallest MSE. 
 ```{r}
@@
 ```
-
+The ridge regression model shows that some countries have a higher influence on the model than others.
+At first glance,it seems like the Y variable, `pov`, depends heavily on whether or not it belongs to certain countries.To further illustrate the significance of `country.code`, we will plot the ridge regression graph for different lambda values.
 ```{r}
@@
 ```
-
-```{r}
-plot(ridge_model, xvar = "lambda", label = TRUE)
-
-#customise the plot with labels
-add_lbs <- function(fit, offset_x = 2.5) {
-  L <- length(fit$lambda[L]) + offset_x
-  y <- fit$beta[,L]
-  labs <- names(y)
-  text(x, y , labels = labs, cex = 1.5)
-
-  title(main = "Coefficients of Predictors vs Log(Lambda)",cex.main = 1.4, adj = 0, line = 3)
-  title(sub = "Number of Predictors", cex.sub = 1.1, adj = 0.55, line = -21.5)
-  add_lbs(ridge_model)
-  legend("topright", lwd = 1, col = 1:6, legend = colnames(train.x), cex = .7)
-
-}
-```
+Plotting the coefficients against different lambda values.
 
@@
 ```
-```{r}
-#cross - validation
-set.seed(123)
-cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0, type.measure = "mse")
 
-```
+Interesting! For a low lambda value, the model exhibits `cntr.MDG`, which represents Madagascar, followed by `cntr.ZMB`, which represents Zimbabwe,  as key predictor variables.
+This means that the model depends highly on whether the country is Madagascar or Zimbabwe, followed by some other significant countries.
+The other variables that previously defined the model is now undermined by the presence of the `country.code` variable and does not show high influence.
 
-```{r}
-cv_ridge
-cv_ridge$lambda.min
-cv_ridge$lambda.1se
-```
+We then proceeded to evaluate the results
 
-```{r}
-plot(cv_ridge)
-abline(v=log(cv_ridge$lambda.min), col = "red", lty=5)
-text(log(cv_ridge$lambda.min)+1, 1.4, 
-     labels= paste0("lambda_min = ", 
-                    round(cv_ridge$lambda.min, digits = 3)), 
-     cex = 1.1)
-
-# Label for lambda.1se
-abline(v=log(cv_ridge$lambda.1se), 
-       col = "red", lty=5)
-text(log(cv_ridge$lambda.1se)+1, 1.25, 
-     labels= paste0("lambda_1se = ", 
-                    round(cv_ridge$lambda.1se, digits = 3)), cex = 1.1)
-
-```
-
-```{r}
-#train the ridge regression model using lambda.min
-
-glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
-t(coef(glm_Ridge))
-```
 ```{r}
@@
 ```
+The training dataset produced a much higher value compared to the regular multiple regression model shown above.
+
+Fitting the model to the test dataset,
 
@@
 ```
+A slightly lower R^2 value, but still proves that the model is good enough.
 
+Below is the summary of the results of the training and testing dataset. 
 ```{r}
-summary_Ridge <- rbind(summary_Ridge_train[-5], summary_Ridge_test[-5])
+summary_Ridge <- rbind(summary_Ridge_train, summary_Ridge_test)
 rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
@@
 ```
-
-```{r}
-#Obtain fitted and true values
-
-fit <- predict(glm_Ridge, train.x)
-true <- train.y
-
-```
-
-```{r}
-#Plot residuals s
-plot(fit, true - fit, 
-     main = "Residual Plot of Ridge Regression Model on the Training dataset",
-     xlab = "Fitted value",
-     ylab = "Standardised Residuals")
-abline(h = 0, col = "red", lwd = 2)
-abline(h = 1, col = "red", lty = 2, lwd = 2)
-abline(h = -1, col = "red", lty = 2, lwd = 2)
-
-```
-Constant varince assumption not satisfied. What to do? Most likely need to transform the variable
-TO TRANSFORM TO BOXCOX
-
+Next, we attempted LASSO regression as well. The steps are essentially similar to that of Ridge regression. 
 ```{r}
@@
 ```
+It is evident that the number of variables included in the model is less than that of ridge regression.
+This is in accordance with the nature of the LASSO regression model that reduces the coefficients of certain predictor variables to zero for model interpretability.
+
+```{r}
+lambda <- 10^seq(-6, 2, length = 100)
+lasso_model = glmnet(train.x, train.y, alpha = 1, lambda = lambda)
+
+t(coef(lasso_model))
+```
+```{r}
+plot_glmnet(lasso_model)
+```
+```{r}
+plot_glmnet(ridge_model)
+```
 
@@
 summary_Lasso_train
+```
+```{r}
+plot_glmnet(lasso_model)
 ```
@@
 ```
+```{r}
+summary_lasso <- rbind(summary_Lasso_train, summary_Lasso_test)
+rownames(summary_lasso) <- c("LASSO_train", "LASSO_test")
+knitr::kable(summary_lasso, digits = 3)
+
+```
+###  3.2.2. Assessment
+
+This section recapitulates the three models that we have generated so far through the method of gradually removing variables from the original dataset.
+
+```{r}
+knitr::kable(summary_lasso, digits = 3)
+```
+
+```{r}
+knitr::kable(summary_Ridge, digits = 3)
+```
+```{r}
+summary(lm_4)
+```
+
+Our first model, `lm_4`, generated with regular multiple linear regression exhibits the lowest R^2 value.
+The model shows that `incomeL` is the most significant predictor variable with the highest coefficient.
+One interesting observation is that `incomeLM` and `incomeUM` has a negative correlation with the pov level, in contrast to `incomeL`.
+This shows that countries that have lower income tend to have high `pov` value.
+On the other hand, countries that are classified to have lower middle income and upper middle income seem to have lower `pov` values, as shown by their negative coefficients in the regression model.
+
+Apart from income, some other variables that contribute more significantly to the model are `lgdp.pc`, with a coefficient of -3.13114, exhibiting a stronger negative correlation, and `pop.gwth.urban`, exhibiting a stronger positive correlation with a coefficient of 1.90680.
+This is counter-intuitive, as the notion is that population growth in cities will help drive positive economic change due to the number of productive people living in the city.
+The model also shows that `pop.gwth.rural` gives a less significant impact on `pov` as compared to  `pop.gwth.rural`.
+
+Both the Ridge regression and LASSO regression models exhibit similar result metrics (MSE, MAE, MAPE, R^2). This result is expected.
+
+```{r}
+plot_glmnet(lasso_model)
+```
+
+```{r}
+plot_glmnet(ridge_model)
+```
+###  3.2.3. Interpretation
+
+The comparison between the three models suggests that `country.code` affects the models generated in a significant way.

### 2022-11-17 src/Selectively_remove_variables.Rmd
Update Selectively_remove_variables.Rmd

@@
 ```{r}
-knitr::kable(summary_lasso, digits = 3)
-```
-
-```{r}
-knitr::kable(summary_Ridge, digits = 3)
-```
-```{r}
 summary(lm_4)
@@
 
-Both the Ridge regression and LASSO regression models exhibit similar result metrics (MSE, MAE, MAPE, R^2). This result is expected.
+```{r}
+knitr::kable(summary_lasso, digits = 3)
+```
+
+```{r}
+knitr::kable(summary_Ridge, digits = 3)
+```
+The MSE for the LASSO regression model is slightly higher than that of the Ridge regression model.
+This to be expected, as the Ridge regression model aims for an improved accuracy while the LASSO model aims for an improved interpretability.
+Both models seem to exhibit very high R^2 value for both the training and testing dataset.
 
@@
 plot_glmnet(lasso_model)
+```
+```{r}
+
 ```
@@
 ```
-###  3.2.3. Interpretation
-
-The comparison between the three models suggests that `country.code` affects the models generated in a significant way.
+```{r}
+plot(lm_4, which = 5)
+```
+The `Residual vs Leverage` plot of `lm_4` suggests that heteroskedasticity might be present in the model.
+This means that the variance of the residuals is unequal over a range of observed values.
+This problem exists in the dataset rather than the models generated.

### 2022-11-17 src/draft.Rmd
update
Update draft.Rmd

--- 2022-11-16 src/draft.Rmd
@@
     df_print: paged
-  word_document: default
 bibliography: citations.bib
@@
 ```{r setup, include=FALSE}
-library(knitr)
 knitr::opts_chunk$set(echo = TRUE, warning = F)
 knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
-# inline hook to 2 digits
-inline_hook <- function(x) {
-    if(is.numeric(x)) x <- round(x, 2)
-    paste(as.character(x), collapse=", ")
-}
-knit_hooks$set(inline = inline_hook)
-# code chunk out put to 
-options(digits = 5)
 ```
@@
 ```{r, message = F}
+# Package names
+# packages <- c("MASS", "knitr", "readr", "tidyr", "dplyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "needs")
+# 
+# # Install packages not yet installed
+# installed_packages <- packages %in% rownames(installed.packages())
+# if (any(installed_packages == FALSE)) {
+#   install.packages(packages[!installed_packages])
+# }
+# 
+# # Packages loading
+# invisible(lapply(packages, library, character.only = TRUE))
+# 
+# prioritize(dplyr)
+
 library(needs)
 
-needs("dplyr", "MASS", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce", "lmtest")
+needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet")
 
@@
 # 1. Introduction (To be done)
+The project aim is to explore different models using the same dataset to evaluate which method will be the best to build a model for the dependent variable.
+The dataset we are using for this project has 1901 data values from 23 different variables.
+Our plan for this project is to first build a model with all the variables in the dataset.
+We will then follow it up by builidng a model without the missing values and build a ridge regression and Lasso regression model.
+Our second approach will be to selectively remove variables from the model and to build the best model  using regularisation and regression.
+Our next approach was to add the variables into the model one at a time to see if a different model can be achieved.
+Lastly, we performed imputation on our model to tackle the missing values present in our dataset.We then compared all of this approaches to come to a conclusion.
 
@@
 ggplot(countries, aes(x = pov)) +
-  geom_histogram(aes(y = ..density..), bins = 30) +
+  geom_histogram(aes(y = ..density..)) +
   geom_density(alpha = .2, fill = "blue")
@@
 ggplot(poverty.headcount, aes(x = year)) +
-  geom_histogram(bins = 30)
+  geom_histogram()
 
@@
 ggplot(pov.98.18, aes(x = pov)) +
-  geom_histogram(aes(y = ..density..), bins = 30) +
+  geom_histogram(aes(y = ..density..)) +
   geom_density(alpha = .2, fill = "blue") +
@@
   geom_point() +
-  geom_smooth(formula = y~x, method = loess)
+  geom_smooth()
 ```
@@
   geom_point() +
-  geom_smooth(formula = y~x, method = loess) +
+  geom_smooth() +
   facet_grid(rows = vars(type))
@@
 
-## 2.4. Assumption Check
-We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data.
-[Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
-**1. Linear relationship:**  
-Use correlation matrix to check linearity
-```{r, warning = FALSE}
-# select only numeric data
-countries.num <- countries %>% select(where(is.numeric))
-chart.Correlation(countries.num, histogram = TRUE, pch = 19)
-```
-Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
-```{r}
-# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
-countries.num2 <- countries.num %>% 
-  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
-  filter(!is.na(value) & !is.na(pov))
-
-drawGraph <- function(indvar, data) {
-  ggplot(data %>% 
-           filter(variable == indvar),
-         aes(x = value, y = pov)) +
-    geom_point() +
-    geom_smooth(method = lm) +
-    labs(title = paste("Relationship pov ~", indvar),
-         x = indvar,
-         y = "pov")
-}
-
-ivs <- distinct(countries.num2, variable)[[1]]
-
-# for (indvar in ivs) {
-#   print(
-#     ggplot(countries.num2 %>% 
-#            filter(variable == indvar),
-#          aes(x = value, y = pov)) +
-#     geom_point() +
-#     geom_smooth(formula = y~x, method = lm) +
-#     labs(title = paste("Relationship pov ~", indvar),
-#          x = indvar,
-#          y = "pov")
-#   )
-# }
-
-p <- ggplot(countries.num2, 
-     aes(x = value, y = pov)) +
-geom_point() +
-geom_smooth(formula = y~x, method = lm) +
-labs(title = paste("pov ~ variable"),
-     x = "variable",
-     y = "pov") +
-facet_wrap_paginate(~ variable,
-                    ncol = 3,
-                    nrow = 1,
-                    scales = "free")
-
-requiredPages <- n_pages(p)
-
-for (i in 1:requiredPages) {
-  ggplot(countries.num2,
-         aes(x = value, y = pov)) +
-  geom_point() +
-  geom_smooth(formula = y~x, method = lm) +
-  labs(title = paste("pov ~ variable"),
-       x = "variable",
-       y = "pov") +
-  facet_wrap_paginate(~ variable,
-                      ncol = 3,
-                      nrow = 1,
-                      scales = "free",
-                      page = i) -> p
-  print(p)
-}
-```
-
-Most variables display linear relationship with some exceptions, which are more appropriate to assume a logarithmic relationship. Gender equality should be viewed as a categorical data.
-```{r}
-logIvs <- c("fdi", "gdp.dflt", "gdp.pc")
-
-for (indvar in logIvs) {
-  print(
-    ggplot(countries.num2 %>% 
-           filter(variable == indvar),
-         aes(x = log(value), y = pov)) +
-    geom_point() +
-    geom_smooth(method = lm) +
-    labs(title = paste0("pov ~ log(", indvar, ")"),
-         x = paste0("log(", indvar, ")"),
-         y = "pov")
-  )
-}
-```
-
-The relationship are more linear after the transformation.
-
-```{r}
-gdr.eql <- countries.num2 %>% 
-  filter(variable == "gdr.eql") %>% 
-  mutate(value = factor(value))
-
-ggplot(gdr.eql,
-       aes(x = value, y = pov)) +
-  geom_boxplot() +
-  labs(title = paste("pov ~ gdr.eql"),
-       x = "gdr.eql",
-       y = "pov")
-```
-There's a significant correlation between gender equality and poverty.  
-Correlation coefficients between `pov` and predictors.
-
-```{r}
-# transform some variables
-countries.num3 <- countries.num %>% 
-  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
-  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
-
-# name of the independent variables
-cn <- colnames(countries.num3)[-c(1,2)]
-
-# correlation of independent variable and pov
-corWithPov <- function(indevar, data) {
-  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
-}
-
-cor.pov <- sapply(cn, corWithPov, data = countries.num3)
-kable(cor.pov)
-```
-
-`edu.ter`, `mil`, `fdi`, `lbr.part`, `unemp`, `gdp.dflt`, `gcf` have negligible correlation with `pov`.
-`lgdp.pc`, `lfdi`, and `lgdp.dflt` have stronger linear relationship with `pov` than their un-transformed counterparts.
-
-```{r}
-countries1 <- countries1 %>% 
-  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
-  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
-```
-
-**2. Predictors are independent**
-There should be no correlation/multicolinearity between each pair of predictors.
-
-```{r}
-countries.arr <- simplify2array(countries.num %>% 
-                                  select(-c("year", "pov")))
-
-cor.mtx <- rcorr(countries.arr, type = "spearman")
-
-round(cor.mtx$r, 2)
-
-corrplot(cor.mtx$r, type = "upper", order = "hclust", 
-         tl.col = "black", tl.srt = 45)
-```
-Some significant correlations can be found between `gdr.eql` with `edu.pri`, `pop.gwth.total`, and `fdi` with `mil`, etc.
-We expect these variables to be eliminated in the VIF test.
-This might be a good property for imputation (3.2.4)
-
-## 2.5. Data Source
+## 2.4. Data Source
 
@@
 # 3. Model Selection and Interpretation
-
-We can now conduct linear regressions following the approaches mentioned above to address missing values issues.
-
-## 3.1. Use Complete Cases (To be done)
-###  3.1.1. Model Fitting
-###  3.1.2. Assessment
-###  3.1.3. Interpretation
-
-## 3.2. Selectively remove variables with high missing rate (To be done)
-**1. About selectively removing variables with high missing rate**
-
-Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables.
-This section attempts to build a model on variables that have missing rates of below 35%.
-
-The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models.
-Hence, we transformed the data with the mutate() function before building the machine learning model.
-```{r}
-subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
-
-subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
-
-subset1 <- subset(subset1, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
-
-subset1 <- na.omit(subset1)
-```
-We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step.
-
-```{r}
-#helper function
+## 3.1. Assumption Check
+We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data.
+[Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
+**1. Linear relationship:**  
+Use correlation matrix to check linearity
+```{r, warning = FALSE}
+# select only numeric data
+countries.num <- countries %>% select(where(is.numeric))
+chart.Correlation(countries.num, histogram = TRUE, pch = 19)
+```
+Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
+```{r}
+# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
+countries.num2 <- countries.num %>% 
+  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
+  filter(!is.na(value) & !is.na(pov))
+
+drawGraph <- function(indvar, data) {
+  ggplot(data %>% 
+           filter(variable == indvar),
+         aes(x = value, y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste("Relationship pov ~", indvar),
+         x = indvar,
+         y = "pov")
+}
+
+ivs <- distinct(countries.num2, variable)[[1]]
+
+for (indvar in ivs) {
+  print(
+    ggplot(countries.num2 %>% 
+           filter(variable == indvar),
+         aes(x = value, y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste("Relationship pov ~", indvar),
+         x = indvar,
+         y = "pov")
+  )
+}
+```
+
+Most variables display linear relationship with some exceptions, which are more appropriate to assume a logarithmic relationship. Gender equality should be viewed as a categorical data.
+```{r}
+logIvs <- c("fdi", "gdp.dflt", "gdp.pc")
+
+for (indvar in logIvs) {
+  print(
+    ggplot(countries.num2 %>% 
+           filter(variable == indvar),
+         aes(x = log(value), y = pov)) +
+    geom_point() +
+    geom_smooth(method = lm) +
+    labs(title = paste0("Relationship pov ~ log(", indvar, ")"),
+         x = paste0("log(", indvar, ")"),
+         y = "pov")
+  )
+}
+```
+
+The relationship are more linear after the transformation.
+
+```{r}
+gdr.eql <- countries.num2 %>% 
+  filter(variable == "gdr.eql") %>% 
+  mutate(value = factor(value))
+
+ggplot(gdr.eql,
+       aes(x = value, y = pov)) +
+  geom_boxplot() +
+  geom_smooth(method = lm) +
+  labs(title = paste("Relationship pov ~ gdr.eql"),
+       x = "gdr.eql",
+       y = "pov")
+```
+There's a significant correlation between gender equality and poverty.  
+Correlation coefficients between `pov` and predictors.
+
+```{r}
+# transform some variables
+countries.num3 <- countries.num %>% 
+  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
+  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
+
+# name of the independent variables
+cn <- colnames(countries.num3)[-c(1,2)]
+
+# correlation of independent variable and pov
+corWithPov <- function(indevar, data) {
+  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
+}
+
+cor.pov <- sapply(cn, corWithPov, data = countries.num3)
+kable(cor.pov)
+```
+
+`edu.ter`, `mil`, `fdi`, `lbr.part`, `unemp`, `gdp.dflt`, `gcf` have negligible correlation with `pov`.
+`lgdp.pc`, `lfdi`, and `lgdp.dflt` have stronger linear relationship with `pov` than their un-transformed counterparts.
+
+```{r}
+countries1 <- countries1 %>% 
+  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
+  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
+```
+
+**2. Predictors are independent**
+There should be no correlation/multicolinearity between each pair of predictors.
+
+```{r}
+countries.arr <- simplify2array(countries.num %>% 
+                                  select(-c("year", "pov")))
+
+cor.mtx <- rcorr(countries.arr, type = "spearman")
+
+round(cor.mtx$r, 2)
+
+corrplot(cor.mtx$r, type = "upper", order = "hclust", 
+         tl.col = "black", tl.srt = 45)
+```
+Some significant correlations can be found between `gdr.eql` with `edu.pri`, `pop.gwth.total`, and `fdi` with `mil`, etc.
+We expect these variables to be eliminated in the VIF test.
+This might be a good property for imputation (3.2.4)
+
+## 3.2. Ordinary Multiple Linear Regression
+
+We conduct a normal linear regression, following the approaches mentioned above to address missing values issues.
+
+### 3.2.1. Use Complete Cases (To be done)
+#### 3.2.1.1. Model Fitting
+#### 3.2.1.2. Assessment
+#### 3.2.1.3. Interpretation
+
+```{r}
+library(tidyverse) # for data manipulations and visualisations
+library(ggpubr) # for correlation coefficient on plot
+library(gridExtra) # for arranging plots 
+library(e1071) # for skewness
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(corrplot)
+library(car)
+library(DescTools)
+library(glmnet)
+```
+
+First we will write the function "eval_results(true,predict) to evaluate the models
+```{r}
+eval_results <- function(true, predict) {
+  actual <- data.matrix(true)
+  SSE <- sum((predict - actual)^2)
+  SST <- sum((actual - mean(actual))^2)
+  R_square <- 1 - SSE / SST
+  data.frame(
+    MSE = MSE(predict, true),
+    MAE = MAE(predict, true),
+    RMSE = RMSE(predict, true),
+    Rsquare = R_square
+  )
+}
+```
+First we read the csv file for the countries containing the dataset. We will then put the variables we need in to subset1.
+```{r}
+df <-read.csv('../data/countries1.csv')
+df
+subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
+subset1
+```
+
+Next we will build a SLR model. We will put 80% of dataset to train and the rest to test.
+```{r}
+set.seed(10)
+index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
+train <- df[index, ]
+test <- df[-index, ]
+```
+
+The model output will be placed in lm.slr and can be viewed in summary(lm.slr). THe r-squared value is 0.9619.
+```{r}
+lm.slr <- lm(pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
+summary(lm.slr)
+
+```
+Next we will be performing standardisation.However not all the variables are numerical.
+Therefore to overcome it we build the function below.
+This will allow us to perform standardisation to the dataset.
+```{r}
+subset1$income <- factor(subset1$income)
+subset1$country.code <- factor(subset1$country.code)
+
 sd0 <- function(vct) {
@@
 }
-
-df.countries0 <- standardise(subset1)
-df.countries <- df.countries0
-head(df.countries)
-```
-
-###  3.2.1. Model Fitting
-##### A. Ordinary Linear Regression
-First, we split the data into training and testing data. The training dataset contains 80% of the initial dataset, and the testing data contains 20% of the whole dataset. 
-```{r}
+```
+
+Next we will perform standardistation to the dataset.
+```{r}
+str(subset1)
+df0 <- subset1
+subset1
+apply(subset1, 2, sd)
+
+```
+
+Next we will plot a pov vs income scatter plot to see the effect of income on pov.
+From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov.
+This is consistent with our intuition that countries that are richer do not have as much poverty issue as the poorer countries.
+```{r}
+ggplot(data = subset1, aes(x = income, y = pov)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  labs(title = "Pov vs. Income ",
+       x = "income",
+       y = "pov")
+
+```
+Next we will build a MLR model with data from subset 1 to see the relationship between dependent variable:pov and the other variables
+
+```{r}
+model1 <- lm(pov ~ ., data = subset1)
+summary(model1)
+vif(model1)
+
+```
+Next we will perform regularisation. We will split the data to predictor variable "train.x" and response variables "train.y"
+```{r}
+subset2 <- subset(subset1,select =c("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
+subset2 <- na.omit(subset2)
+subset2
+train.x <- data.matrix(subset(subset2, select = -pov))
+train.y = data.matrix(subset2$pov)
+```
+
+Next we will perform splitting of data to train and test.
+```{r}
+split_train_test <- function(data) {
+set.seed(1984)
+# split data
+seeds <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() == 1)
+
+plants <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() != 1)
+
+isComplete <- which(complete.cases(plants))
+idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
+
+training <- plants[-idx,] %>% 
+  rbind(seeds)
+test <- plants[idx,]
+
+return(list(training, test))
+}
+
+training <- split_train_test(subset2)[[1]]
+test <- split_train_test(subset2)[[2]]
+
+ntrain <- 1:nrow(training)
+combine <- rbind(training, test)
+combine.mtrx <- model.matrix(pov ~ ., data = combine)
+train.mtrx <- combine.mtrx[ntrain,]
+test.mtrx <- combine.mtrx[-ntrain,]
+
+train.x <- train.mtrx[,-1]
+train.y <- training$pov
+
+test.x <- test.mtrx[,-1]
+text.y <- test$pov
+```
+
+We then build a MLR model with the training model to train it.
+
+```{r}
+model3 <- lm(pov ~., data = training)
+summary(model3)
+
+```
+
+```{r}
+fit <- model3$fitted.values
+true <- training$pov
+summary_MLR3_train <- eval_results(true, fit)
+```
+
+We then evaluate the model between train and test
+```{r}
+fit <- predict(model3, newdata=test)
+true <- test$pov
+summary_MLR3_test <- eval_results(true, fit)
+```
+
+We can see the evaluation of the model from the function below. From the train model, the R^squared value will be 0.965 and for the test model it will be 0.881.
+
+```{r}
+summary_MLR <- rbind(summary_MLR3_train[-5], 
+                     summary_MLR3_test[-5])
+rownames(summary_MLR) <- c("Baseline MLR_Train", 
+                           "Baseline MLR_Test" )
+knitr::kable(summary_MLR, digits = 3)
+```
+Next we will be performing ridge and lasso regression model using the following code. We will split the data to train and test.
+```{r}
+# Training Data
+data <- training
+train.x <- data.matrix(data[,-ncol(data)])
+train.y <- data.matrix(data[,ncol(data)])
+
+# Test Data
+data <- test
+test.x <- data.matrix(data[,-ncol(data)])
+test.y <- data.matrix(data[,ncol(data)])
+```
+
+Next we will perform the Ridge regression model.
+```{r}
+# Fit Ridge Regression Model
 set.seed(123)
-index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
-train_1 <- subset1[index,]
-test_1 <- subset1[-index,]
-
-```
-Our first model includes all the variables with missing rate lower than 35%. This model is assigned to `lm_1`.
-```{r}
-lm_1 <- lm(data= train_1, formula = pov ~ .)
-summary(lm_1)
-```
-The summary of `lm_1` suggests that the model still contains predictor variables with pr(>|t|) > 0.05.
-We can imply that these variables do not provide a high significance to our model and are thus able to be removed.
-To avoid removing significant variables, we shall conduct the removal of the variables deliberately.
-In `lm_1`, we notice that the variable `lgdp.dflt` has a high Pr(>|t|).
-Hence, we remove `lgdp.dflt`.
-```{r}
-lm_2 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
-summary(lm_2)
-```
-Removing `lfdi`,
-```{r}
-lm_3 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
-summary(lm_3)
-```
-Further removing `hlth`,
-```{r}
-lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
-summary(lm_4)
-plot(lm_4, which = 5)
-```
-Our model shows that among all the 13 predictor variables that we started with, only 10 of them showed significance in building the model.
-
-##### B. VIF
-
-##### c. Regularization
-Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model. 
-Later on, we will compare the results of all the models we generated.
-```{r}
-data0 <- subset1
-data <- na.omit(data0)
-train.x <- data.matrix(subset(data, select = -pov))
-train.y <- data.matrix(data$pov)
-
-```
-Dividing into training and testing datasets.
-```{r}
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x, train.y, 
+                    alpha = 0, lambda = cv_ridge$lambda.min)
+
+# Evaluate Ridge Regression Model on Training Set
+fit <- predict(glm_ridge, newx = train.x, type = "response")
+true <- train.y
+summary_ridge_train <- eval_results(true, fit)
+
+# Evaluate Ridge Regression Model on Test Set
+fit <- predict(glm_ridge, newx = test.x, type = "response")
+true <- test.y
+summary_ridge_test <- eval_results(true, fit)
+
+# Summarise Ridge Regression Model Performance into Table
+summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
+rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
+```
+
+Next we will perform Lasso Regression model
+
+```{r}
+# Fit LASSO Model
 set.seed(123)
-rate <- 0.6
-train.size <- round(nrow(data)*rate)
-sample(sample(nrow(data), train.size))
-training <- data[sample, ]
-test <- data[-sample, ]
-```
-Further dividing into `train.x`, `train.y`, `test.x`, and `test.y`.
-```{r}
-training <- na.omit(training)
-train.x <- model.matrix(pov ~., data = training)[,-1]
-train.y <- training$pov
-test.x <- model.matrix(pov ~., data = test)[,-1]
-test.y <- test$pov
-```
-Using the lambda value that yields the lowest MSE value. 
-```{r}
-cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
-glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
-
-t(coef(glm_ridge))
-```
-
-```{r}
-#train the ridge regression model using lambda.min
-
-glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
-t(coef(glm_Ridge))
-```
-
-Function to evaluate the results 
-```{r}
-eval_results <- function(fit, true) {
-  actual <- data.matrix(true)
-  SSE <- sum((actual - fit)^2)
-  SST <- sum((actual - mean(actual))^2)
-  R_square <- 1 - SSE/SST
-  data.frame(
-    MSE = MSE(fit, true),
-    MAE = MAE(fit, true),
-    RMSE = RMSE(fit, true),
-    MAPE = MAPE(fit, true),
-    R2  = R_square
-
-  )
-}
-```
-
-```{r}
-fit <- predict(glm_Ridge, train.x)
+cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
+glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
+                    lambda = cv_lasso$lambda.min)
+
+# Evaluate LASSO Model on Training Set
+fit <- predict(glm_lasso, newx = train.x)
 true <- train.y
-summary_Ridge_train <- eval_results(fit, true)
-
-```
-
-```{r}
-fit <- predict(glm_Ridge, test.x)
+summary_lasso_train <- eval_results(true, fit)
+
+# Evaluate LASSO Model on Test Set
+fit <- predict(glm_lasso, newx = test.x)
 true <- test.y
-summary_Ridge_test <- eval_results(fit, true)
-
-```
-
-```{r}
-summary_Ridge <- rbind(summary_Ridge_train[-5], summary_Ridge_test[-5])
-rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
-knitr::kable(summary_Ridge, digits = 3)
-
-```
-
-###  3.2.2. Assessment
-###  3.2.3. Interpretation
-
-## 3.3. Update the data set as we select variables (To be done)
-###  3.3.1. Model Fitting
-###  3.3.2. Assessment
-###  3.3.3. Interpretation
-
-## 3.4. Imputation
+summary_lasso_test <- eval_results(true, fit)
+
+# Summarise LASSO Model Performance into Table
+summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
+rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
+```
+
+We will compare the 3 models that we have build, Multi linear ,Ridge and Lasso regression models.
+We evaluate the 3 models together and compare both train and test models of each of the 3 regression model.
+
+```{r}
+summary_3models <- rbind(summary_MLR3_train, 
+                         summary_ridge_train,                               
+                         summary_lasso_train, 
+                         summary_MLR3_test,                             
+                         summary_ridge_test, 
+                         summary_lasso_test)
+
+rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
+                               "LASSO Model_Train",  "Baseline MLR_Test",  
+                               "Ridge Model_Test", "LASSO Model_Test" )
+
+knitr::kable(summary_3models[1:3,-5], digits = 3)
+knitr::kable(summary_3models[4:6,-5], digits = 3)
+
+```
+From the 3 models we can see that the original multi linear regression model is the best one with the highest R^squared value.
+However the Lasso Regression Model has the lowest MSE, MAE and RMSE.
+We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting as many values will get removed.
+Next we will build other models to see if they work better.
+
+### 3.2.2. Selectively remove variables with high missing rate (To be done)
+#### 3.2.2.1. Model Fitting
+#### 3.2.2.2. Assessment
+#### 3.2.2.3. Interpretation
+
+### 3.2.3. Update the data set as we select variables (To be done)
+#### 3.2.3.1. Model Fitting
+#### 3.2.3.2. Assessment
+#### 3.2.3.3. Interpretation
+
+### 3.2.4. Imputation
 **1. What is imputation**  
@@
 a variable).
-The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), and use other variables to predict the missing values (much like regression).
-Since we observed some correlations during descriptive analysis, we expect the imputation algorithm to work well.
-
-We use **Multivariate Imputation By Chained Equations (MICE)** as primary tool for this technique, because of its wide acceptance in scientific studies (@imputWorks).
+The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), or predict the value using other variables.
+
+We use **Multivariate Imputation By Chained Equations (MICE)** as our primary tool for this technique, because of its wide acceptance in scientific studies (@imputWorks).
 In order to use this technique, we have to assume that the "missingness" of a field can be explained by the values in other columns, (e.g., If the countries is in North America, it's more likely to be missing as we have discovered in 2.2).
-The general ideas of the algorithm is to fill in the missing values, and improve it iteratively until predicted values converse to a stable point.
-The algorithmic details of MICE is very concisely (and enthrallingly) explained in @micealgo.
-
-We use relatively small parameters in the interest of time.
+The general ideas of the algorithm is iteratively predict the missing values based on other variables, and improve it after each iterations until predicted value converse to a stable point.
+The algorithmic details of MICE is very concisely explained in @micealgo.
+
+Our parameter for this algorithm is fairly standard:
 
@@
 
-###  3.4.1. Model Fitting
+#### 3.2.4.1. Model Fitting
 ##### A. Ordinary Linear Regression
-We generated 3 sets of imputated (training) data.
-We can build separate models using each data set, and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
+We generated 5 sets of imputated (training) data.
+We can build our model on each of these data sets and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details).
 With all the generated data sets.
@@
 ```{r}
-formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'
-
-# Using single imputed data set
-fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
+formula.str <- 'pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'
+
+# Using only the first imputed data set
+fit1 <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
+summary(fit1)
+fit1.pred <- predict(fit1, countries.test.4)
 
@@
 fit2.combined <- pool(fit2)
+summary(fit2.combined)
 # dummy lm model
-fit2.dummy <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
-# replace coefficients of dummy model
-name.coef <- names(fit2.dummy$coefficients)
+fit2.dummy <- lm(as.formula(formula.str), data = countries1)
+# replace coefficients of dummy model & predict
 fit2.dummy$coefficients <- fit2.combined$pooled$estimate
-names(fit2.dummy$coefficients) <- name.coef
+fit2.pred <- predict(fit2.dummy, countries.test.4)
 ```
@@
    RSS <- sum((pred - orig)^2)
-   TSS <- sum((orig - mean(orig))^2)
+   TSS <- sum((pred - mean(orig))^2)
    R2 <- 1 - RSS/TSS
@@
 ```{r}
-fit1.summ <- lapply(fit1, \(f) {
-  # number of variables
-  k <- length(f$coefficients) - 1
-  # predict value of test
-  pred <- predict(f, countries.test.4)
-  test.resid <- pred - countries.test.4$pov
-  data.frame(
-    `Train MSE` = mean(summary(f)$residuals^2),
-    `Test MSE` = mean(test.resid^2),
-    `Train MAE` = mean(abs(summary(f)$residuals)),
-    `Test MAE` = mean(abs(test.resid)),
-    `Train Adj.R2` = summary(f)$adj.r.squared,
-    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
-  )
-})
-
-# number of variables
-k <- length(fit2.dummy$coefficients) - 1
-# number of variables
-fit2.pred <- predict(fit2.dummy, countries.test.4)
-test.resid <- fit2.pred - countries.test.4$pov
-fit2.summ <- data.frame(
-  `Train MSE` = mean(summary(fit2.dummy)$residuals^2),
-  `Test MSE` = mean(test.resid^2),
-  `Train MAE` = mean(abs(summary(fit2.dummy)$residuals)),
-  `Test MAE` = mean(abs(test.resid)),
-  `Train Adj.R2` = pool.r.squared(fit2, adjusted = T)[,'est'],
-  `Test Adj.R2` = adjR2(fit2.pred, countries.test.4$pov, k)
-)
-
-res <- do.call(rbind.data.frame, fit1.summ)
-res <- rbind(res, fit2.summ)
-res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)
-
+k <- length(fit1$coefficients) - 1
+fit1.train.r2 <- summary(fit1)$adj.r.squared
+fit1.test.r2 <- adjR2(fit1.pred, countries.test.4$pov, k)
+fit2.train.r2 <- pool.r.squared(fit2, adjusted = T)[,'est']
+fit2.test.r2 <- adjR2(fit2.pred, countries.test.4$pov, k)
+res <- data.frame(model = c("Single", "Pooled"), train = c(fit1.train.r2, fit2.train.r2), test = c(fit1.test.r2, fit2.test.r2))
 kable(res)
@@
 
-Pooled model slightly improve performance. Test data fitting has better MSE and MAE than train data. However, Adjusted R-Squared is lower in test data. There might be some over-fitting in our models.
-
-```{r}
-# find coef of variable other than country.code
-var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","pop.gwth.urban","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
-
-res <- do.call(rbind, lapply(fit1, \(f) f$coefficients[var]))
-res <- rbind(res, fit2.dummy$coefficients[var])
-res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)
-kable(res)
-```
-
-In contradiction to expectation, `hlth` is positively correlated with `pov`, and lower-middle income countries (incomeLM) are less poor than high income countries (incomeH).
-`pop.gwth.total` is negatively related to `pov`, while `pop.gwth.rural` and `pop.gwth.urban` are positively related.
-They might be indications of over-fitting model.
-
-```{r}
-# fit model on imputated set no. 1
-summary(fit1[[1]])
-```
-
-There are also some weak variable, we should perform some variable selection.
-
-##### B. VIF
-
-Our data set contains a lots of variables. We can perform some variable selection to reduce over-fitting.
-
-```{r}
-gvif <- lapply(fit1, vif)
-gvif
-```
-
-[Some forum](https://stats.stackexchange.com/questions/70679/which-variance-inflation-factor-should-i-be-using-textgvif-or-textgvif) suggested that we should use the standard $GVIF^{1/(2 \cdot df)} < 2$ as equivalent to $GVIF < 4$ to account for high degree of freedom of some variables.
-
-```{r}
-# adjusted gvif higher than 3
-lapply(gvif, \(g) {
-  g[g[,3] > 3, 3]
-})
-```
-
-Remove the highest-ranked variables (`pop.gwth.total`, `pop.gwth.urban`, `pop.gwth.rural`, `gdp.pc`, `lgdp.pc`) and rebuild the models.
-
-```{r}
-formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+trade+lfdi+lgdp.dflt'
-
-# Using single imputed data set
-fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
-```
-
-Re-run GVIF analysis
-
-```{r}
-gvif <- lapply(fit1, vif)
-
-lapply(gvif, \(g) {
-  g[g[,3] > 2, 3]
-})
-```
-
-There're still some variable with adjusted GVIF > 2.
-We are interested in the effects of `hlth` (expenditure in Healthcare), and `mil` (expenditure in Military), so we won't remove those variables.
-We can remove `trade`.
-
-```{r}
-# remove trade
-formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+lfdi+lgdp.dflt'
-
-# Using single imputed data set
-fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
-
-# gvif
-gvif <- lapply(fit1, vif)
-
-lapply(gvif, \(g) {
-  g[g[,3] > 2, 3]
-})
-```
-
-The adjusted GVIF are acceptably low.
-
-Coefficients overview.
-
-```{r}
-# find coef of variable other than country.code
-var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","fdi","lbr.part","unemp","gdp.dflt","gcf","lfdi","lgdp.dflt")
-
-res <- do.call(rbind, lapply(fit1, \(f) f$coefficients[var]))
-kable(res)
-```
-
-There are several sign-switching in `income` coefficients, indicating the effect of removing some multicollinearity.
-
-```{r}
-fit1.summ <- lapply(fit1, \(f) {
-  k <- length(f$coefficients) - 1
-  pred <- predict(f, countries.test.4)
-  data.frame(
-    `Train Adj.R2` = summary(f)$adj.r.squared,
-    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
-  )
-})
-
-res <- do.call(rbind.data.frame, fit1.summ)
-res <- cbind(data.frame(Set = 1:3), res)
-kable(res)
-```
-
-Directly removing variables seems to be penalising our test results.
-As our interest is to investigate the effect of certain variables on poverty, the yielded R-Squared is within acceptable range.
-We can continue variable selection on the new models.
-
-```{r}
-formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+lfdi+lgdp.dflt'
+The models seem to be over-fitting. We should perform model simplification.
+
+##### B. Regularization
+
+Prepare data
+
+```{r}
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
+
+  return(sd(vct, na.rm = T))
+}
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+
+  return(vct/scl)
+}
+
+standardise <- function(train, test) {
+  data <- rbind(train, test)
+  scaler <- unlist(lapply(data, sd0))
+
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}
+```
+
+```{r}
+optimalModel <- function(train) {
+  #standardized
+}
 ```
@@
 \\
+We have no idea how to pool coefficients for AIC model,
 
@@
 fitAIC <- function(i) {
-  set.seed(1984)
   fit <- lm(as.formula(formula.str), complete(countries1.imputed, i))
@@
 aic.fits <- lapply(1:countries1.imputed$m, fitAIC)
-
-# check final terms of each model
-lapply(aic.fits, \(mod) formula(mod$terms))
-```
-
-The most commonly removed variables are:
-
-- `gdp.dflt`
-- `ldgp.dflt`
-- `unemp`
-- `fdi`
-
-```{r}
 # predict with each model
@@
 
-We can compare Adjusted R-Squared to estimate relative over-fitting in the new models.
+Adjusted R-Squared
 
@@
 
-Performance on train set and test set are not affected. However, we have simplified the model by removing some unnecessary terms. We can update our models.
-
-```{r}
-formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+lbr.part+gcf+lfdi'
-
-fit3 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
-
-# Build models with all generated data set and pool the estimates
-fit4 <- with(data = countries1.imputed, exp = lm(as.formula(formula.str)))
-fit4.combined <- pool(fit4)
-# dummy lm model
-fit4.dummy <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
-# replace coefficients of dummy model & predict
-fit4.dummy$coefficients <- fit4.combined$pooled$estimate
-
-fit3.summ <- lapply(fit3, \(f) {
-  k <- length(f$coefficients) - 1
-  pred <- predict(f, countries.test.4)
-  test.resid <- pred - countries.test.4$pov
-  data.frame(
-    `Train MSE` = mean(summary(f)$residuals^2),
-    `Test MSE` = mean(test.resid^2),
-    `Train MAE` = mean(abs(summary(f)$residuals)),
-    `Test MAE` = mean(abs(test.resid)),
-    `Train Adj.R2` = summary(f)$adj.r.squared,
-    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
-  )
-})
-
-k <- length(fit4.dummy$coefficients) - 1
-fit4.pred <- predict(fit4.dummy, countries.test.4)
-test.resid <- fit4.pred - countries.test.4$pov
-fit4.summ <- data.frame(
-  `Train MSE` = mean(summary(fit4.dummy)$residuals^2),
-  `Test MSE` = mean(test.resid^2),
-  `Train MAE` = mean(abs(summary(fit4.dummy)$residuals)),
-  `Test MAE` = mean(abs(test.resid)),
-  `Train Adj.R2` = pool.r.squared(fit4, adjusted = T)[,'est'],
-  `Test Adj.R2` = adjR2(fit4.pred, countries.test.4$pov, k)
-)
-
-res <- do.call(rbind.data.frame, fit3.summ)
-res <- rbind(res, fit4.summ)
-res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)
-
-kable(res)
-```
-
-Coefficients overview.
-
-```{r}
-var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","lbr.part","gcf","lfdi")
-
-res <- do.call(rbind, lapply(fit3, \(f) f$coefficients[var]))
-kable(res)
-```
-
-We can further treat over-fitting with regularization.
-
-##### C. Regularization
-
-Helper functions to standardise data.
-
-```{r}
-sd0 <- function(vct) {
-  if (!is.numeric(vct)) {
-    return(NA)
-  }
-
-  return(sd(vct, na.rm = T))
-}
-
-std0 <- function(vct, scl) {
-  if (!is.numeric(vct)) {
-    return(vct)
-  }
-
-  return(vct/scl)
-}
-
-standardise <- function(train, test) {
-  # save number of train and combine both data set
-  trainCases <- 1:nrow(train)
-  data <- rbind(train, test)
-  # calc scaler
-  scaler <- unlist(lapply(data, sd0))
-  # get the numeric columns
-  numCols <- which(unlist(lapply(data, is.numeric)))
-  # divide numeric columns by scalers, and combine with factor columns
-  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
-  fct <- data[,-numCols]
-  final <- cbind(fct, num)
-  # split to train and test
-  trainSet <- final[trainCases,]
-  testSet <- final[-trainCases,]
-
-  res <- list(final = final, train = trainSet, test = testSet, scaler = scaler)
-  return(res)
-}
-```
-
-Helper function to build the optimal model.
-
-```{r}
-optimalModel <- function(formula, train, test) {
-  set.seed(1984)
-  # standise and save number of train cases (to split later)
-  std.data <- standardise(train, test)
-  trainCases <- 1:nrow(train)
-  # matrix-ise
-  xs <- model.matrix(formula, std.data$final)[,-1]
-  ys <- std.data$final$pov
-  # split data
-  train.x <- xs[trainCases,]
-  train.y <- ys[trainCases]
-  test.x <- xs[-trainCases,]
-  test.y <- ys[-trainCases]
-  # list of tested alpha, resolution = 0.1
-  alphas <- seq(0, 1, 0.1)
-  # build a bunch of models
-  models <- lapply(alphas, \(a) cv.glmnet(train.x, train.y, type.measure = "mse", alpha = a))
-  cv.error <- unlist(lapply(models, \(model) model$cvm[model$lambda == model$lambda.min]))
-  # best model
-  best.model.idx <- which.min(cv.error)
-  # optimal alpha and lambda
-  alpha.opt <- alphas[best.model.idx]
-  lambda.opt <- models[[best.model.idx]]$lambda.min
-  best.model <- glmnet(train.x, train.y, alpha = alpha.opt, lambda = lambda.opt)
-  # predict for test data
-  train.fit <- predict(best.model, train.x)
-  test.fit <- predict(best.model, test.x)
-  # evaluation - train
-  k <- best.model$df
-  train.r2 <- best.model$dev.ratio
-  train.adjR2 <- adjR2(train.fit, train.y, k)
-  train.resid <- train.fit - train.y
-  train.mse <- mean(train.resid^2)
-  train.rmse <- sqrt(train.mse)
-  train.mae <- mean(abs(train.resid))
-  train.mape <- mean(abs(train.resid / train.y))
-  # evaluation - test
-  test.r2 <- r2(test.fit, test.y)
-  test.adjR2 <- adjR2(test.fit, test.y, k)
-  test.resid <- test.fit - test.y
-  test.mse <- mean(test.resid^2)
-  test.rmse <- sqrt(test.mse)
-  test.mae <- mean(abs(test.resid))
-  test.mape <- mean(abs(test.resid / test.y))
-  # final results
-  results <- list(
-    train = list(
-      data = train,
-      r2 = train.r2,
-      adj.r2 = train.adjR2,
-      residuals = train.resid,
-      mse  = train.mse,
-      rmse = train.rmse,
-      mae  = train.mae,
-      mape = train.mape
-    ),
-    test = list(
-      data = test,
-      r2 = test.r2,
-      adj.r2 = test.adjR2,
-      residuals = test.resid,
-      mse  = test.mse,
-      rmse = test.rmse,
-      mae  = test.mae,
-      mape = test.mape
-    ),
-    scaler = std.data$scaler,
-    alpha = alpha.opt,
-    lambda = lambda.opt,
-    model = best.model
-  )
-
-  return(results)
-}
-```
-
-Build the optimal models using each imputated data set.
-
-```{r}
-# helper function to build from a specific imputated data
-buildWith <- function(set, imputed, test, formula, fun) {
-  train <- complete(imputed, set)
-  model <- fun(formula, train, test)
-  return(model)
-}
-
-set.seed(1984)
-
-models <- lapply(1:countries1.imputed$m, buildWith, imputed = countries1.imputed, test = countries.test.4, formula = as.formula(formula.str), fun = optimalModel)
-```
-
-Display parameters and evaluation.
-
-```{r}
-result.list <- lapply(models, \(mod) {
-  data.frame(
-    alpha = mod$alpha,
-    lambda = mod$lambda,
-    `Train MSE` = mod$train$mse,
-    `Test MSE` = mod$test$mse,
-    `Train MAE` = mod$train$mae,
-    `Test MAE` = mod$test$mae,
-    `Train R2` = mod$train$r2,
-    `Test R2` = mod$test$r2,
-    `Train Adj.R2` = mod$train$adj.r2,
-    `Test Adj.R2` = mod$test$adj.r2
-  )
-})
-
-result.df <- do.call(rbind.data.frame, result.list)
-result.df <- cbind(data.frame(Set = 1:countries1.imputed$m), result.df)
-kable(result.df)
-```
-
-There are some improvement on the test data performance based on adjusted R-Squared. MSE and MAE remain low on both data sets.
-
-Coefficients overview.
-
-```{r}
-var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","lbr.part","gcf","lfdi")
-
-res <- do.call(rbind, lapply(models, \(f) coefficients(f$model)[var,]))
-kable(res)
-```
-
-All non-country.code variables are non-zero. We can observe that the regularisation models just remove the effect of `country.code`, for countries that have similar "baseline".
-
-###  3.4.2. Assessment
-
-Now we are evaluating some performance metrics, checking assumptions, and remedy some potential problems.
-
-##### A. Homoscedasticity
-
-Plot the residuals ~ pov
-
-```{r}
-for (mod in models) {
-  resid <- mod$train$residuals
-  fitted <- mod$train$data$pov + mod$train$residuals
-  # plot(resid ~ pov, main = paste("Model", i))
-  print(ggplot(mapping = aes(y = resid, x = fitted)) +
-    geom_point() +
-    geom_smooth(formula = y ~ x, method = loess))
-}
-```
-
-There seems to be some larger variance with higher values of `pov`. We can conduct a statistical test to confirm the present of heteroscedasticity.
-
-The Goldfeld-Quandt test is performed by eliminating a certain number of observations from the dataset’s center, then comparing the spread of residuals between the two datasets on either side of the central observations.
-
-The Goldfeld-Quandt test examines two submodels’ variances divided by a defined breakpoint and rejects if the variances disagree.
-
-Under H0, the Goldfeld-Quandt test’s test statistic follows an F distribution with degrees of freedom as specified in the parameter.
-
-- **Null (H0):** Heteroscedasticity is not present.
-- **Alternative (H1):** Heteroscedasticity is present.
-
-```{r}
-lapply(models, \(m) {
-  resid <- m$train$residuals
-  # apply Goldfeld-Quandt test
-  gqtest(formula = resid ~ 1, order.by = m$train$data$pov)
-})
-```
-
-There is heteroscedasticity in our models.
-[Generalized Least Squares With Unknown For of Variance](https://rpubs.com/cyobero/187387) is a possible remedy of this problem, but at the cost of model interpretation.
-The high variance in higher end of `pov` can be explained by the relative volatile economics and politic climate in highly poor countries ($pov > 0.25$), leading to unstable effect of predictors.
-
-##### B. Independence
-
-Some possible source of dependency are poverty measured on the same country, or in the same year. These are accounted in our models by controlling those variables.
-
-Some countries have been engaging in international wars, being under the influences of foreign forces such as Iraq, and Afghanistan.
-They possess similar (and different) political climates.
-Some other groups of economical alliance, or enjoying similar natural resources: OPEC, EU, ASEAN, etc.
-might have similar characteristics that make them dependent.
-
-A solution to resolve this is to conduct the study on countries with similar geographical characteristics.
-In our study, however, we are interested in the general effects of certain variable across the globe.
-Although there are some dependence, the large number of entities with various features, we hypothesize, will cancel each other out, resulting in a net effect that gives an overview of the influences of variables.
-
-##### C. Normally Distributed Residuals
-
-```{r}
-for (m in models) {
-  p <- ggplot(mapping = aes(x = m$train$residuals)) +
-    geom_histogram(aes(y = ..density..), bins = 30) +
-    geom_density(alpha = .2, fill = "blue")
-
-  print(p)
-}
-```
-
-Although quite negatively skewed, the residuals are visibly normal.
-
-##### D. Outliers
-
-As elastic net model produced in
-
-###  3.4.3. Interpretation
+-> AIC on one model [link](https://rforpoliticalscience.com/2020/10/23/choose-model-variables-by-aic-in-a-stepwise-algorithm-with-the-mass-package-in-r/)
+
+-> stepwise selection on mice  [link](https://stefvanbuuren.name/fimd/sec-stepwise.html)
+
+#### 3.2.4.2. Assessment
+Check assumption
+outliers
+#### 3.2.4.3. Interpretation
+
+## 3.3. Panel Data Analysis (To be done)
 

### 2022-11-17 src/draft.Rmd

--- 2022-11-16 src/draft.Rmd
@@
 
-## 3.1. Use Complete Cases (Imtiyaaz
-###  3.1.1. Model Fitting
-###  3.1.2. Assessment
-###  3.1.3. Interpretation
+## 3.1. Use Complete Cases (Imtiyaaz)
+```{r}
+library(tidyverse) # for data manipulations and visualisations
+library(ggpubr) # for correlation coefficient on plot
+library(gridExtra) # for arranging plots 
+library(e1071) # for skewness
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(corrplot)
+library(car)
+library(DescTools)
+library(glmnet)
+```
+
+First we will write the function "eval_results(true,predict) to evaluate the models
+```{r}
+eval_results <- function(true, predict) {
+  actual <- data.matrix(true)
+  SSE <- sum((predict - actual)^2)
+  SST <- sum((actual - mean(actual))^2)
+  R_square <- 1 - SSE / SST
+  data.frame(
+    MSE = MSE(predict, true),
+    MAE = MAE(predict, true),
+    RMSE = RMSE(predict, true),
+    Rsquare = R_square
+  )
+}
+```
+First we read the csv file for the countries containing the dataset. We will then put the variables we need in to subset1.
+```{r}
+df <-read.csv('../data/countries1.csv')
+df
+subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
+subset1
+```
+
+Next we will build a SLR model. We will put 80% of dataset to train and the rest to test.
+```{r}
+set.seed(10)
+index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
+train <- df[index, ]
+test <- df[-index, ]
+```
+
+The model output will be placed in lm.slr and can be viewed in summary(lm.slr). THe r-squared value is 0.9619.
+```{r}
+lm.slr <- lm(pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
+summary(lm.slr)
+
+```
+Next we will be performing standardisation.However not all the variables are numerical.
+Therefore to overcome it we build the function below.
+This will allow us to perform standardisation to the dataset.
+```{r}
+subset1$income <- factor(subset1$income)
+subset1$country.code <- factor(subset1$country.code)
+
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
+
+  return(sd(vct, na.rm = T))
+}
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+
+  return(vct/scl)
+}
+
+scale <- function(data) {
+  unlist(lapply(data, sd0))
+}
+
+standardise <- function(data) {
+  scaler <- scale(data)
+
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}
+```
+
+Next we will perform standardistation to the dataset.
+```{r}
+str(subset1)
+df0 <- subset1
+subset1
+apply(subset1, 2, sd)
+
+```
+
+Next we will plot a pov vs income scatter plot to see the effect of income on pov.
+From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov.
+This is consistent with our intuition that countries that are richer do not have as much poverty issue as the poorer countries.
+```{r}
+ggplot(data = subset1, aes(x = income, y = pov)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  labs(title = "Pov vs. Income ",
+       x = "income",
+       y = "pov")
+
+```
+Next we will build a MLR model with data from subset 1 to see the relationship between dependent variable:pov and the other variables
+
+```{r}
+model1 <- lm(pov ~ ., data = subset1)
+summary(model1)
+vif(model1)
+
+```
+Next we will perform regularisation. We will split the data to predictor variable "train.x" and response variables "train.y"
+```{r}
+subset2 <- subset(subset1,select =c("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
+subset2 <- na.omit(subset2)
+subset2
+train.x <- data.matrix(subset(subset2, select = -pov))
+train.y = data.matrix(subset2$pov)
+```
+
+Next we will perform splitting of data to train and test.
+```{r}
+split_train_test <- function(data) {
+set.seed(1984)
+# split data
+seeds <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() == 1)
+
+plants <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() != 1)
+
+isComplete <- which(complete.cases(plants))
+idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
+
+training <- plants[-idx,] %>% 
+  rbind(seeds)
+test <- plants[idx,]
+
+return(list(training, test))
+}
+
+training <- split_train_test(subset2)[[1]]
+test <- split_train_test(subset2)[[2]]
+
+ntrain <- 1:nrow(training)
+combine <- rbind(training, test)
+combine.mtrx <- model.matrix(pov ~ ., data = combine)
+train.mtrx <- combine.mtrx[ntrain,]
+test.mtrx <- combine.mtrx[-ntrain,]
+
+train.x <- train.mtrx[,-1]
+train.y <- training$pov
+
+test.x <- test.mtrx[,-1]
+text.y <- test$pov
+```
+
+We then build a MLR model with the training model to train it.
+
+```{r}
+model3 <- lm(pov ~., data = training)
+summary(model3)
+
+```
+
+```{r}
+fit <- model3$fitted.values
+true <- training$pov
+summary_MLR3_train <- eval_results(true, fit)
+```
+
+We then evaluate the model between train and test
+```{r}
+fit <- predict(model3, newdata=test)
+true <- test$pov
+summary_MLR3_test <- eval_results(true, fit)
+```
+
+We can see the evaluation of the model from the function below. From the train model, the R^squared value will be 0.965 and for the test model it will be 0.881.
+
+```{r}
+summary_MLR <- rbind(summary_MLR3_train[-5], 
+                     summary_MLR3_test[-5])
+rownames(summary_MLR) <- c("Baseline MLR_Train", 
+                           "Baseline MLR_Test" )
+knitr::kable(summary_MLR, digits = 3)
+```
+Next we will be performing ridge and lasso regression model using the following code. We will split the data to train and test.
+```{r}
+# Training Data
+data <- training
+train.x <- data.matrix(data[,-ncol(data)])
+train.y <- data.matrix(data[,ncol(data)])
+
+# Test Data
+data <- test
+test.x <- data.matrix(data[,-ncol(data)])
+test.y <- data.matrix(data[,ncol(data)])
+```
+
+Next we will perform the Ridge regression model.
+```{r}
+# Fit Ridge Regression Model
+set.seed(123)
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x, train.y, 
+                    alpha = 0, lambda = cv_ridge$lambda.min)
+
+# Evaluate Ridge Regression Model on Training Set
+fit <- predict(glm_ridge, newx = train.x, type = "response")
+true <- train.y
+summary_ridge_train <- eval_results(true, fit)
+
+# Evaluate Ridge Regression Model on Test Set
+fit <- predict(glm_ridge, newx = test.x, type = "response")
+true <- test.y
+summary_ridge_test <- eval_results(true, fit)
+
+# Summarise Ridge Regression Model Performance into Table
+summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
+rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
+```
+
+Next we will perform Lasso Regression model
+
+```{r}
+# Fit LASSO Model
+set.seed(123)
+cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
+glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
+                    lambda = cv_lasso$lambda.min)
+
+# Evaluate LASSO Model on Training Set
+fit <- predict(glm_lasso, newx = train.x)
+true <- train.y
+summary_lasso_train <- eval_results(true, fit)
+
+# Evaluate LASSO Model on Test Set
+fit <- predict(glm_lasso, newx = test.x)
+true <- test.y
+summary_lasso_test <- eval_results(true, fit)
+
+# Summarise LASSO Model Performance into Table
+summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
+rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
+```
+
+We will compare the 3 models that we have build, Multi linear ,Ridge and Lasso regression models.
+We evaluate the 3 models together and compare both train and test models of each of the 3 regression model.
+
+```{r}
+summary_3models <- rbind(summary_MLR3_train, 
+                         summary_ridge_train,                               
+                         summary_lasso_train, 
+                         summary_MLR3_test,                             
+                         summary_ridge_test, 
+                         summary_lasso_test)
+
+rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
+                               "LASSO Model_Train",  "Baseline MLR_Test",  
+                               "Ridge Model_Test", "LASSO Model_Test" )
+
+knitr::kable(summary_3models[1:3,-5], digits = 3)
+knitr::kable(summary_3models[4:6,-5], digits = 3)
+
+```
+From the 3 models we can see that the original multi linear regression model is the best one with the highest R^squared value.
+However the Lasso Regression Model has the lowest MSE, MAE and RMSE.
+We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting as many values will get removed.
+Next we will build other models to see if they work better.
 
@@
 gdp per capital, income class and region of the country can provide a model to predict the poverty headcount ratio in a country.
+The error rate of the model is high on the testing data set, showing it may not be a good model.
 

### 2022-11-17 src/draft.Rmd
Update draft.Rmd

@@
 ## 3.2. Selectively remove variables with high missing rate (Callista)
+---
+title: "Selectively Remove Variables"
+author: "Callista Stephine Yu"
+date: "2022-11-07"
+output: html_document
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r}
+library(knitr)
+library(readr)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(stats)
+library(tidyverse)
+library(e1071)
+library(car)
+library(glmnet)
+library(corrplot)
+```
+
+```{r}
+countries <- read.csv("../data/countries.csv")
+```
+
 ###  3.2.1. Model Fitting
+##### A. Ordinary Linear Regression
+Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables.
+This section attempts to build a model on variables that have missing rates of below 35%.
+
+The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models.
+Hence, we transformed the data with the mutate() function before building the machine learning model.
+```{r}
+
+subset0 <- subset(countries, select = c("country.code", "year","edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
+
+subset1_0 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
+
+```
+subset1 contains the variables below 35% of missing values. 
+```{r}
+subset1 <- subset(subset1_0, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
+
+subset1 <- na.omit(subset1)
+```
+We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step. 
+```{r}
+sd0 <- function(vct) {
+  if (!is.numeric(vct)) {
+    return(NA)
+  }
+
+  return(sd(vct, na.rm = T))
+}
+
+std0 <- function(vct, scl) {
+  if (!is.numeric(vct)) {
+    return(vct)
+  }
+
+  return(vct/scl)
+}
+
+scale <- function(data) {
+  unlist(lapply(data, sd0))
+}
+
+standardise <- function(data) {
+  scaler <- scale(data)
+
+  numCols <- which(unlist(lapply(data, is.numeric)))
+  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
+  fct <- data[,-numCols]
+  return(cbind(fct, num))
+}
+
+```
+
+```{r}
+
+df.countries0 <- standardise(subset1)
+df.countries <- df.countries0
+```
+
+```{r}
+#Divide each feature/target by its standard deviation
+
+head(df.countries)
+```
+
+First, we split the data into training and testing data. 
+```{r}
+set.seed(123)
+index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
+train_1 <- subset1[index,]
+test_1 <- subset1[-index,]
+
+```
+
+Our first attempt 
+```{r}
+#Everything
+lm_1 <- lm(data= train_1, formula = pov ~ .)
+summary(lm_1)
+```
+
+The summary of `lm_1` suggests that the model still contains predictor variables with pr(>|t|) > 0.05.
+We can imply that these variables do not provide a high significance to our model and are thus able to be removed.
+To avoid removing significant variables, we shall conduct the removal of the variables deliberately.
+In `lm_1`, we notice that the variable `lgdp.dflt` has a high Pr(>|t|).
+Hence, we remove `lgdp.dflt`.
+
+```{r}
+#remove hlth 
+lm_2 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
+summary(lm_2)
+plot(lm_2, which = 5)
+```
+
+```{r}
+#Remove lgdp.dflt
+lm_3 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
+summary(lm_3)
+plot(lm_3, which = 5)
+```
+```{r}
+#remove lfdi
+lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
+summary(lm_4)
+plot(lm_4, which = 5)
+```
+Our model shows that among all the 13 predictor variables that we started with, only 11 of them showed significance in building the model.
+
+We then evaluated the results of the model. 
+```{r}
+#evaluating lm_4
+eval.metrics.linreg <- function(actual, predicted) {
+  residual <- actual - predicted
+  mse <- mean(residual ^ 2)
+  mae <- mean(abs(residual))
+  rmse <-  sqrt(mse)
+  mape <- mean(abs(residual / actual)) * 100
+
+  data.frame(
+    MSE = mse,
+    MAE = mae,
+    RMSE = rmse,
+    MAPE = mape
+  )
+}
+```
+
+```{r}
+actual <- train_1$pov
+predicted <- predict(lm_4, newdata = train_1)
+eval.metrics.linreg(actual, predicted)
+```
+The MAPE shows `Inf` due to the presence of 0 values in the Y variable.
+
+##### B. Regularization
+Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model.
+
+Next, we will perform Ridge regression and LASSO regression in a dataset that contains `country.code` and `year` in addition to the previous dataset that we used for the previous linear regression model.
+```{r}
+#Subsetting the data
+
+data0 <- subset(subset1_0, select = c("country.code","year","edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
+data <- na.omit(data0)
+train.x <- data.matrix(subset(data, select = -pov))
+train.y <- data.matrix(data$pov)
+
+```
+
+Splitting the aforementioned dataset into training and testing data.
+
+```{r}
+split_train_test <- function(data) {
+set.seed(1984)
+# split data
+seeds <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() == 1)
+
+plants <- data %>% 
+  group_by(country.code) %>% 
+  filter(row_number() != 1)
+
+isComplete <- which(complete.cases(plants))
+idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
+
+training <- plants[-idx,] %>% 
+  rbind(seeds)
+test <- plants[idx,]
+
+return(list(training, test))
+}
+```
+
+```{r}
+training <- split_train_test(data)[[1]]
+test <- split_train_test(data)[[2]]
+```
+Further splitting the training and testing data into  `train.x` and `train.y`
+```{r}
+ntrain <- 1:nrow(training)
+combine <- rbind(training, test)
+combine.mtrx <- model.matrix(pov ~ ., data = combine)
+train.mtrx <- combine.mtrx[ntrain,]
+test.mtrx <- combine.mtrx[-ntrain,]
+
+train.x <- train.mtrx[,-1]
+train.y <- training$pov
+
+test.x <- test.mtrx[,-1]
+text.y <- test$pov
+```
+
+Building a model with the lambda that outputs the smallest MSE. 
+```{r}
+cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
+glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
+
+t(coef(glm_ridge))
+```
+The ridge regression model shows that some countries have a higher influence on the model than others.
+At first glance,it seems like the Y variable, `pov`, depends heavily on whether or not it belongs to certain countries.To further illustrate the significance of `country.code`, we will plot the ridge regression graph for different lambda values.
+```{r}
+#generate sequence of lambda values 
+lambda <- 10^seq(-6, 2, length = 100)
+ridge_model = glmnet(train.x, train.y, alpha = 0, lambda = lambda)
+
+t(coef(ridge_model))
+```
+Plotting the coefficients against different lambda values.
+
+```{r}
+plot_glmnet(ridge_model)
+```
+
+Interesting! For a low lambda value, the model exhibits `cntr.MDG`, which represents Madagascar, followed by `cntr.ZMB`, which represents Zimbabwe,  as key predictor variables.
+This means that the model depends highly on whether the country is Madagascar or Zimbabwe, followed by some other significant countries.
+The other variables that previously defined the model is now undermined by the presence of the `country.code` variable and does not show high influence.
+
+We then proceeded to evaluate the results
+
+```{r}
+eval_results <- function(fit, true) {
+  actual <- data.matrix(true)
+  SSE <- sum((actual - fit)^2)
+  SST <- sum((actual - mean(actual))^2)
+  R_square <- 1 - SSE/SST
+  data.frame(
+    MSE = MSE(fit, true),
+    MAE = MAE(fit, true),
+    RMSE = RMSE(fit, true),
+    MAPE = MAPE(fit, true),
+    R2  = R_square
+
+  )
+}
+```
+
+```{r}
+fit <- predict(glm_Ridge, train.x)
+true <- train.y
+summary_Ridge_train <- eval_results(fit, true)
+summary_Ridge_train
+```
+The training dataset produced a much higher value compared to the regular multiple regression model shown above.
+
+Fitting the model to the test dataset,
+
+```{r}
+fit <- predict(glm_Ridge, test.x)
+true <- test.y
+summary_Ridge_test <- eval_results(fit, true)
+summary_Ridge_test
+```
+A slightly lower R^2 value, but still proves that the model is good enough.
+
+Below is the summary of the results of the training and testing dataset. 
+```{r}
+summary_Ridge <- rbind(summary_Ridge_train, summary_Ridge_test)
+rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
+knitr::kable(summary_Ridge, digits = 3)
+
+```
+Next, we attempted LASSO regression as well. The steps are essentially similar to that of Ridge regression. 
+```{r}
+#Train a LASSO regression model
+set.seed(123)
+cv_lasso <- cv.glmnet(train.x, train.y, alpha = 1, type.measure = "mse")
+```
+
+```{r}
+cv_lasso
+```
+
+```{r}
+plot(cv_lasso)
+```
+
+```{r}
+glm_Lasso <- glmnet(train.x, train.y, alpha = 1, lambda = cv_lasso$lambda.min)
+t(coef(glm_Lasso))
+```
+It is evident that the number of variables included in the model is less than that of ridge regression.
+This is in accordance with the nature of the LASSO regression model that reduces the coefficients of certain predictor variables to zero for model interpretability.
+
+```{r}
+lambda <- 10^seq(-6, 2, length = 100)
+lasso_model = glmnet(train.x, train.y, alpha = 1, lambda = lambda)
+
+t(coef(lasso_model))
+```
+```{r}
+plot_glmnet(lasso_model)
+```
+```{r}
+plot_glmnet(ridge_model)
+```
+
+```{r}
+fit2 <- predict(glm_Lasso, train.x)
+true2 <- train.y
+summary_Lasso_train <- eval_results(fit2, true2)
+summary_Lasso_train
+```
+
+```{r}
+plot_glmnet(lasso_model)
+```
+
+```{r}
+fit3 <- predict(glm_Lasso, test.x)
+true3 <- test.y
+summary_Lasso_test <- eval_results(fit3, true3)
+summary_Lasso_test
+```
+```{r}
+summary_lasso <- rbind(summary_Lasso_train, summary_Lasso_test)
+rownames(summary_lasso) <- c("LASSO_train", "LASSO_test")
+knitr::kable(summary_lasso, digits = 3)
+
+```
 ###  3.2.2. Assessment
-###  3.2.3. Interpretation
+
+This section recapitulates the three models that we have generated so far through the method of gradually removing variables from the original dataset.
+
+```{r}
+summary(lm_4)
+```
+
+Our first model, `lm_4`, generated with regular multiple linear regression exhibits the lowest R^2 value.
+The model shows that `incomeL` is the most significant predictor variable with the highest coefficient.
+One interesting observation is that `incomeLM` and `incomeUM` has a negative correlation with the pov level, in contrast to `incomeL`.
+This shows that countries that have lower income tend to have high `pov` value.
+On the other hand, countries that are classified to have lower middle income and upper middle income seem to have lower `pov` values, as shown by their negative coefficients in the regression model.
+
+Apart from income, some other variables that contribute more significantly to the model are `lgdp.pc`, with a coefficient of -3.13114, exhibiting a stronger negative correlation, and `pop.gwth.urban`, exhibiting a stronger positive correlation with a coefficient of 1.90680.
+This is counter-intuitive, as the notion is that population growth in cities will help drive positive economic change due to the number of productive people living in the city.
+The model also shows that `pop.gwth.rural` gives a less significant impact on `pov` as compared to  `pop.gwth.rural`.
+
+```{r}
+knitr::kable(summary_lasso, digits = 3)
+```
+
+```{r}
+knitr::kable(summary_Ridge, digits = 3)
+```
+The MSE for the LASSO regression model is slightly higher than that of the Ridge regression model.
+This to be expected, as the Ridge regression model aims for an improved accuracy while the LASSO model aims for an improved interpretability.
+Both models seem to exhibit very high R^2 value for both the training and testing dataset.
+
+```{r}
+plot_glmnet(lasso_model)
+```
+```{r}
+
+```
+
+```{r}
+plot_glmnet(ridge_model)
+```
+```{r}
+plot(lm_4, which = 5)
+```
+The `Residual vs Leverage` plot of `lm_4` suggests that heteroskedasticity might be present in the model.
+This means that the variance of the residuals is unequal over a range of observed values.
+This problem exists in the dataset rather than the models generated.
 

### 2022-11-17 src/Group 1_Project Code.Rmd
zmb is zambia

--- 2022-11-17 src/draft.Rmd
@@
 ---
-title: "Draft Documents"
-author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
+title: "Public Expenditure and Poverty"
+author: "Team 1: Le Van Minh, Callista Stephine Yu, Muhammad Imtiyaaz B M A M , Thum Jun Long Issac"
 date: "`r Sys.Date()`"
@@
 # code chunk out put to 
-options(digits = 5)
-```
-
-```{r, message = F}
-library(needs)
-
-needs("dplyr", "MASS", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce", "lmtest")
-
-prioritize(dplyr)
-```
-
-# 1. Introduction (To be done)
+options(digits = 3)
+```
+
+# 1. Introduction
 The project aim is to explore different models using the same dataset to evaluate which method will be the best to build a model for the dependent variable.
@@
 
+Github repo: [https://github.com/IntrovertHedgehog/PublicExpenditureAndPoverty](https://github.com/IntrovertHedgehog/PublicExpenditureAndPoverty)
+
 # 2. Data Characteristic
+
+```{r, message = F}
+library(needs)
+
+needs("dplyr", "MASS", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce", "lmtest")
+
+prioritize(dplyr)
+```
 
@@
 22. **gcf**: Gross Capital Formation\
-23. **trade**: Trade = import + export (% of GDP)
+23. **trade**: Trade = import + export (% of GDP)\
 24. **gdp.pc**: GDP per capita (current US$)
@@
 
-1.  **Use complete cases**: Only use the complete cases for the analysis. This is a straightforward approach, but doesn't resolve the bias resulted from the mass loss of data.
-2.  **Selectively remove variables with high missing rate**: The same as we did before, but this process should be carried out carefully as we run the chance of dropping an important variable.
-3.
-**Update the data set as we select variables**: As we drop insignificant variables (in backward selection), the number of NAs are changed as well.
-We can utilize the extra complete cases to build the next model in the steps.
+1.  **Use complete cases**: Only use the complete cases for the analysis. This is a straightforward approach, but doesn't resolve the bias resulted from the massive loss of data.
+2.
+**Selectively remove variables with high missing rate**: Start with all the variable in the data set, then step-by-step remove insignificant, or frequently missing variables to use as much data as possible.
+3.  **Forward variable selection**: Start with 0 variable, then selectively add potentially important variable to the data set to achieve optimal model.
 4.
 **Imputation**: The idea is to replace the missing observations on the response or the predictors with artificial values that try to preserve the data set structure.
-This is a quite complex topic of its own, but we think why not.
+This is a quite complex topic of its own.
 You can read more from @imputation.
@@
 
-## 3.1. Use Complete Cases (Imtiyaaz)
-```{r}
-library(tidyverse) # for data manipulations and visualisations
-library(ggpubr) # for correlation coefficient on plot
-library(gridExtra) # for arranging plots 
-library(e1071) # for skewness
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-library(readxl)
-library(corrplot)
-library(car)
-library(DescTools)
-library(glmnet)
+## 3.1. Use Complete Cases
+```{r}
+needs("tidyverse","ggpubr","gridExtra","e1071","ggplot2","dplyr","tidyr","readxl","corrplot","car","DescTools","glmnet")
 ```
@@
 df <-read.csv('../data/countries1.csv')
-df
+head(df)
 subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
-subset1
+head(subset1)
 ```
@@
 
-The model output will be placed in lm.slr and can be viewed in summary(lm.slr). THe r-squared value is 0.9619.
+The model output will be placed in lm.slr and can be viewed in summary(lm.slr).
 ```{r}
@@
 df0 <- subset1
-subset1
+head(subset1)
 apply(subset1, 2, sd)
-
 ```
@@
 vif(model1)
-
 ```
@@
 subset2 <- na.omit(subset2)
-subset2
+head(subset2)
 train.x <- data.matrix(subset(subset2, select = -pov))
@@
 test.x <- test.mtrx[,-1]
-text.y <- test$pov
+test.y <- test$pov
 ```
@@
 summary(model3)
-
 ```
@@
 knitr::kable(summary_3models[4:6,-5], digits = 3)
-
 ```
@@
 
-## 3.2. Selectively remove variables with high missing rate (Callista)
----
-title: "Selectively Remove Variables"
-author: "Callista Stephine Yu"
-date: "2022-11-07"
-output: html_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-```{r}
-library(knitr)
-library(readr)
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(stats)
-library(tidyverse)
-library(e1071)
-library(car)
-library(glmnet)
-library(corrplot)
+## 3.2. Selectively remove variables with high missing rate
+
+```{r}
+needs("knitr","readr","tidyr","dplyr","ggplot2","stats","tidyverse","e1071","car","glmnet","corrplot","plotmo")
 ```
@@
 ###  3.2.1. Model Fitting
-##### A. Ordinary Linear Regression
+#### A. Ordinary Linear Regression
 Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables.
@@
 ```{r}
-
 subset0 <- subset(countries, select = c("country.code", "year","edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
@@
 subset1_0 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
-
 ```
@@
   fct <- data[,-numCols]
-  return(cbind(fct, num))
-}
-
-```
-
-```{r}
-
-df.countries0 <- standardise(subset1)
+  return(list(cbind(fct, num), scaler))
+}
+
+```
+
+```{r}
+df.countries0 <- standardise(subset1)[[1]]
+scaler2 <- standardise(subset1)[[2]]
 df.countries <- df.countries0
@@
 #Divide each feature/target by its standard deviation
-
 head(df.countries)
@@
 test_1 <- subset1[-index,]
-
 ```
@@
 ```
-Our model shows that among all the 13 predictor variables that we started with, only 11 of them showed significance in building the model.
+Our model shows that among all the 14 predictor variables that we started with, only 11 of them showed significance in building the model.
 
@@
 
-##### B. Regularization
+#### B. Regularization
 Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model.
@@
 train.y <- data.matrix(data$pov)
-
 ```
@@
 cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
-glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
-
-t(coef(glm_ridge))
+glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
+
+t(coef(glm_Ridge))
 ```
@@
 
-t(coef(ridge_model))
+head(t(coef(ridge_model)))
 ```
@@
 
-Interesting! For a low lambda value, the model exhibits `cntr.MDG`, which represents Madagascar, followed by `cntr.ZMB`, which represents Zimbabwe,  as key predictor variables.
-This means that the model depends highly on whether the country is Madagascar or Zimbabwe, followed by some other significant countries.
+Interesting! For a low lambda value, the model exhibits `cntr.MDG`, which represents Madagascar, followed by `cntr.ZMB`, which represents Zambia,  as key predictor variables.
+This means that the model depends highly on whether the country is Madagascar or Zambia, followed by some other significant countries.
 The other variables that previously defined the model is now undermined by the presence of the `country.code` variable and does not show high influence.
@@
     R2  = R_square
-
   )
@@
 knitr::kable(summary_Ridge, digits = 3)
-
 ```
@@
 ```
-```{r}
-
-```
 
@@
 
-## 3.3. Update the data set as we select variables (Isaac)
+## 3.3. Forward variable selection
 ###  3.3.1. Model Fitting 
@@
 countries.test.4 <- plants[idx,]
+
+meth <- c(rep("",4), "cart", "", rep("cart", 16))
+names(meth) <- colnames(countries1)
+meth
 
@@
 ```
-Adjusted R-Squared on train and test
+R-Squared on train and test
 ```{r}
@@
     `Test MAE` = mean(abs(test.resid)),
-    `Train Adj.R2` = summary(f)$adj.r.squared,
-    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
+    `Train R2` = summary(f)$r.squared,
+    `Test R2` = r2(pred, countries.test.4$pov)
   )
@@
   `Test MAE` = mean(abs(test.resid)),
-  `Train Adj.R2` = pool.r.squared(fit2, adjusted = T)[,'est'],
-  `Test Adj.R2` = adjR2(fit2.pred, countries.test.4$pov, k)
+  `Train R2` = pool.r.squared(fit2)[,'est'],
+  `Test R2` = r2(fit2.pred, countries.test.4$pov)
 )
@@
 
-Pooled model slightly improve performance. Test data fitting has better MSE and MAE than train data. However, Adjusted R-Squared is lower in test data. There might be some over-fitting in our models.
+Pooled model slightly improve performance. Test data fitting has better MSE and MAE than train data. However,  R-Squared is lower in test data. There might be some over-fitting in our models.
 
@@
 
-There're still some variable with adjusted GVIF > 2.
-We are interested in the effects of `hlth` (expenditure in Healthcare), and `mil` (expenditure in Military), so we won't remove those variables.
-We can remove `trade`.
+There're still some variable with adjusted GVIF > 2. We are interested in the effects of `hlth` (expenditure in Healthcare), so we won't remove those variables. We can remove `trade`.
 
@@
   data.frame(
-    `Train Adj.R2` = summary(f)$adj.r.squared,
-    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
+    `Train Adj.R2` = summary(f)$r.squared,
+    `Test Adj.R2` = r2(pred, countries.test.4$pov)
   )
@@
 k <- lapply(aic.fits, \(m) length(m$coefficients))
-aic.test.r2 <- unlist(mapply(adjR2, aic.pred, k = k, MoreArgs = list(orig = countries.test.4$pov)))
+aic.test.r2 <- unlist(mapply(r2, aic.pred, MoreArgs = list(orig = countries.test.4$pov)))
 
@@
 
-We can compare Adjusted R-Squared to estimate relative over-fitting in the new models.
+We can compare R-Squared to estimate relative over-fitting in the new models.
 
@@
   `Test MAE` = mean(abs(test.resid)),
-  `Train Adj.R2` = pool.r.squared(fit4, adjusted = T)[,'est'],
-  `Test Adj.R2` = adjR2(fit4.pred, countries.test.4$pov, k)
+  `Train R2` = pool.r.squared(fit4)[,'est'],
+  `Test R2` = r2(fit4.pred, countries.test.4$pov)
 )
@@
     `Train R2` = mod$train$r2,
-    `Test R2` = mod$test$r2,
-    `Train Adj.R2` = mod$train$adj.r2,
-    `Test Adj.R2` = mod$test$adj.r2
+    `Test R2` = mod$test$r2
   )
@@
 
-Some obvious outliers are point 1122, 1327, 1395, 1396, 1424, 1474.
-
-```{r}
-out <- c(1122, 1327, 1395, 1396, 1424, 1474)
+Some obvious outliers are point 1112, 1329, 1395, 1396, 1424, 147, 1486, 1508.
+
+```{r}
+out <- c(1112, 1329, 1395, 1396, 1424, 147, 1486, 1508)
 var <- c("country.code","country.name","year","pov","income","edu.total","hlth","mil","lbr.part","gcf","lfdi")
@@
 # 4. Conclusion
-# 5. Appendix
-# 6. References
+
+Approach 1, 2 and 4 seem to have the best performace.
+
+First Approach:
+```{r}
+knitr::kable(summary_3models[1:3,-5], digits = 3)
+knitr::kable(summary_3models[4:6,-5], digits = 3)
+```
+
+Second Approach:
+
+```{r}
+knitr::kable(summary_lasso, digits = 3)
+knitr::kable(summary_Ridge, digits = 3)
+```
+
+Fourth Approach:
+```{r}
+kable(result.df[c('alpha', 'lambda', 'Train.R2', 'Test.R2')])
+```
+
+We can deduce our final interpretation from these models.
+
+```{r}
+var <- c('incomeL', 'incomeLM', 'incomeUM', 'edu.total', 'hlth', 'mil')
+result1 <- coefficients(model3)[var]
+# result1[names(scaler2)] <- result1[names(scaler2)] / scaler2[names(scaler2)]
+# result1 <- result1 * scaler2['pov']
+# result1 <- result1[var]
+
+result2 <- coefficients(glm_Ridge)[,'s0']
+result2[names(scaler2)] <- result2[names(scaler2)] / scaler2[names(scaler2)]
+result2 <- result2 * scaler2['pov']
+result2 <- result2[var]
+
+result4 <- res[2,var]
+
+# combine
+finalResults <- data.frame(Approach = c(1,2,4))
+finalResults <- finalResults %>% 
+  cbind(rbind(result1, result2, result4))
+
+kable(finalResults)
+```
+
+Some important interpretation:
+
+- Low income countries generally have higher poverty index than High income countries.
+- As suggested by 1st approach, Lower-Middle income countries have higher poverty index than High income countries. While 2nd and 4th approach suggest otherwise.
+- Upper-Middle income countries in general have lower poverty index than High income countries.
+
+As suggested by these models.
+Upper-Middle income countries have lowest poverty rate.
+This can be attributed to that these countries tend to have more even wealth distribution than high income countries, and have enough money to alleviate poor people.
+
+- Spending on education is negatively correlated with poverty.
+On average, 1% (of GDP) more spending in education results in 3 point decrease in poverty index.
+This suggests that public spending on education can reduce poverty.
+- Spending on Healthcare is negatively correlated with poverty. On average, 1% (of GDP) more spending in education results in 0.5 point decrease in poverty index.
+- Spending on military has ambiguous effect on poverty.
+Different model results in different interpretation.
+This can be accounted to the difference in the data used in each approach.
+Approach 1 only use complete cases, which caused the model to predict positive relationship.
+However, as more data is utilised in approach 2&4 use more data, the effect tend to be neutral/negative.
+
+# 5. References

### 2022-11-17 src/draft.Rmd
minor changes

@@
 ## 3.2. Selectively remove variables with high missing rate (Callista)
----
-title: "Selectively Remove Variables"
-author: "Callista Stephine Yu"
-date: "2022-11-07"
-output: html_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-```{r}
-library(knitr)
-library(readr)
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(stats)
-library(tidyverse)
-library(e1071)
-library(car)
-library(glmnet)
-library(corrplot)
+
+```{r}
+needs("knitr","readr","tidyr","dplyr","ggplot2","stats","tidyverse","e1071","car","glmnet","corrplot","plotmo")
 ```
@@
 # 4. Conclusion
-# 5. Appendix
-# 6. References
+# 5. References

### 2022-11-17 src/Selectively_remove_variables.Rmd

@@
 library(corrplot)
+library(plotmo)
 ```
@@
 
-```{r}
-head(countries)
-```
 ###  3.2.1. Model Fitting

### 2022-11-17 src/Selectively_remove_variables.Rmd
probably final result

@@
 ```{r}
-fit <- predict(glm_Ridge, train.x)
+fit <- predict(glm_ridge, train.x)
 true <- train.y
@@
 ```{r}
-fit <- predict(glm_Ridge, test.x)
+fit <- predict(glm_ridge, test.x)
 true <- test.y

### 2022-11-17 src/draft.Rmd
probably final result

@@
 title: "Draft Documents"
-author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
+author: "Team 1: Le Van Minh, Callista Stephine Yu, Muhammad Imtiyaaz B M A M , Thum Jun Long Issac"
 date: "`r Sys.Date()`"
@@
 
-## 3.2. Selectively remove variables with high missing rate (Callista)
+## 3.2. Selectively remove variables with high missing rate
 
@@
 # 4. Conclusion
+
+Approach 1, 2 and 4 seem to have the best performace.
+
+First Approach:
+```{r}
+knitr::kable(summary_3models[1:3,-5], digits = 3)
+knitr::kable(summary_3models[4:6,-5], digits = 3)
+```
+
+Second Approach:
+
+```{r}
+knitr::kable(summary_lasso, digits = 3)
+knitr::kable(summary_Ridge, digits = 3)
+```
+
+Fourth Approach:
+```{r}
+kable(result.df[c('alpha', 'lambda', 'Train.R2', 'Test.R2')])
+```
+
+We can deduce our final interpretation from these models.
+
+```{r}
+var <- c('incomeL', 'incomeLM', 'incomeUM', 'edu.total', 'hlth', 'mil')
+result1 <- coefficients(model3)[var]
+# result1[names(scaler2)] <- result1[names(scaler2)] / scaler2[names(scaler2)]
+# result1 <- result1 * scaler2['pov']
+# result1 <- result1[var]
+
+result2 <- coefficients(glm_Ridge)[,'s0']
+result2[names(scaler2)] <- result2[names(scaler2)] / scaler2[names(scaler2)]
+result2 <- result2 * scaler2['pov']
+result2 <- result2[var]
+
+result4 <- res[2,var]
+
+# combine
+finalResults <- data.frame(Approach = c(1,2,4))
+finalResults <- finalResults %>% 
+  cbind(rbind(result1, result2, result4))
+
+kable(finalResults)
+```
+
+Some important interpretation:
+
+- Low income countries generally have higher poverty index than High income countries.
+- As suggested by 1st approach, Lower-Middle income countries have higher poverty index than High income countries. While 2nd and 4th approach suggest otherwise.
+- Upper-Middle income countries in general have lower poverty index than High income countries.
+
+As suggested by these models.
+Upper-Middle income countries have lowest poverty rate.
+This can be attributed to that these countries tend to have more even wealth distribution than high income countries, and have enough money to alleviate poor people.
+
+- Spending on education is negatively correlated with poverty.
+On average, 1% (of GDP) more spending in education results in 3 point decrease in poverty index.
+This suggests that public spending on education can reduce poverty.
+- Spending on Healthcare is negatively correlated with poverty. On average, 1% (of GDP) more spending in education results in 0.5 point decrease in poverty index.
+- Spending on military has ambiguous effect on poverty.
+Different model results in different interpretation.
+This can be accounted to the difference in the data used in each approach.
+Approach 1 only use complete cases, which caused the model to predict positive relationship.
+However, as more data is utilised in approach 2&4 use more data, the effect tend to be neutral/negative.
+
 # 5. References

### 2022-11-17 src/temp.Rmd
probably final result

--- 2022-11-17 src/draft.Rmd
@@
 ---
-title: "Draft Documents"
+title: "Temp"
 author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
 date: "`r Sys.Date()`"
-output:
-  pdf_document: default
-  html_document:
-    df_print: paged
-  word_document: default
-bibliography: citations.bib
-urlcolor: blue
+output: html_document
 ---
-
-```{r setup, include=FALSE}
-library(knitr)
-knitr::opts_chunk$set(echo = TRUE, warning = F)
-knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
-# inline hook to 2 digits
-inline_hook <- function(x) {
-    if(is.numeric(x)) x <- round(x, 2)
-    paste(as.character(x), collapse=", ")
-}
-knit_hooks$set(inline = inline_hook)
-# code chunk out put to 
-options(digits = 5)
-```
-
-```{r, message = F}
-library(needs)
-
-needs("dplyr", "MASS", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce", "lmtest")
-
-prioritize(dplyr)
-```
-
-# 1. Introduction (To be done)
-The project aim is to explore different models using the same dataset to evaluate which method will be the best to build a model for the dependent variable.
-The dataset we are using for this project has 1901 data values from 23 different variables.
-Our plan for this project is to first build a model with all the variables in the dataset.
-We will then follow it up by builidng a model without the missing values and build a ridge regression and Lasso regression model.
-Our second approach will be to selectively remove variables from the model and to build the best model  using regularisation and regression.
-Our next approach was to add the variables into the model one at a time to see if a different model can be achieved.
-Lastly, we performed imputation on our model to tackle the missing values present in our dataset.We then compared all of this approaches to come to a conclusion.
-
-# 2. Data Characteristic
-
-## 2.1. Nature of Data
-
-The data set is collection The World Bank Data, the variables of interest are extracted from the raw data files and combined into a single data frame for analysis. The final data set includes:
-
-1.  **country.code**: Country code\
-2.  **country.name**: Country name\
-3.  **year**: Year\
-4.  **income**: Income class
--   Low income (L)\
--   Lower middle income (LM)\
--   Upper middle income (UM)\
--   High income (H)\
-5.  **reg**: Region\
-6.  **pov**: Poverty headcount ratio based on cut-off value of $2.15 per day\
-7.  **mpi**: Multidimensional Poverty Index\
-8.  **edu.total**: Total expenditure on education (% of GDP)\
-9.  **edu.pri**: Total expenditure on primary education (% of total education expenditure)\
-10. **edu.sec**: Total expenditure on secondary education (% of total education expenditure)\
-11. **edu.ter**: Total expenditure on tertiary education (% of total education expenditure)\
-12. **hlth**: Total expenditure on health (% of GDP)\
-13. **mil**: Total expenditure on military (% of GDP)\
-14. **fdi**: Foreign Direct Investment\
-15. **lbr.part**: Labour force participation (% of population ages 15+)\
-16. **unemp**: Unemployment rate\
-17. **pop.gwth.total**: Total population growth rate\
-18. **pop.gwth.rural**: Total rural population growth rate\
-19. **pop.gwth.urban**: Total urban population growth rate\
-20. **gdp.dflt**: GDP deflator\
-21. **gdr.eql**: Gender equality rating\
-22. **gcf**: Gross Capital Formation\
-23. **trade**: Trade = import + export (% of GDP)
-24. **gdp.pc**: GDP per capita (current US$)
-
-Data imports and combining:
-
-```{r}
-# helper functions
-importWDI <- function(filepath, value_name) {
-  df <- read_csv(filepath, skip = 4)
-
-  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))
-
-  df <- df %>% 
-    pivot_longer(5:ncol(.), names_to = "year", values_to = "value") %>% 
-    filter(!is.null(value) & !is.na(value)) %>% 
-    mutate(country.code = factor(country.code), 
-           country.name = factor(country.name),
-           year = as.numeric(year)) %>% 
-    select(country.code, country.name, year, value)
-
-  colnames(df)[4] <- value_name
-
-  df
-}
-
-importRegionClass <- function(filepath) {
-  df <- read_csv(filepath, skip = 4)
-
-  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))
-
-  df %>% mutate(country.name = factor(country.name),
-                region = factor(region)) %>% 
-    select(country.name, reg = region)
-}
-
-importIncomeClass <- function(filepath) {
-  df <- read_csv(filepath, skip = 4)
-
-  colnames(df) <- tolower(gsub(" ", ".", colnames(df)))
-
-  df %>% 
-    pivot_longer(3:ncol(.), names_to = "year", values_to = "income") %>% 
-    filter(!is.null(income) & !is.na(income)) %>% 
-    mutate(country.code = factor(country.code), 
-           country.name = factor(country.name), 
-           year = as.numeric(year), 
-           income = factor(income)) %>% 
-    select(country.code, country.name, year, income)
-}
-```
-
-```{r, message = F}
-# import data
-setwd("../data")
-
-poverty.headcount <- importWDI("poverty.headcount.215dollar.csv", "pov")
-mpi <- importWDI("mpi.csv", "mpi")
-education.expenditure.total <- importWDI("total.education.expenditure.csv", "edu.total")
-education.expenditure.primary <- importWDI("primary.education.expenditure.csv", "edu.pri")
-education.expenditure.secondary <- importWDI("secondary.education.expenditure.csv", "edu.sec")
-education.expenditure.tertiary <- importWDI("tertiary.education.expenditure.csv", "edu.ter")
-health.expenditure <- importWDI("health.expenditure.csv", "hlth")
-military.expenditure <- importWDI("military.expenditure.csv", "mil")
-fdi <- importWDI("fdi.csv", "fdi")
-labour.force.participation <- importWDI("labour.force.participation.csv", "lbr.part")
-unemployment.rate <- importWDI("unemployment.csv", "unemp")
-population.growth <- importWDI("population.growth.csv", "pop.gwth.total")
-rural.population.growth <- importWDI("rural.population.growth.csv", "pop.gwth.rural")
-urban.population.growth <- importWDI("urban.population.growth.csv", "pop.gwth.urban")
-gdp.deflator <- importWDI("gdp.deflator.csv", "gdp.dflt")
-gender.equality <- importWDI("gender.equality.csv", "gdr.eql")
-gross.capital.formation <- importWDI("gross.capital.formation.csv", "gcf")
-trade <- importWDI("trade.csv", "trade")
-region.class <- importRegionClass("region.class.csv")
-income.class <- importIncomeClass("income.class.csv")
-gdp.pc <- importWDI("gdp.pc.csv", "gdp.pc")
-
-setwd("../src")
-```
-
-We found that the data sets collected from [World Bank's Data helpdesk](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups) and [The World Bank's Data](https://data.worldbank.org/) have different naming convention for certain countries (e.g.
-"Czechia" vs.
-"Czechnia Republic").
-So we need to rename these countries to avoid some error when joining.
-
-Furthermore, WDI's data sets rate also account for non-country (e.g.
-country.name = "Low income" or "South Asia").
-These special groups are not in our scope of interest, which is national, so we eliminate them.
-
-```{r}
-# using poverty.headcount as a naming standard (as other data from WDI also use this convention)
-# join a subset of data to process the names
-d <- poverty.headcount %>%
-  select(country.name) %>%
-  mutate(inPov = T) %>%  
-  full_join(income.class %>% 
-              select(country.name) %>% 
-              mutate(inIncome = T), by = "country.name") %>%  
-  full_join(region.class %>%
-              select(country.name) %>% 
-              mutate(inReg = T), by = "country.name") %>% 
-  mutate(inPov = !is.na(inPov), inIncome = !is.na(inIncome), inReg = !is.na(inReg))
-
-d
-```
-
-First, remove special economic groups from `poverty.headcount`.
-We figured these regions will not appear in `income.class` or `region.class`, so we might find something from looking at the countries **only** appear in `poverty.headcount`.
-
-```{r}
-d %>% filter(inPov & (!inIncome | !inReg)) %>% distinct(country.name)
-```
-
-Lucky! We can look through these 18 results and compose a list of special regions.
-
-```{r}
-spec.reg <- c("Fragile and conflict affected situations", "IDA total", "World", "East Asia & Pacific", "Europe & Central Asia", "Latin America & Caribbean", "Middle East & North Africa", "South Asia", "Sub-Saharan Africa", "Low income", "Low & middle income", "Lower middle income", "Upper middle income", "High income")
-```
-
-Then, we rename those countries with inconsistent naming convention. Since we should only care about countries whose poverty headcount is available, reusing the list generated above, we can identify:
-
-1.  Cote d'Ivoire (also Côte d'Ivoire)
-2.  Czechia (also Czechoslovakia or Czech Republic)
-3.  Curacao (also Curaçao)
-4.  Turkiye (formerly known as Turkey, also Türkiye)
-5.  Sao Tome and Principe (also São Tomé and Príncipe)
-
-```{r}
-# mapping standard name and variation
-nameMap <- tibble(standard = c("Cote d'Ivoire", "Czechia", "Czechia", "Curacao", "Turkiye", "Turkiye", "Sao Tome and Principe"), variation = c("Côte d'Ivoire", "Czechoslovakia", "Czech Republic", "Curaçao", "Turkey", "Türkiye", "São Tomé and Príncipe"))
-
-correctName <- function(name) {
-  tibble(name = name) %>% 
-    left_join(nameMap, by = c("name" = "variation")) %>% 
-    mutate(standard = ifelse(is.na(standard), name, standard)) %>% 
-    select(standard) %>% 
-    pull()
-}
-
-orig.name <- c("Vietnam", "China", "Turkey", "Czechia Republic")
-correctName(orig.name)
-```
-
-Let's test this out!
-
-```{r}
-d <- poverty.headcount %>%
-  # correct name here
-  mutate(country.name = correctName(country.name)) %>% 
-  select(country.name) %>%
-  mutate(inPov = T) %>%
-  full_join(income.class %>% 
-              # correct name here
-              mutate(country.name = correctName(country.name)) %>% 
-              select(country.name) %>% 
-              mutate(inIncome = T), by = "country.name") %>%  
-  full_join(region.class %>%
-              # correct name here
-              mutate(country.name = correctName(country.name)) %>% 
-              select(country.name) %>% 
-              mutate(inReg = T), by = "country.name") %>% 
-  filter(!(country.name %in% spec.reg)) %>% 
-  mutate(inPov = !is.na(inPov), inIncome = !is.na(inIncome), inReg = !is.na(inReg))
-
-# countries not in region list, but is in Pov list
-d %>% 
-  filter(!inReg & inPov) %>% 
-  distinct(country.name) %>% 
-  nrow()
-# countries not in income list, but is in Pov list
-d %>% 
-  filter(!inIncome & inPov) %>% 
-  distinct(country.name) %>% 
-  nrow()
-```
-
-We are *pretty* confident that there's no inconsistent naming left unprocessed in the data sets.
-
-```{r}
-# Rename countries in all data sets.
-poverty.headcount <- poverty.headcount %>%
-  mutate(country.name = correctName(country.name))
-mpi <- mpi %>%
-  mutate(country.name = correctName(country.name))
-education.expenditure.total <- education.expenditure.total %>%
-  mutate(country.name = correctName(country.name))
-education.expenditure.primary <- education.expenditure.primary %>%
-  mutate(country.name = correctName(country.name))
-education.expenditure.secondary <- education.expenditure.secondary %>%
-  mutate(country.name = correctName(country.name))
-education.expenditure.tertiary <- education.expenditure.tertiary %>%
-  mutate(country.name = correctName(country.name))
-health.expenditure <- health.expenditure %>%
-  mutate(country.name = correctName(country.name))
-military.expenditure <- military.expenditure %>%
-  mutate(country.name = correctName(country.name))
-fdi <- fdi %>%
-  mutate(country.name = correctName(country.name))
-labour.force.participation <- labour.force.participation %>%
-  mutate(country.name = correctName(country.name))
-unemployment.rate <- unemployment.rate %>%
-  mutate(country.name = correctName(country.name))
-population.growth <- population.growth %>%
-  mutate(country.name = correctName(country.name))
-rural.population.growth <- rural.population.growth %>%
-  mutate(country.name = correctName(country.name))
-urban.population.growth <- urban.population.growth %>%
-  mutate(country.name = correctName(country.name))
-gdp.deflator <- gdp.deflator %>%
-  mutate(country.name = correctName(country.name))
-gender.equality <- gender.equality %>%
-  mutate(country.name = correctName(country.name))
-gross.capital.formation <- gross.capital.formation %>%
-  mutate(country.name = correctName(country.name))
-trade <- trade %>%
-  mutate(country.name = correctName(country.name))
-region.class <- region.class %>%
-  mutate(country.name = correctName(country.name))
-income.class <- income.class %>%
-  mutate(country.name = correctName(country.name))
-gdp.pc <- gdp.pc %>%
-  mutate(country.name = correctName(country.name))
-```
-
-Join the data
-
-```{r, message = F}
-countries <- poverty.headcount %>% 
-  # We used a full join here so we can conduct a separate analysis on mpi later
-  full_join(mpi, by = c("country.name", "country.code", "year")) %>%
-  left_join(income.class, c("country.name", "country.code", "year")) %>% 
-  left_join(region.class, by = "country.name") %>%
-  left_join(education.expenditure.total, by = c("country.name", "country.code", "year")) %>%
-  left_join(education.expenditure.primary, by = c("country.name", "country.code", "year")) %>%
-  left_join(education.expenditure.secondary, by = c("country.name", "country.code", "year")) %>%
-  left_join(education.expenditure.tertiary, by = c("country.name", "country.code", "year")) %>%
-  left_join(health.expenditure, by = c("country.name", "country.code", "year")) %>%
-  left_join(military.expenditure, by = c("country.name", "country.code", "year")) %>%
-  left_join(fdi, by = c("country.name", "country.code", "year")) %>%
-  left_join(labour.force.participation, by = c("country.name", "country.code", "year")) %>%
-  left_join(unemployment.rate, by = c("country.name", "country.code", "year")) %>%
-  left_join(population.growth, by = c("country.name", "country.code", "year")) %>%
-  left_join(rural.population.growth, by = c("country.name", "country.code", "year")) %>%
-  left_join(urban.population.growth, by = c("country.name", "country.code", "year")) %>%
-  left_join(gdp.deflator, by = c("country.name", "country.code", "year")) %>%
-  left_join(gender.equality, by = c("country.name", "country.code", "year")) %>%
-  left_join(gross.capital.formation, by = c("country.name", "country.code", "year")) %>%
-  left_join(trade, by = c("country.name", "country.code", "year")) %>% 
-  left_join(gdp.pc, by = c("country.name", "country.code", "year")) %>% 
-  # filter special groups
-  filter(!(country.name %in% spec.reg))
-```
-
-Data preview
-
-```{r}
-head(countries)
-str(countries)
-summary(countries)
-```
-
-There's no NA in `reg`, which is a sign that all naming in the data is remedied.
-There's some expected NAs in `income` and `pov`, as these data are collected by year.
-There's a substantial amount of missing data in `mpi`, as this is a relative new concept.
-We will address the nature, and processing of missing data in the next sections.
-
-## 2.2. Missing values
-
-As observed from the summary above, the data set contains a lot of missing values in some of the variables.
-
-```{r}
-mean(is.na(countries))
-```
-
-About 19% of the data set is missing.
-
-```{r}
-nCompleteObs <- sum(complete.cases(countries))
-print(paste("No. of complete cases:", nCompleteObs))
-```
-
-There are only 3 complete cases where all the variable is available.
-This is nowhere near acceptable to conduct any meaningful analysis.
-Therefore, we need to eliminate some variables for a more balance data set.
-
-```{r}
-missings <- colMeans(is.na(countries))
-ggplot(mapping = aes(x = names(missings), y = missings, fill = missings < 0.35)) +
-  geom_bar(stat = "identity") +
-  ggtitle("Missing rate") +
-  xlab("variables") +
-  ylab("% missing") +
-  theme(axis.text.x = element_text(size=9, angle=90))
-
-missings[missings > 0.35]
-```
-
-There are 5 variables with missing rate \>35%.
-
-expenditure in primary, secondary, and tertiary edication can be very useful and relevant information to predict poverty reduction [@pubspend&pov].
-However, we would like to exclude these variables from some first analyses to make use of the richer set of data.
-We can conduct a separate analysis with these variable to gain more insight.
-
-```{r}
-# variables with high missing rate
-hMiss <- names(missings[missings > 0.35])
-# exclude these variables in countries1
-countries1 <- countries %>% 
-  select(!hMiss) %>% 
-  filter(!is.na(pov))
-str(countries1)
-```
-
-Re-evaluate the `countries1` set.
-
-```{r}
-mean(is.na(countries1))
-sum(complete.cases(countries1))
-mean(complete.cases(countries1))
-```
-
-On average, each column has 6% missing rate, results in 937 complete data point (i.e.
-49%).
-This can be a sufficient number for the analysis.
-However, the missing data can induce loss of power due to the reduced sample size, and some other biases depending on which variables is missing.
-
-```{r}
-# complete rate of data by regions
-countries1 %>% 
-  mutate(isComplete = complete.cases(.)) %>% 
-  group_by(reg) %>% 
-  summarise(complete.rate = mean(isComplete)) %>% 
-  arrange(desc(complete.rate))
-```
-
-Countries from North America, Sub-Saharan Africa, and South Asia have the highest rate of missing data.
-We suspect that Sub-Saharan Africa, and South Asia are comparably less accessible regions.
-We also know that Americans don't like filling out forms, so their high rate of missing data is understandable as well.
-
-Still, we need to find a way to address this issue. we propose several approaches:
-
-1.  **Use complete cases**: Only use the complete cases for the analysis. This is a straightforward approach, but doesn't resolve the bias resulted from the mass loss of data.
-2.  **Selectively remove variables with high missing rate**: The same as we did before, but this process should be carried out carefully as we run the chance of dropping an important variable.
-3.
-**Update the data set as we select variables**: As we drop insignificant variables (in backward selection), the number of NAs are changed as well.
-We can utilize the extra complete cases to build the next model in the steps.
-4.
-**Imputation**: The idea is to replace the missing observations on the response or the predictors with artificial values that try to preserve the data set structure.
-This is a quite complex topic of its own, but we think why not.
-You can read more from @imputation.
-
-## 2.3. Descriptive Analytics
-Distribution of the predicted variable `pov`
-```{r}
-ggplot(countries, aes(x = pov)) +
-  geom_histogram(aes(y = ..density..), bins = 30) +
-  geom_density(alpha = .2, fill = "blue")
-```
-The graph displays a decreasing rate as poverty indicator increasing.
-This might not be representative of the current state of poverty in the world, but of the number presented in our data.
-For example, more recent data is likely to be more inclusive than ancient data, when poverty is more prevalent.
-We should look at data from the same period.
-
-```{r}
-# number of data point available from 1967 to 2021
-ggplot(poverty.headcount, aes(x = year)) +
-  geom_histogram(bins = 30)
-
-# pov data from 1998 and 2018
-pov.98.18 <- poverty.headcount %>%
-  filter(year == 1998 | year == 2018)
-pov.98.18 %>% 
-  group_by(year) %>% 
-  summarise(sum = n())
-
-ggplot(pov.98.18, aes(x = pov)) +
-  geom_histogram(aes(y = ..density..), bins = 30) +
-  geom_density(alpha = .2, fill = "blue") +
-  facet_grid(cols = vars(year))
-```
-
-The graph for 1998 has a much gentler slope, meaning poverty was more popular during that time, as predicted from out intuition. What about the general progress of the world?
-
-```{r}
-# Re-import pov and only take special regions
-# geographic
-geo.regs <- c("EAS", "ECS", "LCN", "MEA", "SAS", "SSF", "WLD")
-# economics
-eco.regs <- c("HIC", "LIC", "LMC", "LMY", "UMC")
-
-pov.reg <- importWDI("../data/poverty.headcount.215dollar.csv", "pov") %>% 
-  filter(country.code %in% c(geo.regs, eco.regs)) %>% 
-  mutate(type = ifelse(country.code %in% geo.regs, "Geographic", "Economics"))
-
-pov.reg %>%
-  distinct(country.code, country.name, type) %>% 
-  arrange(type)
-
-# World
-ggplot(pov.reg %>% filter(country.code == "WLD"),
-       aes(x = year, y = pov)) +
-  geom_point() +
-  geom_smooth(formula = y~x, method = loess)
-```
-
-An overall very steady decrease of poverty. How about each region?
-
-```{r}
-ggplot(pov.reg %>% filter(country.code != "WLD"),
-       aes(x = year, y = pov, color = country.name)) +
-  geom_point() +
-  geom_smooth(formula = y~x, method = loess) +
-  facet_grid(rows = vars(type))
-```
-
-There's a general steady, but distinct decline of poverty over time in each type region of respective type.
-Latin America & Caribbean, Europe & Central Asia, Middle East & North Africa, and High Income group has a more gradual decline as they are not very poor to begin with.
-
-Among the income groups, Low & Middle Income, Lower & Middle Income, and Upper Middle Income have quite similar in term of poverty indicator and slope over the year.
-While these values vary greatly among different geographical regions.
-
-Let's see some important statistics
-
-```{r}
-stats <- poverty.headcount %>% 
-  summarise(count = n(), skewness = skewness(pov), kurtosis = kurtosis(pov), std.deviation = sd(pov))
-
-kable(stats)
-```
-
-## 2.4. Assumption Check
-We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data.
-[Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
-**1. Linear relationship:**  
-Use correlation matrix to check linearity
-```{r, warning = FALSE}
-# select only numeric data
-countries.num <- countries %>% select(where(is.numeric))
-chart.Correlation(countries.num, histogram = TRUE, pch = 19)
-```
-Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
-```{r}
-# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
-countries.num2 <- countries.num %>% 
-  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
-  filter(!is.na(value) & !is.na(pov))
-
-drawGraph <- function(indvar, data) {
-  ggplot(data %>% 
-           filter(variable == indvar),
-         aes(x = value, y = pov)) +
-    geom_point() +
-    geom_smooth(method = lm) +
-    labs(title = paste("Relationship pov ~", indvar),
-         x = indvar,
-         y = "pov")
-}
-
-ivs <- distinct(countries.num2, variable)[[1]]
-
-# for (indvar in ivs) {
-#   print(
-#     ggplot(countries.num2 %>% 
-#            filter(variable == indvar),
-#          aes(x = value, y = pov)) +
-#     geom_point() +
-#     geom_smooth(formula = y~x, method = lm) +
-#     labs(title = paste("Relationship pov ~", indvar),
-#          x = indvar,
-#          y = "pov")
-#   )
-# }
-
-p <- ggplot(countries.num2, 
-     aes(x = value, y = pov)) +
-geom_point() +
-geom_smooth(formula = y~x, method = lm) +
-labs(title = paste("pov ~ variable"),
-     x = "variable",
-     y = "pov") +
-facet_wrap_paginate(~ variable,
-                    ncol = 3,
-                    nrow = 1,
-                    scales = "free")
-
-requiredPages <- n_pages(p)
-
-for (i in 1:requiredPages) {
-  ggplot(countries.num2,
-         aes(x = value, y = pov)) +
-  geom_point() +
-  geom_smooth(formula = y~x, method = lm) +
-  labs(title = paste("pov ~ variable"),
-       x = "variable",
-       y = "pov") +
-  facet_wrap_paginate(~ variable,
-                      ncol = 3,
-                      nrow = 1,
-                      scales = "free",
-                      page = i) -> p
-  print(p)
-}
-```
-
-Most variables display linear relationship with some exceptions, which are more appropriate to assume a logarithmic relationship. Gender equality should be viewed as a categorical data.
-```{r}
-logIvs <- c("fdi", "gdp.dflt", "gdp.pc")
-
-for (indvar in logIvs) {
-  print(
-    ggplot(countries.num2 %>% 
-           filter(variable == indvar),
-         aes(x = log(value), y = pov)) +
-    geom_point() +
-    geom_smooth(formula = y ~ x, method = lm) +
-    labs(title = paste0("pov ~ log(", indvar, ")"),
-         x = paste0("log(", indvar, ")"),
-         y = "pov")
-  )
-}
-```
-
-The relationship are more linear after the transformation.
-
-```{r}
-gdr.eql <- countries.num2 %>% 
-  filter(variable == "gdr.eql") %>% 
-  mutate(value = factor(value))
-
-ggplot(gdr.eql,
-       aes(x = value, y = pov)) +
-  geom_boxplot() +
-  labs(title = paste("pov ~ gdr.eql"),
-       x = "gdr.eql",
-       y = "pov")
-```
-There's a significant correlation between gender equality and poverty.  
-Correlation coefficients between `pov` and predictors.
-
-```{r}
-# transform some variables
-countries.num3 <- countries.num %>% 
-  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
-  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
-
-# name of the independent variables
-cn <- colnames(countries.num3)[-c(1,2)]
-
-# correlation of independent variable and pov
-corWithPov <- function(indevar, data) {
-  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
-}
-
-cor.pov <- sapply(cn, corWithPov, data = countries.num3)
-kable(cor.pov)
-```
-
-`edu.ter`, `mil`, `fdi`, `lbr.part`, `unemp`, `gdp.dflt`, `gcf` have negligible correlation with `pov`.
-`lgdp.pc`, `lfdi`, and `lgdp.dflt` have stronger linear relationship with `pov` than their un-transformed counterparts.
-
-```{r}
-countries1 <- countries1 %>% 
-  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
-  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
-```
-
-**2. Predictors are independent**
-There should be no correlation/multicolinearity between each pair of predictors.
-
-```{r}
-countries.arr <- simplify2array(countries.num %>% 
-                                  select(-c("year", "pov")))
-
-cor.mtx <- rcorr(countries.arr, type = "spearman")
-
-round(cor.mtx$r, 2)
-
-corrplot(cor.mtx$r, type = "upper", order = "hclust", 
-         tl.col = "black", tl.srt = 45)
-```
-Some significant correlations can be found between `gdr.eql` with `edu.pri`, `pop.gwth.total`, and `fdi` with `mil`, etc.
-We expect these variables to be eliminated in the VIF test.
-This might be a good property for imputation (3.2.4)
-
-## 2.5. Data Source
-
-- [poverty.headcount](https://data.worldbank.org/indicator/SI.POV.DDAY)
-- [mpi](https://data.worldbank.org/indicator/SI.POV.MDIM.XQ)
-- [education.expenditure.primary](https://data.worldbank.org/indicator/SE.XPD.PRIM.ZS)
-- [education.expenditure.secondary](https://data.worldbank.org/indicator/SE.XPD.SECO.ZS)
-- [education.expenditure.tertiary](https://data.worldbank.org/indicator/SE.XPD.TERT.ZS)
-- [education.expenditure.total](https://data.worldbank.org/indicator/SE.XPD.TOTL.GD.ZS)
-- [health.expenditure](https://data.worldbank.org/indicator/SH.XPD.GHED.GD.ZS)
-- [military.expenditure](https://data.worldbank.org/indicator/MS.MIL.XPND.GD.ZS)
-- [fdi](https://data.worldbank.org/indicator/BX.KLT.DINV.CD.WD)
-- [unemployment.rate](https://data.worldbank.org/indicator/SL.UEM.TOTL.NE.ZS)
-- [labour.force.participation](https://data.worldbank.org/indicator/SL.TLF.CACT.NE.ZS)
-- [gender.equality](https://data.worldbank.org/indicator/IQ.CPA.GNDR.XQ)
-- [population.growth](https://data.worldbank.org/indicator/SP.POP.GROW)
-- [urban.population.growth](https://data.worldbank.org/indicator/SP.URB.GROW)
-- [rural.population.growth](https://data.worldbank.org/indicator/SP.RUR.TOTL.ZG)
-- [gdp.deflator](https://data.worldbank.org/indicator/NY.GDP.DEFL.KD.ZG)
-- [gross capital formation](https://data.worldbank.org/indicator/NE.GDI.TOTL.ZS)
-- [trade](https://data.worldbank.org/indicator/NE.TRD.GNFS.ZS)
-- [region.class](https://datatopics.worldbank.org/world-development-indicators/the-world-by-income-and-region.html)
-- [income.class](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)
-- [gross.capital.formation](https://data.worldbank.org/indicator/NE.GDI.TOTL.ZS)
-
-# 3. Model Selection and Interpretation
-
-We can now conduct linear regressions following the approaches mentioned above to address missing values issues.
-
-## 3.1. Use Complete Cases (Imtiyaaz)
-```{r}
-library(tidyverse) # for data manipulations and visualisations
-library(ggpubr) # for correlation coefficient on plot
-library(gridExtra) # for arranging plots 
-library(e1071) # for skewness
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-library(readxl)
-library(corrplot)
-library(car)
-library(DescTools)
-library(glmnet)
-```
-
-First we will write the function "eval_results(true,predict) to evaluate the models
-```{r}
-eval_results <- function(true, predict) {
-  actual <- data.matrix(true)
-  SSE <- sum((predict - actual)^2)
-  SST <- sum((actual - mean(actual))^2)
-  R_square <- 1 - SSE / SST
-  data.frame(
-    MSE = MSE(predict, true),
-    MAE = MAE(predict, true),
-    RMSE = RMSE(predict, true),
-    Rsquare = R_square
-  )
-}
-```
-First we read the csv file for the countries containing the dataset. We will then put the variables we need in to subset1.
-```{r}
-df <-read.csv('../data/countries1.csv')
-df
-subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
-subset1
-```
-
-Next we will build a SLR model. We will put 80% of dataset to train and the rest to test.
-```{r}
-set.seed(10)
-index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
-train <- df[index, ]
-test <- df[-index, ]
-```
-
-The model output will be placed in lm.slr and can be viewed in summary(lm.slr).
-```{r}
-lm.slr <- lm(pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
-summary(lm.slr)
-
-```
-Next we will be performing standardisation.However not all the variables are numerical.
-Therefore to overcome it we build the function below.
-This will allow us to perform standardisation to the dataset.
-```{r}
-subset1$income <- factor(subset1$income)
-subset1$country.code <- factor(subset1$country.code)
-
-sd0 <- function(vct) {
-  if (!is.numeric(vct)) {
-    return(NA)
-  }
-
-  return(sd(vct, na.rm = T))
-}
-
-std0 <- function(vct, scl) {
-  if (!is.numeric(vct)) {
-    return(vct)
-  }
-
-  return(vct/scl)
-}
-
-scale <- function(data) {
-  unlist(lapply(data, sd0))
-}
-
-standardise <- function(data) {
-  scaler <- scale(data)
-
-  numCols <- which(unlist(lapply(data, is.numeric)))
-  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
-  fct <- data[,-numCols]
-  return(cbind(fct, num))
-}
-```
-
-Next we will perform standardistation to the dataset.
-```{r}
-str(subset1)
-df0 <- subset1
-subset1
-apply(subset1, 2, sd)
-
-```
-
-Next we will plot a pov vs income scatter plot to see the effect of income on pov.
-From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov.
-This is consistent with our intuition that countries that are richer do not have as much poverty issue as the poorer countries.
-```{r}
-ggplot(data = subset1, aes(x = income, y = pov)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  labs(title = "Pov vs. Income ",
-       x = "income",
-       y = "pov")
-
-```
-Next we will build a MLR model with data from subset 1 to see the relationship between dependent variable:pov and the other variables
-
-```{r}
-model1 <- lm(pov ~ ., data = subset1)
-summary(model1)
-vif(model1)
-
-```
-Next we will perform regularisation. We will split the data to predictor variable "train.x" and response variables "train.y"
-```{r}
-subset2 <- subset(subset1,select =c("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
-subset2 <- na.omit(subset2)
-subset2
-train.x <- data.matrix(subset(subset2, select = -pov))
-train.y = data.matrix(subset2$pov)
-```
-
-Next we will perform splitting of data to train and test.
-```{r}
-split_train_test <- function(data) {
-set.seed(1984)
-# split data
-seeds <- data %>% 
-  group_by(country.code) %>% 
-  filter(row_number() == 1)
-
-plants <- data %>% 
-  group_by(country.code) %>% 
-  filter(row_number() != 1)
-
-isComplete <- which(complete.cases(plants))
-idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
-
-training <- plants[-idx,] %>% 
-  rbind(seeds)
-test <- plants[idx,]
-
-return(list(training, test))
-}
-
-training <- split_train_test(subset2)[[1]]
-test <- split_train_test(subset2)[[2]]
-
-ntrain <- 1:nrow(training)
-combine <- rbind(training, test)
-combine.mtrx <- model.matrix(pov ~ ., data = combine)
-train.mtrx <- combine.mtrx[ntrain,]
-test.mtrx <- combine.mtrx[-ntrain,]
-
-train.x <- train.mtrx[,-1]
-train.y <- training$pov
-
-test.x <- test.mtrx[,-1]
-text.y <- test$pov
-```
-
-We then build a MLR model with the training model to train it.
-
-```{r}
-model3 <- lm(pov ~., data = training)
-summary(model3)
-
-```
-
-```{r}
-fit <- model3$fitted.values
-true <- training$pov
-summary_MLR3_train <- eval_results(true, fit)
-```
-
-We then evaluate the model between train and test
-```{r}
-fit <- predict(model3, newdata=test)
-true <- test$pov
-summary_MLR3_test <- eval_results(true, fit)
-```
-
-We can see the evaluation of the model from the function below. From the train model, the R^squared value will be 0.965 and for the test model it will be 0.881.
-
-```{r}
-summary_MLR <- rbind(summary_MLR3_train[-5], 
-                     summary_MLR3_test[-5])
-rownames(summary_MLR) <- c("Baseline MLR_Train", 
-                           "Baseline MLR_Test" )
-knitr::kable(summary_MLR, digits = 3)
-```
-Next we will be performing ridge and lasso regression model using the following code. We will split the data to train and test.
-```{r}
-# Training Data
-data <- training
-train.x <- data.matrix(data[,-ncol(data)])
-train.y <- data.matrix(data[,ncol(data)])
-
-# Test Data
-data <- test
-test.x <- data.matrix(data[,-ncol(data)])
-test.y <- data.matrix(data[,ncol(data)])
-```
-
-Next we will perform the Ridge regression model.
-```{r}
-# Fit Ridge Regression Model
-set.seed(123)
-cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
-glm_ridge <- glmnet(train.x, train.y, 
-                    alpha = 0, lambda = cv_ridge$lambda.min)
-
-# Evaluate Ridge Regression Model on Training Set
-fit <- predict(glm_ridge, newx = train.x, type = "response")
-true <- train.y
-summary_ridge_train <- eval_results(true, fit)
-
-# Evaluate Ridge Regression Model on Test Set
-fit <- predict(glm_ridge, newx = test.x, type = "response")
-true <- test.y
-summary_ridge_test <- eval_results(true, fit)
-
-# Summarise Ridge Regression Model Performance into Table
-summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
-rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
-```
-
-Next we will perform Lasso Regression model
-
-```{r}
-# Fit LASSO Model
-set.seed(123)
-cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
-glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
-                    lambda = cv_lasso$lambda.min)
-
-# Evaluate LASSO Model on Training Set
-fit <- predict(glm_lasso, newx = train.x)
-true <- train.y
-summary_lasso_train <- eval_results(true, fit)
-
-# Evaluate LASSO Model on Test Set
-fit <- predict(glm_lasso, newx = test.x)
-true <- test.y
-summary_lasso_test <- eval_results(true, fit)
-
-# Summarise LASSO Model Performance into Table
-summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
-rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
-```
-
-We will compare the 3 models that we have build, Multi linear ,Ridge and Lasso regression models.
-We evaluate the 3 models together and compare both train and test models of each of the 3 regression model.
-
-```{r}
-summary_3models <- rbind(summary_MLR3_train, 
-                         summary_ridge_train,                               
-                         summary_lasso_train, 
-                         summary_MLR3_test,                             
-                         summary_ridge_test, 
-                         summary_lasso_test)
-
-rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
-                               "LASSO Model_Train",  "Baseline MLR_Test",  
-                               "Ridge Model_Test", "LASSO Model_Test" )
-
-knitr::kable(summary_3models[1:3,-5], digits = 3)
-knitr::kable(summary_3models[4:6,-5], digits = 3)
-
-```
-From the 3 models we can see that the original multi linear regression model is the best one with the highest R^squared value.
-However the Lasso Regression Model has the lowest MSE, MAE and RMSE.
-We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting as many values will get removed.
-Next we will build other models to see if they work better.
-
-## 3.2. Selectively remove variables with high missing rate (Callista)
-
-```{r}
-needs("knitr","readr","tidyr","dplyr","ggplot2","stats","tidyverse","e1071","car","glmnet","corrplot","plotmo")
-```
-
-```{r}
-countries <- read.csv("../data/countries.csv")
-```
-
-###  3.2.1. Model Fitting
-##### A. Ordinary Linear Regression
-Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables.
-This section attempts to build a model on variables that have missing rates of below 35%.
-
-The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models.
-Hence, we transformed the data with the mutate() function before building the machine learning model.
-```{r}
-
-subset0 <- subset(countries, select = c("country.code", "year","edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))
-
-subset1_0 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))
-
-```
-subset1 contains the variables below 35% of missing values. 
-```{r}
-subset1 <- subset(subset1_0, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
-
-subset1 <- na.omit(subset1)
-```
-We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step. 
-```{r}
-sd0 <- function(vct) {
-  if (!is.numeric(vct)) {
-    return(NA)
-  }
-
-  return(sd(vct, na.rm = T))
-}
-
-std0 <- function(vct, scl) {
-  if (!is.numeric(vct)) {
-    return(vct)
-  }
-
-  return(vct/scl)
-}
-
-scale <- function(data) {
-  unlist(lapply(data, sd0))
-}
-
-standardise <- function(data) {
-  scaler <- scale(data)
-
-  numCols <- which(unlist(lapply(data, is.numeric)))
-  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
-  fct <- data[,-numCols]
-  return(cbind(fct, num))
-}
-
-```
-
-```{r}
-
-df.countries0 <- standardise(subset1)
-df.countries <- df.countries0
-```
-
-```{r}
-#Divide each feature/target by its standard deviation
-
-head(df.countries)
-```
-
-First, we split the data into training and testing data. 
-```{r}
-set.seed(123)
-index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
-train_1 <- subset1[index,]
-test_1 <- subset1[-index,]
-
-```
-
-Our first attempt 
-```{r}
-#Everything
-lm_1 <- lm(data= train_1, formula = pov ~ .)
-summary(lm_1)
-```
-
-The summary of `lm_1` suggests that the model still contains predictor variables with pr(>|t|) > 0.05.
-We can imply that these variables do not provide a high significance to our model and are thus able to be removed.
-To avoid removing significant variables, we shall conduct the removal of the variables deliberately.
-In `lm_1`, we notice that the variable `lgdp.dflt` has a high Pr(>|t|).
-Hence, we remove `lgdp.dflt`.
-
-```{r}
-#remove hlth 
-lm_2 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
-summary(lm_2)
-plot(lm_2, which = 5)
-```
-
-```{r}
-#Remove lgdp.dflt
-lm_3 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
-summary(lm_3)
-plot(lm_3, which = 5)
-```
-```{r}
-#remove lfdi
-lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
-summary(lm_4)
-plot(lm_4, which = 5)
-```
-Our model shows that among all the 14 predictor variables that we started with, only 11 of them showed significance in building the model.
-
-We then evaluated the results of the model. 
-```{r}
-#evaluating lm_4
-eval.metrics.linreg <- function(actual, predicted) {
-  residual <- actual - predicted
-  mse <- mean(residual ^ 2)
-  mae <- mean(abs(residual))
-  rmse <-  sqrt(mse)
-  mape <- mean(abs(residual / actual)) * 100
-
-  data.frame(
-    MSE = mse,
-    MAE = mae,
-    RMSE = rmse,
-    MAPE = mape
-  )
-}
-```
-
-```{r}
-actual <- train_1$pov
-predicted <- predict(lm_4, newdata = train_1)
-eval.metrics.linreg(actual, predicted)
-```
-The MAPE shows `Inf` due to the presence of 0 values in the Y variable.
-
-##### B. Regularization
-Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model.
-
-Next, we will perform Ridge regression and LASSO regression in a dataset that contains `country.code` and `year` in addition to the previous dataset that we used for the previous linear regression model.
-```{r}
-#Subsetting the data
-
-data0 <- subset(subset1_0, select = c("country.code","year","edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))
-data <- na.omit(data0)
-train.x <- data.matrix(subset(data, select = -pov))
-train.y <- data.matrix(data$pov)
-
-```
-
-Splitting the aforementioned dataset into training and testing data.
-
-```{r}
-split_train_test <- function(data) {
-set.seed(1984)
-# split data
-seeds <- data %>% 
-  group_by(country.code) %>% 
-  filter(row_number() == 1)
-
-plants <- data %>% 
-  group_by(country.code) %>% 
-  filter(row_number() != 1)
-
-isComplete <- which(complete.cases(plants))
-idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
-
-training <- plants[-idx,] %>% 
-  rbind(seeds)
-test <- plants[idx,]
-
-return(list(training, test))
-}
-```
-
-```{r}
-training <- split_train_test(data)[[1]]
-test <- split_train_test(data)[[2]]
-```
-Further splitting the training and testing data into  `train.x` and `train.y`
-```{r}
-ntrain <- 1:nrow(training)
-combine <- rbind(training, test)
-combine.mtrx <- model.matrix(pov ~ ., data = combine)
-train.mtrx <- combine.mtrx[ntrain,]
-test.mtrx <- combine.mtrx[-ntrain,]
-
-train.x <- train.mtrx[,-1]
-train.y <- training$pov
-
-test.x <- test.mtrx[,-1]
-text.y <- test$pov
-```
-
-Building a model with the lambda that outputs the smallest MSE. 
-```{r}
-cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
-glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
-
-t(coef(glm_Ridge))
-```
-The ridge regression model shows that some countries have a higher influence on the model than others.
-At first glance,it seems like the Y variable, `pov`, depends heavily on whether or not it belongs to certain countries.To further illustrate the significance of `country.code`, we will plot the ridge regression graph for different lambda values.
-```{r}
-#generate sequence of lambda values 
-lambda <- 10^seq(-6, 2, length = 100)
-ridge_model = glmnet(train.x, train.y, alpha = 0, lambda = lambda)
-
-t(coef(ridge_model))
-```
-Plotting the coefficients against different lambda values.
-
-```{r}
-plot_glmnet(ridge_model)
-```
-
-Interesting! For a low lambda value, the model exhibits `cntr.MDG`, which represents Madagascar, followed by `cntr.ZMB`, which represents Zimbabwe,  as key predictor variables.
-This means that the model depends highly on whether the country is Madagascar or Zimbabwe, followed by some other significant countries.
-The other variables that previously defined the model is now undermined by the presence of the `country.code` variable and does not show high influence.
-
-We then proceeded to evaluate the results
-
-```{r}
-eval_results <- function(fit, true) {
-  actual <- data.matrix(true)
-  SSE <- sum((actual - fit)^2)
-  SST <- sum((actual - mean(actual))^2)
-  R_square <- 1 - SSE/SST
-  data.frame(
-    MSE = MSE(fit, true),
-    MAE = MAE(fit, true),
-    RMSE = RMSE(fit, true),
-    MAPE = MAPE(fit, true),
-    R2  = R_square
-
-  )
-}
-```
-
-```{r}
-fit <- predict(glm_Ridge, train.x)
-true <- train.y
-summary_Ridge_train <- eval_results(fit, true)
-summary_Ridge_train
-```
-The training dataset produced a much higher value compared to the regular multiple regression model shown above.
-
-Fitting the model to the test dataset,
-
-```{r}
-fit <- predict(glm_Ridge, test.x)
-true <- test.y
-summary_Ridge_test <- eval_results(fit, true)
-summary_Ridge_test
-```
-A slightly lower R^2 value, but still proves that the model is good enough.
-
-Below is the summary of the results of the training and testing dataset. 
-```{r}
-summary_Ridge <- rbind(summary_Ridge_train, summary_Ridge_test)
-rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
-knitr::kable(summary_Ridge, digits = 3)
-
-```
-Next, we attempted LASSO regression as well. The steps are essentially similar to that of Ridge regression. 
-```{r}
-#Train a LASSO regression model
-set.seed(123)
-cv_lasso <- cv.glmnet(train.x, train.y, alpha = 1, type.measure = "mse")
-```
-
-```{r}
-cv_lasso
-```
-
-```{r}
-plot(cv_lasso)
-```
-
-```{r}
-glm_Lasso <- glmnet(train.x, train.y, alpha = 1, lambda = cv_lasso$lambda.min)
-t(coef(glm_Lasso))
-```
-It is evident that the number of variables included in the model is less than that of ridge regression.
-This is in accordance with the nature of the LASSO regression model that reduces the coefficients of certain predictor variables to zero for model interpretability.
-
-```{r}
-lambda <- 10^seq(-6, 2, length = 100)
-lasso_model = glmnet(train.x, train.y, alpha = 1, lambda = lambda)
-
-t(coef(lasso_model))
-```
-```{r}
-plot_glmnet(lasso_model)
-```
-```{r}
-plot_glmnet(ridge_model)
-```
-
-```{r}
-fit2 <- predict(glm_Lasso, train.x)
-true2 <- train.y
-summary_Lasso_train <- eval_results(fit2, true2)
-summary_Lasso_train
-```
-
-```{r}
-plot_glmnet(lasso_model)
-```
-
-```{r}
-fit3 <- predict(glm_Lasso, test.x)
-true3 <- test.y
-summary_Lasso_test <- eval_results(fit3, true3)
-summary_Lasso_test
-```
-```{r}
-summary_lasso <- rbind(summary_Lasso_train, summary_Lasso_test)
-rownames(summary_lasso) <- c("LASSO_train", "LASSO_test")
-knitr::kable(summary_lasso, digits = 3)
-
-```
-###  3.2.2. Assessment
-
-This section recapitulates the three models that we have generated so far through the method of gradually removing variables from the original dataset.
-
-```{r}
-summary(lm_4)
-```
-
-Our first model, `lm_4`, generated with regular multiple linear regression exhibits the lowest R^2 value.
-The model shows that `incomeL` is the most significant predictor variable with the highest coefficient.
-One interesting observation is that `incomeLM` and `incomeUM` has a negative correlation with the pov level, in contrast to `incomeL`.
-This shows that countries that have lower income tend to have high `pov` value.
-On the other hand, countries that are classified to have lower middle income and upper middle income seem to have lower `pov` values, as shown by their negative coefficients in the regression model.
-
-Apart from income, some other variables that contribute more significantly to the model are `lgdp.pc`, with a coefficient of -3.13114, exhibiting a stronger negative correlation, and `pop.gwth.urban`, exhibiting a stronger positive correlation with a coefficient of 1.90680.
-This is counter-intuitive, as the notion is that population growth in cities will help drive positive economic change due to the number of productive people living in the city.
-The model also shows that `pop.gwth.rural` gives a less significant impact on `pov` as compared to  `pop.gwth.rural`.
-
-```{r}
-knitr::kable(summary_lasso, digits = 3)
-```
-
-```{r}
-knitr::kable(summary_Ridge, digits = 3)
-```
-The MSE for the LASSO regression model is slightly higher than that of the Ridge regression model.
-This to be expected, as the Ridge regression model aims for an improved accuracy while the LASSO model aims for an improved interpretability.
-Both models seem to exhibit very high R^2 value for both the training and testing dataset.
-
-```{r}
-plot_glmnet(lasso_model)
-```
-```{r}
-
-```
-
-```{r}
-plot_glmnet(ridge_model)
-```
-```{r}
-plot(lm_4, which = 5)
-```
-The `Residual vs Leverage` plot of `lm_4` suggests that heteroskedasticity might be present in the model.
-This means that the variance of the residuals is unequal over a range of observed values.
-This problem exists in the dataset rather than the models generated.
-
-## 3.3. Update the data set as we select variables (Isaac)
-###  3.3.1. Model Fitting 
-```{r}
-df_count<-read.csv("../data/countries1.csv")
-set.seed(10)
-index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
-train <- df_count[index, ]
-test <- df_count[-index, ]
-```
-
-Through forward selection, we will do feature selection. We will stop the forward selection when the adjusted r-value>0.75.
-
-Through iteration of different models. A linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201, thus it is selected.
-
-```{r cars}
-summary(lm(formula = pov ~ income,data = train)) #0.6201
-```
-
-Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value.
-reg as a predictor has the highest adjusted r-value=0.7181
-
-```{r}
-summary(lm(formula = pov ~ reg+income,data = train)) #0.7181
-```
-
-The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7492
-```{r}
-summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train)) #0.7492
-```
-
-The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
-```{r}
-summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7703
-```
-
-Fitting of multiple linear regression with predictor gdp.pc, income, reg and lgd.pc.
-```{r}
-lm1<-lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)
-summary(lm1)
-```
-
-###  3.3.2. Assessment
-Assessment of linear model lm1: 
-the predicted model has some NA values which are removed. The error rates are too high, therefore, this method of variable selection may not be suitable
-```{r}
-eval.metrics.linreg <- function(actual, predicted) {
-  residual <- actual - predicted
-  mse <- mean(residual ^ 2)
-  mae <- mean(abs(residual))
-  rmse <-  sqrt(mse)
-  mape <- mean(abs(residual / actual)) * 100
-
-  data.frame(
-    MSE = mse,
-    MAE = mae,
-    RMSE = rmse,
-    MAPE = mape
-  )
-}
-predicted<-predict(lm1,newdata=test)
-actual<-test$pov
-df23<-na.omit(data.frame(actual,predicted))
-ac<-df23$actual
-pr<-df23$predicted
-eval.metrics.linreg(ac,pr)
-```
 
@@
 ```{r}
+library(needs)
+
+needs("dplyr", "MASS", "readr", "tidyr","tibble", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce", "lmtest")
+
+countries1 <- read.csv("../data/countries1.csv") %>% 
+  select(-1) %>% 
+  mutate(country.code = factor(country.code),
+         country.name = factor(country.name),
+         reg = factor(reg),
+         income = factor(income)) %>% 
+  as_tibble()
+
 set.seed(1984)
@@
 isComplete <- which(complete.cases(plants))
-idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))
+idx <- sample(isComplete, replace = F, 0.3 * nrow(plants))
 
@@
 
+meth <- c(rep("",4), "cart", "", rep("cart", 16))
+names(meth) <- colnames(countries1)
+meth
+
 # run the algorithm, or read from the file we generated as this is a very time-consuming process
-# countries1.imputed <- mice(countries.train.4, m = 3, maxit = 20, method = meth)
+countries1.imputed <- mice(countries.train.4, m = 3, maxit = 20, method = meth)
 # saveRDS(countries1.imputed, 'countries1.imputed.RData')

### 2022-11-17 src/Group 1_Project Code.Rmd
sure this is the last commit

@@
 
-Interesting! For a low lambda value, the model exhibits `cntr.MDG`, which represents Madagascar, followed by `cntr.ZMB`, which represents Zambia,  as key predictor variables.
-This means that the model depends highly on whether the country is Madagascar or Zambia, followed by some other significant countries.
+Interesting! For a low lambda value, the model exhibits `cntr.MDG`, which represents Madagascar, followed by `cntr.ZMB`, which represents Zimbabwe,  as key predictor variables.
+This means that the model depends highly on whether the country is Madagascar or Zimbabwe, followed by some other significant countries.
 The other variables that previously defined the model is now undermined by the presence of the `country.code` variable and does not show high influence.

### 2022-11-18 src/Group 1_Project Code.Rmd
updated

@@
 
-t(coef(glm_Ridge))
+head(t(coef(glm_Ridge)))
 ```
@@
 glm_Lasso <- glmnet(train.x, train.y, alpha = 1, lambda = cv_lasso$lambda.min)
-t(coef(glm_Lasso))
+head(t(coef(glm_Lasso)))
 ```
@@
 
-t(coef(lasso_model))
+head(t(coef(lasso_model)))
 ```

### 2022-11-18 src/archives/All variables.Rmd
updated

--- 2022-11-17 src/All variables.Rmd
@@
 ---
+Introduction : The project aim is to explore different models using the same dataset to evaluate which method will be the best to build a model for the dependent variable.
+The dataset we are using for this project has 1901 data values from 23 different variables.
+Our plan for this project is to first build a model with all the variables in the dataset.
+We will then follow it up by builidng a model without the missing values and build a ridge regression and Lasso regression model.
+Our second approach will be to selectively remove variables from the model and to build the best model  using regularisation and regression.
+Our next approach was to add the variables into the model one at a time to see if a different model can be achieved.
+Lastly, we performed imputation on our model to tackle the missing values present in our dataset.We then compared all of this approaches to come to a conclusion.
+
 ```{r}
@@
 However the Lasso Regression Model has the lowest MSE, MAE and RMSE.
-We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting.
+We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting as many values will get removed.
 Next we will build other models to see if they work better.
